# Interpretations of Disjunction {#sempragLit}

```{r global_options2, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, 
                      fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=TRUE, 
                      message=F, sanitize = T)
```

## Introduction

Despite its modest appearance, the word *or* has been a major trouble maker for formal theories of meaning. Perhaps the main reason is that capturing its meaning has always seemed within the reach of formal accounts at the time, yet more examination has shown that *or* is much more complex than expected. Consequently, few words have managed to prove as influential as *or* in advancing formal semantics and pragmatics. This chapter is dedicated to the semantics and pragmatics of disjunction, and specifically *or* in English. My primary goal is to provide an overview of the empirical discoveries in the interpretation of *or* in adult language before moving to disjunction in child language acquisition.

In discussing linguistic meaning, often different words are used to refer to different types of meaning. Here I clarify my terminology. In what follows, I may use the terms "meaning", "implicature", "inference", "implication", and "interpretation". I use the term "meaning" to refer to "literal meaning": the meaning of words and word combinations in their most "literal", least context-dependent, and least enriched version. Most of the time my usage of "meaning" refers to "lexical meaning". I use "implicature" (conversational) when the meaning is considered to be derived from literal meaning via the Gricean cooperative principle and its associated conversational maxims [@grice1989studies]. The term "inference" is used more generally to refer to any type of meaning that is derived from literal meaning using some form of reasoning. I use "implication" to refer to all linguistic meaning regardless of the source and theoretical status. Therefore, literal meanings, implicatures, and inferences are all implications. Finally, I use interpretation in the broadest and most theoretically neutral way to include all types of meaning; linguistic or non-linguistic. 

In what follows, I start with the discussion of disjunction in logic and philosophy. Disjunction has enjoyed a great deal of investigation in these fields and it would not be possible to understand current semantics and pragmatics of disjunction without first understanding its logical background. Then I move to the discussion of disjunction in language, and more specifically *or* in English. I show that since the early linguistic investigations, two different approaches to the meaning of disjunction have co-existed: first lexical ambiguity accounts, and second Gricean accounts. I explain that lexical ambiguity accounts consider *or* ambiguous between several meanings or senses while the Gricean accounts propose a single meaning for *or* and derive different interpretations of a disjunction from other linguistic and conversational factors. Then I enumerate some of the factors that affect the interpretation of disjunction. I argue that these factors provide a starting point for a cue-based acquisition of disjunction.

## Disjunction in Logic {#logic}

The first explicit account of propositional connectives in logic was developed by Stoic philosophers (3BC - 3AD) [@gabbay2012logic]. They divided logical propositions into simple (atomic) and complex (molecular) propositions. Complex propositions were made of simple propositions connected by connective such as conjunction and disjunction. Similar to logical systems today, they defined these connectives in terms of the truth conditions of the propositions they connect. A conjunction was considered true when all propositions were true and false otherwise. Disjunction, however, received two definitions: exclusive disjunction and inclusive disjunction. Exclusive disjunction was true when exactly one proposition was true and the rest were false. Inclusive disjunction was true when at least one proposition was true, the rest could be true or false. Table \@ref(tab:truthtable) shows the truth conditions for binary conjunction ($\land$), exclusive disjunction ($\oplus$), and inclusive disjunction ($\lor$)[^1].

[^1]: Binary exclusive disjunction has an "odd property"" in that linking a sequence of exclusive disjuncts results in a disjunction that is true if and only if odd number of disjuncts are true. This is not the case for the Stoic version of disjunction which is not a binary operator.

|$\phi$|$\psi$|$(\phi \land \psi)$|$(\phi \oplus \psi)$|$(\phi \lor \psi)$|
|:-:|:-:|:-:|:-:|:-:|
|T|T|T|F|T|
|T|F|F|T|T|
|F|T|F|T|T|
|F|F|F|F|F|
Table: (\#tab:truthtable) Truth conditions for conjunction and disjunction in classical logic

It is important to note that there are entailment relations between the connectives in Table \@ref(tab:truthtable). A proposition $\phi$ entails another proposition $\psi$ if in all situations that $\phi$ is true $\psi$ is true as well. Conjunction of two propositions entails their inclusive disjunction but not the other way round. When the conjunction is true (row 1), the inclusive disjunction is also true but there are situations where inclusive disjunction is true but the conjunction is false (rows 2 and 3). Similarly, exclusive disjunction entails inclusive disjunction but not the other way round. This is because in every situation that exclusive disjunction is true, inclusive disjunction is also true but there is one situation where inclusive disjunction is true but exclusive disjunction is not. Conjunction and exclusive disjunction stand in contradiction to each other. If the conjunction of two propositions is true then the exclusive disjunction is false and if the exclusive disjunction is true then their conjunction is false. Figure \@ref(fig:entailmentFigure) shows propositions as sets of situations in which they are true. The entailment relations can be seen as inclusion: a proposition that is contained within another proposition entails it. Exclusive disjunction and conjunction are both contained within inclusive disjunction and they both entail it.

```{r entailmentFigure, fig.align="center", fig.asp=0.5, fig.cap="The relation betwen conjunction, inclusive disjunction, exclusive disjunction, and their negations."}
connective_entailment <- png::readPNG("figs/connective_entailment.png")
grid::grid.raster(connective_entailment)
```

Negation in classical logic is a unary operator that flips the truth value of a proposition. Figure \@ref(fig:entailmentFigure) also shows the negation of conjunction and two types of disjunction. Negated inclusive disjunction is contained within negated conjunction and negated exclusive disjunction; consequently it entails them. Table \@ref(tab:negationtable) shows the meaning of negative atomic propositions ($\lnot \phi$, $\lnot \psi$), negative inclusive disjunctions ($\lnot (\phi \lor \psi)$), negative conjunctions ($\lnot (\phi \land \psi)$), inclusive disjunction of negative propositions ($\lnot \phi \lor \lnot \psi$), the conjunction of negative propositions ($\lnot \phi \land \lnot \psi$), negative exclusive disjunction, and exclusive disjunction of negative propositions. 

|$\phi$|$\psi$|$\lnot \phi$|$\lnot \psi$|$\lnot (\phi \lor \psi)$|$\lnot (\phi \land \psi)$|$\lnot \phi \lor \lnot \psi$|$\lnot \phi \land \lnot \psi$|$\lnot (\phi \oplus \psi)$|$\lnot \phi \oplus \lnot \psi$|
|:-:|:-:|:-:|:-:|:-----:|:-----:|:-----:|:-----:|:------:|:-----:|
|T|T|F|F|F|F|F|F|T|F|
|T|F|F|T|F|T|T|F|F|T|
|F|T|T|F|F|T|T|F|F|T|
|F|F|T|T|T|T|T|T|T|F|
Table: (\#tab:negationtable) Truth conditions for statements that involve the interaction of conjunction and disjunction with negation in classical logic.

Comparing the columns in Table \@ref(tab:negationtable) we can observe at least three equivalence relations. The first two are the ones named after the 19th-century British mathematician Augustus De Morgan[^2]. The negation of an inclusive disjunction is equivalent to the conjunction of their negatives and the negation of a conjunction is equivalent to the disjunction of their negatives.

[^2]: The so called De Morgan laws were known even to Stoics so the discovery of them certainly predate De Morgan's formulation.

De Morgan's Laws:

$\lnot (\phi \lor \psi) \Leftrightarrow \lnot \phi \land \lnot \psi$

$\lnot (\phi \land \psi) \Leftrightarrow \lnot \phi \lor \lnot \psi$

The third equivalence relation concerns exclusive disjunction. The exclusive disjunction of two propositions has similar truth conditions to the exclusive disjunction of their negatives. If exactly one disjunct is true and the other false, flipping the truth values using negation results in exactly one disjunct true and another false again. If the disjuncts have similar truth values, their negatives will also have similar truth values.


$\phi \oplus \psi \Leftrightarrow \lnot \phi \oplus \lnot \psi$


Table \@ref(tab:logicalproperties) shows some other logical properties of exclusive and inclusive disjunction. Truth preservation refers to the property that if both disjuncts are true, then their disjunction is also true. Truth preservation holds for inclusive disjunction but not exclusive disjunction. Falsehood preservation is the opposite: the property that if both disjuncts are false then the disjunction is also false. Both inclusive and exclusive disjunction are falsehood preserving. Idempotency for disjunction refers to the property that the disjunction of a proposition with itself is equivalent in truth conditions to the original proposition. Inclusive disjunction is idempotent but not exclusive disjunction. However, both inclusive and exclusive disjunction are commutative and associative; meaning neither the order of the operands nor the order of the operations alter the outcome of the disjunctions. While inclusive disjunction distributes over conjunction, exclusive disjunction does not.    

|Properties|Inclusive Disjunction $\lor$|Exclusive Disjunction $\oplus$|
|--------------------------|:--------------------------:|:----------------------:|
|Truth Preservation|$a=T,b=T \Rightarrow a \lor b = T$|X|
|Falsehood Preservation|$a=F, b=F \Rightarrow a \lor b = F$|$a=F, b=F \Rightarrow a \oplus b = F$|
|Idempotency|$a \lor a \Leftrightarrow a$|X|
|Commutativity|$a \lor b \Leftrightarrow b \lor a$|$a \oplus b \Leftrightarrow b \oplus a$|
|Associativity|$(a \lor b) \lor c \Leftrightarrow a \lor (b \lor c)$|$(a \oplus b) \oplus c \Leftrightarrow a \oplus (b \oplus c)$|
|Disributivity|$a \lor (b \land c) \Leftrightarrow (a \lor b) \land (a \lor c)$|X|
|Monotonicity|$(a \rightarrow b) \Rightarrow (a \lor c) \rightarrow (b \lor c)$|X|
Table: (\#tab:logicalproperties) Some logical properties of inclusive and exclusive disjunction.

```{r binaryLogicalConnectives, fig.env="figure", fig.align="center", fig.width=5, fig.height=2, fig.cap="The truth table for the 16 binary logical connectives. The rows represent the set of situations where zero, one, or both propositions are true. The columns represent the 16 possible connectives and their truth conditions. Green cells represent true situations."}
binary_connectives<- png::readPNG("figs/binary_connective.png")
grid::grid.raster(binary_connectives)
```
Logical systems often pick one or the other definition of disjunction and give it more primacy. For example, Stoics considered exclusive disjunction as primary. Inclusive disjunction was considered "pseudo-disjunction" or merely the negation of a conjunction. It seems that the main reason for considering exclusive disjunction as primary was that unlike inclusive disjunction, it cannot be concisely redefined using conjunction and negation. Therefore, Stoics thought that exclusive disjunction must have an independent and primary status as an operator while inclusive disjunction is just a derivative. Unlike their Stoic colleagues, medieval philosophers defined disjunction as inclusive. Leibniz dispensed with both definitions of disjunction but later Bolzano (1781-1840) used both. George Boole (1815 - 1864) gives more primacy to exclusive disjunction since it corresponds to addition modulo 2. Later Charles Sanders Peirce (1839-1914) changed Boole's system to use inclusive disjunction instead. Inclusive disjunction has enjoyed a primary role in modern logical systems since then.

The question of whether inclusive disjunction is "primary" in a logical system or exclusive disjunction depends on the goal of the system. The properties of exclusive disjunction make it ideal as an operator for some purposes but not others. When it comes to natural language semantics and modeling the meaning of the word *or*, we can similarly ask whether exclusive disjunction is primary (better represents *or*'s meaning) or inclusive disjunction. In the following sections we will discuss linguistic work that has addressed this issue but before I move to the discussion of disjunction in natural language, I would like to comment on two properties of exclusive disjunction that mismatch the properties of *or* in English.

First the lack of idempotency suggests that the exclusive disjunction of a proposition with itself ($\phi \oplus \phi$) is necessarily false. Applied to the word *or* in natural languages, the prediction is that a sentence such as "Bob is happy or Bob is happy" should be judged as false regardless of whether Bob is happy or not. This intuition seems to be absent in judgments of such sentences with *or*. In fact, there is an intuition that such disjunctions are redundant and have the same truth value as the atomic proposition itself. For example, "the sky is blue or the sky is blue" is redundant and equivalent to "the sky is blue". This intuition is predicted if *or*'s meaning is represented by inclusive disjunction but not exclusive disjunction.

Second, exclusive disjunction of two propositions has the same truth values as the exclusive disjunction of their negatives ($\phi \oplus \psi \Leftrightarrow \lnot\phi \oplus \lnot\psi$). Considering the natural language *or*, this property suggests that the sentence "the door is open or the window is open" should have the same truth conditions as "the door is not open or the window is not open". However, the first seems false intuitively when both are closed while the second is false when both are open. Therefore, two essential properties of exclusive disjunction seem to mismatch the meaning of *or* in English. Inclusive disjunction on the other hand does not face these problems. However, many examples of English *or* such as "He is a basketball player or a soccer player" seem to exclude the possibility of both disjuncts being true. Therefore, exclusive disjunction seems too strong for the meaning of *or* and inclusive disjunction often falls short of the interpretation *or* receives. This tension has lead researchers in semantics and pragmatics to consider inclusive disjunction as the primary meaning of *or* in English, and resort to other mechanisms that pragmatically strengthen the inclusive meaning of *or* to an exclusive interpretation. In the next few sections, I briefly review the history and the current issues in the semantics and pragmatics of disjunction.

## Disjunction in Language

In natural languages, disjunction and conjunction are types of coordination. @haspelmath2007 defines coordination (or coordinate constructions) as: "syntactic constructions in which two or more units of the same type are combined into a larger unit and still have the same semantic relations with other surrounding elements." The example below shows several coordinate constructions in English. 

(1) 
a. The dog barked *and* the cat ran behind the sofa.
b. I will get some coffee *or* tea.
c. He went to bed *because* he was sleepy.
d. She was sick *but* she finished the homework.

The word/affix that marks coordination is called a coordinator. The coordinators are shown in italics in the example above. The units that are marked by the coordinator are called coordinands. "The dog barked" is a coordinand in the first example. In English, a conjunction is a coordination marked by the coordinator *and*, and a disjunction is marked by *or*. The coordinands in conjunctions and disjunctions are called conjuncts and disjuncts respectively. 

While there has been no report of a language that does not have ways to express conjunction and disjunction, several languages lack overt coordinators that mark conjunction or disjunction. In fact, @haspelmath2007 reports that the mere juxtaposition of the conjuncts is a widespread strategy to convey conjunction crosslinguistically. For disjunction, often the juxtaposition is also accompanied by a marker of uncertainty or modality. For example, in Maricopa a sentence such as "John and Bill will come" is expressed as "John, Bill, will-come." while the disjunction "John or Bill will come" is expressed as "John, Bill may-come." (Gil 1991) Similarly in Dyirbal a disjunction such as "A or B" is expressed as "maybe A, maybe B" (Dixon 1972: 363). A similar effect of modals on the interpretation of coordination can also be found in English. Consider the English examples below.

(2) Here is what you will see on the table: a book, a pen, a calculator.

(3) Here is what you may see on the table: a book, a pen, a calculator.

A list of items with no overt coordinator such as "a book, a pen, a calculator" can be interpreted as a conjunction or a disjunction. In the first sentence, the list is interpreted as a conjunction: you will see a book, a pen, and a paper. In the second example, the list is interpreted similar to a disjunction: you may see a book, a pen, or a calculator. The sentences are identical except for the modal verbs. The sentence with a disjunctive interpretation uses the possibility modal *may*. @Ariel2014 reports a similar naturally occurring example with the modal *perhaps*, repeated below. Such examples point to a systematic connection between the notions of disjunction and modality.

(4) Practices of abortion, perhaps, pre-partum, perhaps postpartum (LSAC).

(5) Practices of abortion, pre-partum, or postpartum.

Languages that mark coordination overtly do it in several ways. The most common pattern, at least in European languages, is to place the coordinator in between the coordinands like "A and B" and "A or B" in English. However, other patterns such as "A B-and" where the coordinator appears after the coordination is also attested (see @haspelmath2007 for more details). Another notable and common cross-linguistic pattern is the doubling of the coordinator on each coordinand. For example in Persian (Farsi), a disjunction can be expressed using the disjunction word *ya* like "A ya B" or it can be doubled on each disjunct like "ya A ya B". Similar patterns have been attested in Polish ("albo A albo B"),  Dutch ("of A of B"), Basque ("ala A ala B"), Somali ("ama A ama B"), and French ("ou A ou B") among many others. In English the *either A or B* construction comes close to this pattern. In the next section, I focus on the interpretations of disjunction in English but the discussion is not limited to English and similar results have also been attested in other languages.

## Disjunction in English

The use of the word *or* in English is informally observed to correlate with one or more of the following implications [@Aloni2016]: 

1. Inclusivity (IOR): at least one of the disjuncts is true;
2. Exclusivity (XOR): exactly one disjunct is true;
3. Conjunctivity (AND): both disjuncts are true;
4. Ignorance: the speaker does not know which disjunct is true;
5. Indifference: which disjunct is true does not matter for the purpose of the conversation. 

These five implications do not have equal status with respect to each other. The first three (inclusivity, exclusivity, and conjunctivity) are binary truth-conditional meanings discussed in section \@ref(logic). The fourth, ignorance, is related to the knowledge state of the speaker, and the fifth, indifference,  is related to the role of disjunction within the conversation.

These five implications are also not necessarily caused by the word *or* itself. They may directly stem from the meaning of *or*, or they may be the result of *or* interacting with several other factors that shape the overall communicative message of the utterance. The goal of semantic and pragmatic research on *or* is to separate the implication(s) contributed by *or* from those caused by other words or factors in the communication of meaning. Disentangling the contribution of *or* from the contribution of other communicative elements has proven difficult, yet extremely fruitful in advancing semantic and pragmatic theories. In what follows I try to provide a short overview of the difficulties that *or* poses to theories of meaning.

| Example | IOR | XOR | AND | Ignorance | Indifference |
|--------------------------------------|:-:|:-:|:-:|:-------:|:-------:|
|Study today or tomorrow, to pass the exam! |$\checkmark$|X|X|$\checkmark$|$\checkmark$|
|Bob studied yesterday or the day before. |X|$\checkmark$|X|$\checkmark$|X|
|Students, like Bob or Becky, will pass. |X|X|$\checkmark$| X | X |
Table:(\#tab:implications) Implications of *or* for three example sentences.

Table \@ref(tab:implications) shows three example sentences with the disjunction word *or* and marks the presence or absence of each *or*-implication. A sentence such as "Study today or tomorrow, to pass the exam." implies that the addressee is asked to study today, tomorrow, or possibly both days. The overall message is that the addressee should study at least one of those two days to pass the exam. It does not rule out the possibility of the addressee studying both today and tomorrow, so there is no exclusivity implication. Furthermore, it is not asking the addressee to both "study today" **and** "study tomorrow", so a conjunctive implication is absent. It is also possible to infer that the speaker does not know or it does not matter to them whether the addressee studies today or tomorrow. The command is complied with as long as the addressee studies one of those two days. 

Compare this to a sentence like "Bob studied yesterday or the day before." The overall message is that Bob studied one day and that day may have been yesterday or the day before. It suggests that Bob did not study both yesterday and the day before, so there is an exclusivity implication and no conjunctivity implication. It is also implied that the speaker does not know which day exactly Bob studied so we have an ignorance implication, but it is not implied that the speaker is indifferent towards which day Bob actually studied. Now compare the previous sentences to a sentence such as "students like Bob or Becky will pass the exam." This sentence does not imply inclusivity or exclusivity. It does not communicate anything like "Either students like Bob will pass, or students like Becky will pass, or both." It does not exclude the possibility of students like both Bob and Becky passing at all. In fact to the contrary, the main message is that "both students like Becky **and** students like Bob will pass". In this example, *or* has a conjunctive implication. Even replacing *or* with *and* does not substantially alter the meaning of the sentence: "students like Bob and Becky will pass the exam". No ignorance or indifference implication accompanies this conjunctive implication.

Just looking at Table \@ref(tab:implications) one might conclude that there is no systematicity behind the use of *or* and the implications that accompany it. This was indeed the prevailing view among many logicians and philosophers that considered language messy, fraught with ambiguities, and illogical. For example, while discussing disjunction in his influential introductory book on logic, Alfred Tarski warns readers about the many interpretations of *or* in natural language and explains that there are "quite noticeable differences between the usage of it in everyday language and in logic "[@tarski1941logic]. He first points out that "the word *or* in everyday language has at least two different meanings". A child may ask us to be taken to a hike in the morning and a theater in the afternoon, but we may respond: "No, we are going on a hike or we are going to the theater". He explains that the interpretation of *or* in this example is exclusive because "we intend to comply with only one of the two requests" and not both. However, a disjunction may also have an inclusive interpretation like the following example: "Customers who are teachers or college students are entitled to a special reduction". Tarski explains that *or* in this example is inclusive "since it is not intended to refuse reduction to a teacher who is at the same time a college student." He advises readers to avoid this exclusive vs. inclusive ambiguity by reserving the word *or* for the "logical" (inclusive) sense and use the construction "either ... or ..." for the exclusive one. 

However, he immediately notes that there are other interpretations of *or* that go beyond the exclusive/inclusive distinction. I include his discussion here directly since in addition to addressing the ignorance implication of *or*, it foreshadows future developments in semantics and pragmatics initiated years later by Paul Grice.

>In commmon language, two sentences are joined by the word *or* only when they are in some way connected in form and content. (The same applies, though perhaps to a lesser degree, to the usage of the word *and*) ... anybody unfamiliar with the language of contemporary logic would presumably be little inclined to consider such a phrase as "2+2=5 or New York is a large city" as a meaningful expression, and even less so to accept it as a true sentence. Moreoever, the usage of the word *or* in everyday English is influenced by certain factors of a psychological character. Usually we affirm a disjunction of two sentences only if we believe that one of them is true but wonder which one. If, for example, we look upon a lawn in normal light, it will not enter our mind to say that the lawn is green or blue, since we are able to affirm something simpler, and at the same time, stronger, namely that the lawn is green. Sometimes even, we take the utterance of a disjunction as an admission by the speaker that he does not know which of the members of the disjunction is true. And if we later arrive at the conviction that he knew at the time that one -- and specifically, which -- of the members was false, we are inclined to look upon the whole disjunction as a false sentence, even should the other member be undoutedly true. Let us imagine, for instance, that a friend of ours, upon being asked when he is leaving town, answers that he is going to do so today, tomorrow, or the day after. Should we then later assertain that, at that time, he had already decided to leave the same day, we shall probably get the impression that we were delibrately misled and that he told us a lie. The creators of contemporary logic, when introducing the word *or* into their considerations, desired, perhaps unconsciously, to simplify its meaning and to render the latter clearer and independent of psychological factors. [@tarski1941logic]

Tarski ended his discussion by saying that these linguistic and psychological factors should not enter logical considerations. The view that language is not amenable to the tools of logic remained prevalent but faced substantial challenges from philosophers that argued for more attention to ordinary language use. Finally, the work of two philosophers, Richard Montague, a student of Tarski, and Paul Grice, tilted arguments towards the "ordinary language" philosophers. Montague's focus was on quantification in logic and natural language while Grice focused on the logical connectives such as *or* and *if*. @grice1989studies argued that the perceived differences between the meaning of natural language words such as *or* and the operators in formal logic such as inclusive disjunction "arise from inadequate attention to the nature and importance of the conditions governing conversation." He set out to show that when we take conversational factors into account, the meaning of linguistic connectives are similar to the common definitions of their counterparts in formal logic.

An important contribution of Grice was his typology of meaning that lead to the establishment of semantics and pragmatics. He differentiated between "what is said" and "what is implied". "What is said" refers to the literal meanings of words; the conventional association between words and meanings that is independent of any particular context. "What is said" is the most primary type of meaning and not derived from any other process. The field of semantics is concerned with this type of meaning and all else that relies on context of use is relegated to pragmatics. On the other hand, "what is implied" (conversationally) or as he called them "conversational implicatures" refer to meanings that are created by using words and sentences in specific contexts[^3].  Conversational implicatures are the result of refining the literal meanings of words (what was said) to suit the assumptions and purposes of the conversational context. As such, implicatures are derivative and not primary. 

[^3]: I set aside conventional implicatures here since they do not come up in the semantics and pragmatics of disjunction.  

Grice was also influential because he pioneered the methodology of studying linguistic meaning. He viewed the interpretation of an utterance as the composite of many interacting factors. In order to understand the literal meaning of a word, a semanticist needs to examine it in different contexts and understand its interaction with other factors that shape the general interpretation of the utterance. Utterances containing a word like *or* may be interpreted differently in different contexts, giving the impression that the word is polysemous. However, it is possible that the word has one underlying meaning (i.e. its semantics) and that in interaction with other elements of the sentence and the conversation, it gives rise to a variety of interpretations. In such cases, the underlying meaning can be recovered by detecting the factors that influence interpretation and reversing their effects. In Grice's view, true polysemy is stable polysemy across a wide range of contexts.

In addition to context-sensitivity, Grice suggested cancellability and computability as other properties of pragmatic enrichment. Cancellability refers to the following observation: a speaker explicitly denying a pragmatic enrichment of what they said does not result in the speaker contradicting themselves. For example, if I say that "the chocolate box is in the fridge or the cupboard", you may infer that I do not know where it is (ignorance implication). But, I can continue by saying "I know where it is but I'm not going to tell you" without contradicting what I said before. According to Grice, this suggests that the ignorance implication is not part of the literal meaning of what I said. The idea is that my continuation cancelled the pragmatic inference that I do not know where the chocolate box is. With literal meaning on the other hand, a speaker explicitly denying what they literally said results in a contradiction. For example, if I say "the chocolate box is in the fridge or the cupboard", I cannot add "it is in neither place" without sounding contradictory. Grice also emphasized on calculability of non-literal (non-conventional) meaning (i.e. conversational implicatures). He believed that if researchers propose a certain interpretation to be non-literal (more accurately non-conventional), they should be able to explain the mechanism that gave rise to that interpretation. Finally, it is important to note that Grice did not mean to introduce an exhaustive and definite set of diagnostics to distinguish literal and implicated meaning. He emphasized that in many cases, our intuitions will be the guide on what is said literally and what is merely implicated by what is said. 

## Factors Involved in the Interpretation of Disjunction

This section presents a brief list of factors that influence the interpretation of disjunction in English. This list is not meant to be exhaustive; it only provides a window into the factors that an account of disjunction acquisition needs to consider. The factors discussed in this section include: conversational principles, entailment environment, modality, semantics of the disjuncts, metalinguistic communication (definitions and repairs), syntactic categories of the disjuncts, question intonation, embedded imperatives, and alternative unconditional constructions. I start this section with Grice's discussion of conversational principles that contribute to the overall interpretation of an utterance.

### Conversational Principles

Grice's focus was on how conversational factors enrich utterance interpretation. He saw conversation as a cooperative social activity in which participants follow a certain set of rules or "maxims". Below I list the overarching "cooperative principle" as well as its maxims. Grice contended that these maxims further enrich the primary meaning of words and sentences in a given context. He used the term *conversational implicature* to refer to implications that are derived from the cooperative principle and the conversational maxims. 

* **Cooperative Principle**: Make your conversational contribution such as is required, at the stage at which it occurs, by the accepted purpose or direction of the talk exchange in which you are engaged.
    + **Maxim of Quality**: Try to make your contribution one that is true. Do not say what you believe to be false. Do not say that for which you lack adequate evidence.
    + **Maxim of Quantity**: Make your contribution as informative as required. Do not make your contribution more informative than is required.
    + **Maxim of Relation**: Be relevant.
    + **Maxim of Manner**: Be perspicuous. Avoid obscurity of expression. Avoid ambiguity. Be brief. Be orderly. 

Grice's account provided some informal explanations for the divergences between the interpretation of *or* and inclusive disjunction in logic. For example, Tarski pointed out that sentences that are connected with *or* should be related with respect to their content. He reported that English speakers find a sentence like "2+2=5 or New York is a large city" very odd and hard to judge as true, even though one of the disjuncts is certainly true. Grice's "maxim of relation" provides an independently motivated explanation for this observation. In conversations, speakers are expected to provide information relevant to the topic of the conversation. It is extremely hard to think of a context where "2+2=5" and "New York is a large city" are relevant alternatives to the conversational topic. Tarski himself noticed, however, that this phenomenon is not limited to the word *or*. Other connectives face the similar problem: "2+2=4 and New York is a large city" or "if 2+2=4, then New York is a large city". 

Tarski also pointed out that a disjunction such as "A or B" is often odd when the speaker knows which alternative is true. In other words, a disjunction often implies that the speaker is ignorant with respect to the truth of the disjuncts. The definition of disjunction in classical logic lacks this aspect. Grice's theory provides a separate mechanism that is responsible for the ignorance implication. The maxim of quality requires the speaker to say what they believe to be true and the maxim of quantity requires them to be as informative as required. In many contexts the speaker is expected to be maximally informative. If the speaker knows A then they should say "A". In Tarski's example, if the speaker knows that the grass is green, then the most informative statement would be "the grass is green." Saying that "the grass is green or blue" goes against the maxim of quantity. The speaker is providing less information even though it is clear they know more. 

However, as Grice points out, in some contexts there is a reason for being underinformative and in such cases there is no  ignorance implication for the speaker. For example, he mentions that in a treasure hunt with his children, he may say: "The prize is in the garden or the attic. I know that because I know where I put it, but I'm not going to tell you." In this context the speaker knows which alternative is true. However, the context does not require him to be maximally informative. In fact, given the rules of a treasure hunt, he is required to not divulge the whereabouts of the prize. Therefore, the speaker's underinfomative utterance is warranted. This way Grice shows that the (speaker) ignorance implication is not part of what *or* means but rather the result of *or* interacting with conversational maxims. It is important to point out that a disjunction such as "A *or* B" is unacceptable when both the speaker and the addressee know which disjunct is true. In Grice's example, while Grice knows which disjunct is true (there is no speaker ignorance), the addressee does not (there is addressee ignorance). However, in Tarski's example, both the speaker and the addressee know whether the grass is green or blue. Therefore, the assertion of a disjunction seems to require a common ground that is not settled with respect to A and B when "A or B" is uttered. Experimental studies that test disjunction comprehension need to control for this requirement by making either the speaker or addressee ignorant of which disjunct actually holds.

Tarski discussed the exclusive and inclusive interpretations of *or* as well. He explained that even though in logic a disjunction is commonly defined as inclusive, a sentence like "we are going on a hike or we are going to the theater" has an exclusive interpretation. Grice's account provides an explanation for this supposed "ambiguity" as well. In some contexts, the speaker knows whether both alternatives are true or not (Competence Assumption). If the speaker knows both alternatives to be true and the context requires maximal informativeness, then they should use the connective *and*. If they know that only one alternative is true, then they cannot use *and* because using it would violate quality (the speaker has said something false.) In such contexts using *or* instead of *and* implies that only one alternative is true and not both. The exclusive example provided by Tarski is similar to such contexts. The father knows (and in fact decides) whether both alternatives will be true or only one. If he intended for both to be true he could use the word *and*: "we are going on a hike and we are going to the theater". Since he did not, we can infer that he does not intend to do both. In fact in Tarski's original context, the father says the sentence with *or* in contrast to the child's utterance with *and*.

The Gricean account has been developed and made more explicit by several authors. Here I present the current standard neo-Gricean account which is mostly due to @sauerland2004scalar. Assuming the speaker has uttered a disjunctive assertion such as "P or Q", the reasoning proceeds as follows:

* **Utterance**: the speaker said "P or Q".
* **Alternatives**: the speaker could have said: P, Q, or "P and Q". Why didn't s/he?
    1. **Ignorance**: The speaker is uncertain about the truth of P (i.e. $\lnot B_S (P)$)
    2. **Ignorance**: The speaker is uncertain about the truth of Q (i.e. $\lnot B_S (Q)$)
    3. **Ignorance**: The speaker is uncertain about the truth of P and Q (i.e. $\lnot B_S (P \land Q)$)
* **Competence**: The speaker knows whether both propositions hold or not (i.e. $B_S (P \land Q) \lor B_S \lnot (P \land Q)$)
    4. **Exclusivity**: Given ignorance inference 3 and speaker competence, the speaker believes that only one of the disjuncts is true (i.e. $B_S \lnot (P \land Q)$).
    
Inferences 1-3 are ignorance implications. It is important to notice that the exclusivity inference in step 4 develops from the ignorance inference in step 3 and the competence assumption. For the exclusivity implicature to arise, the context of the utterance must be one where the speaker should know whether both disjuncts hold or not, but s/he is not sure about which one holds. As @geurts2006exclusive argues, it is not clear how common such contexts are and what proportion of exclusivity implications are in fact implicatures derived via this type of pragmatic reasoning. Nevertheless, it is not impossible to think of such contexts. For example, if I say that "Bob went to the shop and bought a cardigan or a shirt", it is reasonable for the addressee to assume that I know whether one item was bought or two. At issue is not whether exclusivity implicatures exist at all but rather how common they are. There seems to be a tacit assumption in the literature on semantics and pragmatics that the majority of exclusivity implications are derived via the Gricean reasoning sketched above.

While ignorance, indifference, and exclusivity implications of *or* can be explained (at least partly) via the Gricean mechanism above, the conjunctive interpretation remains unexplained. @geurts2010quantity revises the Gricean account to cover the conjunctive inferences of free-choice expressions like "you can have coffee or tea". @potts2015negotiating extend the Gricean account to cover definitional uses like "he is a wine lover or an oenophile". I will discuss free-choice inferences and definitional uses below. As far as I know, there are no Gricean accounts for conjunctive interpretations in examples such as "Students, like Bob or Becky, will pass." There are also no Gricean accounts for examples of self-repair with disjunction such as "Bob changed the font, or I mean, the size of the text." These cases will be discussed below as well.

Before moving on, I would like to add that in Gricean pragmatics, the notion of "alternatives" or "what the speaker could have said" plays a crucial role. In the case of the ignorance implicature, the explanation relies on the assumption that instead of "A or B" the speaker could have said something simpler but more informative, namely "A". The Gricean explanation relies on the addressee reasoning about why the speaker did not just say "A". In case of the exclusivity implication, the Gricean approach relies on the assumption that instead of "A or B" the speaker could have used a different connective, namely *and* to say "A and B". The addressee reasoning about why the speaker did not do so derives the exclusivity implicature. Alternatives also play an important role in developmental accounts of pragmatics. @barner2011accessing argue that children are not adult-like in their access to the linguistic alternatives and as a result, they show non-adult-like interpretations. 

An active area of research in Gricean pragmatics is the question of what constitutes an alternative and how implicatures are computed given a set of alternatives to the asserted utterance. The alternatives to a lexical item were traditionally considered as a scale ordered by entailment. @horn1972semantic proposed that words such as *or* and *and* as well as *some* and *all* form scales of alternatives represented as <*or*, *and*> and <*some*, *many*, *most*, *all*>.  This is why the exclusivity implicature of *or* is also called a "scalar implicature". The basic idea is that asserting an utterance with the weaker lexical item on the scale (e.g. *or*) results on the inference that the corresponding utterance with the stronger item (e.g. *and*) would have been false (violated the maxim of quality). However, using the scale <*or*, *and*> will only derive the exclusivity implicature and does not provide an account of ignorance implications. In the standard account provided above, the ignorance implications rely on including each disjunct in the alternative set as well. Due to this problem (among others), the notion of "scale" has been replaced by "the set of alternatives". Nevertheless the term "scalar implicature" is still used to refer to the implicatures of *some*, and *or*. A large body of literature suggests that children do not compute scalar implicatures at the rate that adults do [See @katsos2014scalar for a review]. I discuss this issue more in the next two chapters.

### Entailment Environment

The concept of entailment plays an important role in logic and semantics. The concept is often introduced as a relation among propositions. $P$ entails $Q$ if and only if $P$'s truth makes $Q$ necessarily true. Based on this, two other notions are defined: upward entailing and downward entailing environments. 

* A linguistic environment $\phi$ is upward entailing if and only if for any two expressions $a$ and $b$, if $a$ entails $b$, then $\phi[a]$ entails $\phi[b]$.

* A linguistic environment $\psi$ is downward entailing if and only if for any two expressions $a$ and $b$, if $a$ entails $b$, then $\psi[b]$ entails $\psi[a]$.

Let's consider $a$ as "student", $b$ as "human", $\phi$ as "Bob is a []", and $\psi$ as "Bob isn't a []". Since "student" entails "human", and "Bob is a student" entails "Bob is a human", we can conclude that "Bob is a []" is an upward entailing environment. On the other hand, "student" entails "human" but "Bob isn't a human" entails "Bob isn't a student". Therefore, we can conclude that "Bob isn't a []" is a downward entailing environment. In short, upward entailing environments preserve entailment direction while downward entailing environments reverse it. 

As explained earlier, exclusive disjunction entails inclusive disjunction. In an upward entailing environment, this entailment relation is preserved: "Bob is a student or a teacher (but not both)" entails "Bob is a student or a teacher (or both)." However, in a downward entailing environment this relation is reversed: "Bob isn't a student or a teacher (neither)" entails "Bob isn't a student or a teacher (but may be both or neither)" A common observation is that the exclusive interpretation of *or* is more common in upward entailing environments while the inclusive interpretation is more common in downward entailing environments. For example a disjunction like "Bob is a student or a teacher" is more likely to be interpreted as "exclusive" while its negative counterpart "Bob is not a student or a teacher" is more often interpreted as "neither", which is expected if *or* is inclusive. This is predicted by Grice's maxim of quantity. In an upward entailing environment, exclusive disjunction is stronger and conveys more information. In a downward entailing environment, this pattern is reversed; inclusive disjunction results in a stronger and more informative statement. If a listener expects the speaker to make their contribution as informative as possible, then we can expect a bias towards exclusive disjunction in upward entailing environments and inclusive disjunction in downward entailing ones. 

As explained in the section on logic and disjunction, the interaction of negation with disjunction, and conjunction is captured by the so-called De Morgan laws. It is sometimes argued that the interaction of negative particles with disjunction and conjunction words follow these rules in natural languages as well. For disjunction the argument is that "not A or B" is interpreted as the conjunction of negatives "not A and not B". While this may be true in many cases, there are notable and systematic exceptions. This point is perhaps best illustrated by the negative phrase "not one or the other" in the following naturally occurring examples (6-10). In (6) below, "not one or the other" is used to mean "both". In example (7) it is used to mean "neither". In the (8) and (9) examples, the authors explicitly explain that by "not one or the other" they mean "neither or both". Finally (10) shows that these observations are not isolated to the expression "one or the other".

(6) Speed or Quality? It's Not One or the Other. It's Both! (Online Blogpost title by Chris Manuel)

(7) ... there is no in-between: you're either masculine or feminine because you're either male or female, and if you're not one or the other of these two genders, then there must be something wrong with you. In numerous other cultures, however, there are gender systems that are not binary.

(8) In the above example [containing this expression: $time =~ /(\d+):(\d+)(:(\d+))?/], the third set of parentheses is used to associate both the colon and the digits with "?" - either both should be specified, or neither, but not one or the other. (Response to a programming exercise on an online forum)

(9) Everything is either physical or spiritual ... This is of course the root of the problem ... If things are either A or B, it actually makes sense to assume that once they're not one, then they are the other. But what if they're not A or B? what if, for instance, they're neither? Or ... both? (Medium Article by Doc Ayomide: I'm Christian and I don't believe mental illness is spiritual)

(10) The church of the Lord is not here or there, but everywhere. (online newsletter at newchurch.org)

The examples above suggest that in natural language, the negation of a disjunction is not necessarily the conjunction of negations. These examples are not presented here to suggest that the NOR and AND interpretations of not (A or B) are equally likely. The examples show the importance of linguistic and non-linguistic factors in shifting our biases one way or another. Entailment environment is simply one of these factors.

Nevertheless, research in semantics has shown that the entailment environment is an important factor in the interpretation and distribution of lexical items (see @giannakidou2011negative for a discussion). Based on this, @crain2012emergence has suggested that the entailment environment and its interactions with logical words such as *or* must be an innate property of the human mind. I will cover this nativist model of logical word acquisition in Chapter \@ref(devoLit). In the next section I discuss the role of modality in the interpretation of disjunction.

### Modals

The interaction of *or* with modals, especially possibility modals such as *may* and *can* presented the biggest challenge for the proposal that *or* has the semantics of inclusive disjunction. The issue was first discussed by @kamp1973free in the context of giving permissions. Consider a sentence like: "You may pay using cash or credit card." It has two main interpretations. One is the standard inclusive interpretation which is more accessible if it is followed by an expression of uncertainty: "you may use cash or you may use credit card, or possibly both but I'm not sure". This interpretation is predicted by the standard inclusive account. However, there is a second and often more prominent interpretation for this sentence which suggests customers are free to choose: they are allowed to use cash *and* they are allowed to use credit card. This so called "free choice" interpretation is not expected if *or* means inclusive disjunction and *may* acts like a standard possibility modal. "It is possible that A or it is possible that B" is weaker and less informative than "it is possible that A and it is possible that B". However, it seems that in natural language when *or* appears with possibility modals it can have both interpretations. 
There have been many proposals to tackle this problem. Here I discuss two main proposals and expand on one that has made its way into acquisition research. In one approach, semanticists abandoned the truth-conditional inclusive semantics of *or* and proposed that *or* encodes the conjunction of two possibilities [@zimmermann2000free; @geurts2005entertaining]. In a second approach, they kept the standard semantics of *or* and treated the free choice inference as an implicature [@fox2007free; @geurts2010quantity].

First, @zimmermann2000free made a departure from the truth-functional account of *or* and proposed that *or*'s meaning is inherently modal. "A or B" expresses the conjunction of two possibilities and can be paraphrased as "It is possible that A and it is possible that B," represented in modal logic as $\Diamond A \land \Diamond B$. For example, "Bob paid using cash or credit card" is equivalent to "It is possible that Bob paid using cash and it is possible that Bob paid using credit card." The modal account of *or* predicts a different logical structure for the free choice sentences. A sentence like "Bob is allowed to pay using cash or credit card" will have two layers of possibility modals: $\Diamond\Diamond A \land \Diamond\Diamond B$, roughly paraphrased as "it is possible that Bob is allowed to pay using cash, and it is possible that Bob is allowed to pay using credit card." This is the interpretation without a free-choice implication. The free-choice interpretation is only derived when the speaker is assumed to be an "authority" on the subject. According to the "authority principle", if an authority considers it possible that Bob is allowed to pay in cash, then Bob is allowed to pay in cash ($\Diamond \Diamond A \rightarrow \Diamond A$). Assuming speaker authority, a statement like "it is possible that Bob is allowed to pay using cash and it is possible that Bob is allowed to pay using credit card" reduces to "Bob is allowed to pay using cash and Bob is allowed to pay using credit card" ($\Diamond\Diamond A \land \Diamond\Diamond B \rightarrow \Diamond A \land \Diamond B$). Therefore, the modal account of disjunction provides a straightforward answer to the puzzle of free-choice interpretations, yet it dispenses with the truth-functional account of disjunction.

The pragmatic approach strives to keep the truth-functional meaning of *or* and analyze the free-choice interpretation as an implicature. It takes two observations as its starting point. First, that the free-choice inference is cancellable. As explained before, "you are allowed to drink tea or coffee" can be followed with "but I'm not sure which", therefore not implying that the addressee has free choice between tea and coffee. Second, when we add negation to free-choice sentences, the interpretation is not the negative of the conjunctive free-choice interpretation, but rather the negative of the standard account in which disjunction is inclusive. For example, the main interpretation of a sentence like "you are not allowed to drink tea or coffee" is one where neither option is allowed. This interpretation is expected if negation is operating on the possibility modal and an inclusive disjunction ($\lnot \Diamond (A \lor B)$). Proponents of the pragmatic approach argue that this behavior is not limited to negation and applies to downward entailing environments more generally. Consider a sentence like: if I'm allowed to drink tea or coffee, I'll be happy. We can infer that the speaker will be happy even if they are only allowed to drink tea. However, under a conjunctive interpretation of *or* with *allowed to*, one should expect that the speaker will only be happy if both are allowed. This is not the interpretation we intuit. The pragmatic approach keeps the semantics of *or* as inclusive disjunction and contends that similar mechanisms should explain exclusivity as well as the conjunctive (free-choice) implication [@fox2007free;@chierchia2013logic; @geurts2010quantity]. In what follows, I sketch @fox2007free's analysis of free-choice inferences of *or*. This analysis was later adopted by @Singh2016, who reported that children sometimes interpret disjunction like conjunction in simple declaratives. I talk more about this phenomenon in \@ref(GriceanEmpirical).

Applying the ordinary pragmatic mechanism explained earlier to sentences with modals such as *may* and *allowed to* results in the wrong prediction. The problem lies in the alternatives considered for pragmatic computation. Consider the sentence "you are allowed to drink coffee or tea". The standard alternatives to this sentence are: you are allowed to drink coffee, you are allowed to drink tea, and you are allowed to drink coffee and tea. If we mechanistically apply the Gricean recipe, we derive nothing more than the standard interpretation: "you are allowed to drink coffee or tea; but the speaker is not certain that you are allowed to drink coffee, and the speaker is not certain that you are allowed to drink tea, and the speaker is not certain that you are allowed to drink both coffee and tea". This is not the free-choice inference we have been looking for.

@fox2007free resolves this issue by changing the alternatives. As explained before, the standard set of alternatives to a disjunction like "A or B" is A, B, and "A and B". For the sentence "you are allowed to drink coffee or tea", the standard alternatives are "you are allowed to drink coffee" ($\Diamond A$), "you are allowed to drink tea" ($\Diamond B$), and "you are allowed to drink coffee and tea" ($\Diamond A \land \Diamond B$). @fox2007free proposes that the alternatives are instead the following set: "you are only allowed to drink coffee" ($\Diamond A \land \lnot \Diamond B$), "you are only allowed to drink tea" ($\Diamond B \land \lnot \Diamond A$"), and "you are allowed to drink coffee and tea" ($\Diamond A \land \Diamond B$). What justifies the change of alternatives from "allowed to X" to "only allowed to X"?  In @fox2007free's system this follows from the syntactic structure of these sentences. They contain a silent operator that has similar semantic effects as the word *only*. The process of applying this operator to a linguistic expression is called "exhaustification", and the operator is commonly abbreviated as EXH. In short, EXH($\phi$) asserts that $\phi$ is true and every alternative not entailed by $\phi$ is false. In @fox2007free's account, EXH applies to each disjunct, as well as the disjunction as a whole. The final product is the following implication: "you are allowed to drink coffee or tea; not **only** coffee, not **only** tea, and not both coffee and tea."

<!-- probably needs more explanation -->

@geurts2010quantity uses essentially the same solution as that of @fox2007free, but casts it in a pragmatic framework rather than a syntactic one. He suggests that pragmatic reasoning is intention-based and even though linguistic alternatives are important, it is the set of possible communicative intentions that play the key role in pragmatic computation. When a sentence such as "you are allowed to drink coffee or tea" is uttered, the listener considers four possible communicative intentions: (1) coffee and tea are both allowed, (2) coffee is allowed but tea isn't,(3) tea is allowed but coffee isn't, and (4) neither is allowed. He explains that intention (4) is ruled out because it is in contradiction with the basic meaning of the utterance. Options (2) and (3) are ruled out because if the speaker meant to convey them, they could have said something simpler, namely "you are allowed to drink tea" or "you are allowed to drink coffee". The only intention left that satisfies Gricean maxims is intention (1) which is the desired free-choice inference.

While the literature on the interaction of disjunction with modals has mostly focused on possibility modals such as *may* and the context of giving permissions, similar conjunctive inferences are present with preference/desire modals such as *want*, *like*, and *love*. If we say "Bob likes/loves going on hikes or climbing rocks or staying outdoors", it is clear that Bob likes all the activities listed. Therefore, it is possible to infer that "Bob likes/loves rock climbing." Similarly, we can infer from "Bob would like some coffee or tea" that "Bob **would like** some coffee" and "Bob **would like** some tea" (The inference is easier to access when *would like* is stressed). Similar inferences seem to be valid for "Bob wants some coffee or tea." However, it seems that the conjunctive inferences in such cases are not as strong as the case of permissions like "Bob is allowed to drink coffee or tea." Perhaps the strongest case of conjunctive inferences comes in the context of providing examples with the word "like". In a sentence such as "Students like Bob or Becky never fail", the disjunction word is almost equivalent to a conjunction: "Students like Bob and Becky never fail." These observations suggest that *or* is interpreted as conjunctive in a wider range of environments than commonly discussed in the literature.

### Disjunct Semantics

@geurts2006exclusive points out that even though pragmatic reasoning may be behind some exclusive interpretation of *or*, in many examples deriving the exclusivity interpretation via implicatures may be unnecessary because exclusivity is introduced by the semantics of the disjuncts themselves. Consider an example such as: "Bob is in the kitchen or the bathroom."  In the ordinary world that we live in, both disjuncts in this sentence cannot be true. Bob cannot be in the kitchen and the bathroom at the same time. Given that the inconsistency of the disjuncts is common knowledge to discourse participants, no inclusive interpretation is possible. The only available interpretation of such disjunctions is exclusive but this has little to do with the meaning of *or*. It stems from the semantic relation between the disjuncts. In fact, the exclusive interpretation would be present even when *or* is absent. Suppose someone asks where Bob is and the speaker responds with "Not sure ... in the kitchen ... in the bathroom." The interpretation of this response is similar to exclusive disjunction. 

More generally, our world knowledge provides us with likely relations between different propositions. For example, "Bob fell down" and "Bob hurt himself" are likely to co-occur and are interpreted as causally linked even though they do not have to be. The rich conceptual structure among different propositions can help the interpretation of linguistic connectives. Consider the following naturally occurring example from @Ariel2014: "You come ... you don' come ... it doesn't matter to me". In this example, the speaker uses no connectives between the main three sentences of the utterance, yet the interpretation of the relations among them is clear. The utterance can be paraphrased as "Whether you come or you don't come, it does not matter to me." The rich conceptual structure among coordinands often make it transparent what type of coordinator is required and this can subsequently help children's acquisition of connectives. With respect to disjunction, the disjuncts are often inconsistent in their meanings; only one can be true and not both. It is noteworthy that stoics described disjunction as "an operator for incompatibles", which points to the influence of disjunct semantics on the definition of disjunction in stoic logic.

### Metalinguistic Communication 

In two instances, disjunction is used to communicate about language itself. The first is when a speaker wants to communicate that two expressions have the same meaning (at least for the purposes of the conversation), and the second when a speaker wants to provide a repair; a signal that a linguistic error was made. I discuss these two cases below.

#### Definitions

Definitional or metalinguistic disjunction is the type of disjunction that I just used at the beginning of this sentence! The primary function of such a disjunction is to communicate that two expressions are equivalent in meaning or function, at least for the current purposes of the conversation. For example, a sentence such as "Bob is a wine lover or an oenophile" communicates that "oenophile" is another term for "wine lover" [@potts2015negotiating]. Similar to what we saw with modals, definitional uses give rise to conjunctive interpretations of disjunction. If "Bob is a wine lover or an oenophile", then "Bob is a wine lover" and "Bob is an oenophile". @potts2015negotiating propose that the following sociopragmatic conditions should hold for the definitional interpretations of *or*: First, discourse participants have mutual interest in communicating about the language itself, in addition to interest in communicating about the world. Second, the participants can assume that the speaker has expertise in the relevant domain. Third, that the cost of using a disjunction must justify its verbosity given that a disjunction of A or B is always longer than B itself to communicate the same meaning. 

@potts2015negotiating provide a Gricean account for definitional disjunctions in which "A or B" has the semantics of inclusive disjunction. The key innovation of the account is that conversational participants use language to convey information about the world as well as about language itself. @potts2015negotiating argue that a disjunction can be used to communicate information about a speaker's preferred lexicon, as well as the state of the world. Therefore, when the right sociopragmatic conditions for the definitional use hold, a disjunction such as "A or B" communicates two pieces of meaning: 1. "A" is true and 2. "A and B have the same meaning". The equivalence of A and B in meaning derives the conjunctive interpretation in this account. Given that the speaker has communicated that A and B have the same meaning, they are either both true or both false. Since the speaker also asserted that A is true, then B must be true as well. Therefore, in @potts2015negotiating's account the conjunctive interpretation of definitional uses is primarily the result of disjunct semantics and not *or* itself. In turn, disjunct semantics in definitional uses are the result of sociopragmatic conditions governing the context of the utterance.

It is important to point out that child directed speech often satisfies all three sociopragmatic conditions of definitional interpretations proposed by @potts2015negotiating. First, parents and children have mutual interest in communicating about the language itself given that children are active language learners. Second, in almost all areas, parents are experts with respect to the lexicon compared to children. Finally, it is reasonable to assume that the pedagogical goal of teaching a child the lexicon of a language justifies the verbosity of using a disjunction. Therefore, we may expect that definitional uses will show up commonly in child-directed speech and in fact, this is what we found in our corpus study presented in Chapter \@ref(corpus).

#### Repairs

A fairly unexplored area is the role of disjunction in conversational repairs. Often during casual speech, conversational participants notice a mistake either in their own speech or someone else's. The utterance that signals this mistake and provides correction is called a "repair". Repairs are often classified into "self-repair" and "other-repair". Self-repairs are repairs that are provided by the speaker themselves while other-repairs are provided by discourse participants other than the speaker. For example while discussing news on the flat earth society, a speaker may say "I can't believe there are people who still believe the earth is round" providing an immediate self repair "... I mean flat." Alternatively, someone else in the conversation may provide the repair with "... you mean flat." Repairs have the following three components: reparandum (the part of the original utterance that needs repair like "round"), editing term (a discourse marker like "I mean" that signals a repair), and alteration (the corrected section like "flat") [@heeman1999speech]. Common editing terms include *oh*, *um*, *uh*, *I mean/I meant*, *well*, *sorry*, *no*, *or*, and *let's see*. Not all repairs are accompanied by editing terms, especially those that repeat a large part of the reparandum in the alteration and make a minor modification such as "The boy was ha ... The man was happy."

The disjunction word *or* can be used as an editing term alone or along with other editing terms such as "I mean". The sentences below show a few examples of *or* as an editing term.

(10) "Engine two from Elmi(...) or engine three from Elmira." [example 14 of @heeman1999speech]
(11) "Why can't I change font or I mean size of the typed letters?" (online example)
(12) "I promised to see Nadelka again," said he, "or, I mean she promised to see me." (Stash of the Marsh Country by Harold Waldo)
(13) "John picked us up in his car, or rather his dad's car which he'd borrowed." (online example)
(14) "I met him very late on Friday night, or rather, early on Saturday morning." (online example)

It is reasonable to use a coordinator for repairs given that the reparandum-term-alteration structure is similar to the structure of coordination.  However, the communicated meaning in a repair is not what *or* often communicates. A repair commonly signals that the reparandum was not true or accurate and that the alteration is true and what the speaker is trying to communicate. A disjunction, on the other hand, commonly allows either disjunct to be true and does not rule out any disjunct. How come *or* is used as the connective for repairs?

It is important to note that while *or* does not carry the repair interpretation itself, its meaning -- either as inclusive disjunction or exclusive -- is not incompatible with a repair. The repair interpretation (the first disjunct is false, the second is true) is stronger and more specific but still compatible with the meaning of *or*. This is not the case for *and* since conjunction communicates that both conjuncts are true and this is not ideal for repairs. For example, in the repair sentences provided above, using *and* instead of *or* is infelicitous for communicating a repair. Since *or* is compatible with a repair reading but not strong enough, it is quite possible that similar to previous cases, the weak semantics of *or* (inclusive disjunction) is strengthened by external factors. 

For example, it is possible that *or* simply contributes the meaning that at least one of the disjuncts are true. Then factors that commonly signal repairs such as pauses, intonation, significant overlap between reparandum and alteration, or co-occurring edit terms such as "I mean" can strengthen the inclusive meaning of *or* to communicate that "in fact it is the second disjunct that is true". This account is supported by the fact that *or* is often optional in repairs. In the examples recounted before, if *or* is dropped, other elements such as the pause, the intonation, the phrase "I mean", or the word "rather" can still signal the repair in the utterance. Therefore, it is possible that *or* contributes inclusivity but these repair factors strengthen the coordination to mean that the first disjunct is false while the second is true.  

### Syntactic Units

To my knowledge, there has been no systematic investigation of the effect of syntactic category on disjunction interpretation. However, there is some informal evidence suggesting that it may play a role in generating exclusivity inferences. Compare the following example sentence: 

(16) He likes coffee or tea.
(17) He likes coffee or he likes tea. 

A common intuition is that the second disjunction with sentential disjuncts is more likely to be exclusive than the first disjunction with nominal disjuncts. The clausal vs. sub-clausal distinction also plays a role crosslinguistically. @haspelmath2007 reports that Yapese (an Austronesian language of Micronesia) uses different words for sentential and nominal conjunction. He reports different conjunction words for nominal and event conjunction as a widespread typological phenomenon, especially in African languages. He also reports that in Koromfe (a Gur language of Burkina Faso) a disjunction is only allowed for events so a sentence like "do you want coffee or tea?" must be rephrased as "do you want coffee or do you want tea?" These observations suggest that the syntax of a disjunction may play a role in shaping its interpretation. However, experimental investigations are required before we can confirm the effect of the disjuncts' syntactic categories on the interpretation of the disjunction.

### Question Intonation

|Question| Intonation| Example | Answer | Interpretation |
|---------|--------|--------------------------------|-------|:---------:|
|Polar|Rising|Would you like any tea($\uparrow$) or coffee$\uparrow$? | yes/no| $\lor$|
|Alternative|Rise-Fall|Would you like tea$\uparrow$, or coffee$\downarrow$? | tea/coffee|$\oplus$|
Table:(\#tab:questions) The interpretation of disjunction in polar and alternative questions.

There are two types of questions with the disjunction word *or*: polar questions and alternative questions. These two types of questions differ in the type of intonation and the responses they receive. Table \@ref(tab:questions) provides a summary of the properties of polar and alternative questions. Disjunctions in polar questions are accompanied by an overall rising intonation, or by rising intonation on each disjunct. Disjunctions in alternative questions receive rising intonation on the non-final disjuncts and falling intonation on the last. Polar questions typically receive a yes/no answer followed by one of the alternatives if the alternative matters for the purpose of the conversation. For example, if a waiter approaches and asks "would you like any tea or coffee?" the appropriate answer is typically "yes, tea/coffee please" or simply "no thank you". On the other hand, in the context of asking someone out on a date, a simple "yes/no" response to the same question may suffice given that the choice of alternative does not matter for the purpose of the conversation. When the choice of alternative is extremely relevant, the yes/no part of the response may be left unsaid and simply implied by the mention of the alternative. For example, in response to the waiter's question "would you like any coffee or tea?", the addressee can say "coffee, please". When it comes to alternative questions, a yes/no response is infelicitous. The purpose of an alternative question is to find out which alternative is true and a yes/no response does not do that. The one exception is when the alternatives themselves are positive and negative. For example, the alternative question "would you like coffee or not?" can be felicitously responded with "yes/no" but in such cases the alternatives themselves are "yes" and "no".

Polar questions receive an inclusive interpretation (at least one disjunct is true) while alternative questions receive an exclusive interpretation (exactly one disjunct is true). Intonation plays a crucial role in the interpretation of polar and alternative questions with disjunction. @pruitt2013interpretation recorded 24 disjunctive questions with both final rise intonation and final fall intonation. They asked 37 undergraduate participants to choose between two paraphrases: an inclusive paraphrase and an exclusive paraphrase. For example, a question like "did Sally bring wine or bake a dessert?" had the inclusive paraphrase "did Sally do any of these things: bring wine or bake a dessert?" and the exclusive paraphrase "which of these things did Sally do: bring wine or bake dessert?" They showed that the majority of responses (80%) considered a question with falling final intonation as exclusive and a question with rising final intonation as inclusive. 

<!--
Some languages use different connectives for polar vs. alternative questions. For example Mandarin Chinese uses the connective *haishi* for alternative and *huozhe* for polar questions.-->

### Embedded Imperatives

We can use *and* and *or* to connect imperative and declarative sentences. The two sentences below connect the same imperative and declarative sentences but use different connectives; the first uses *and* and the second *or*. An important observation with respect to the meaning of such utterances is that they can be paraphrased as conditionals. The first sentence with *and* can be paraphrased as "if you go home, you'll miss the fun." The second sentence with *or* can be paraphrased as "if you do not go home, you'll miss the fun." More accurately, the original sentences and their conditional paraphrases are biconditionals or perfected conditionals [@geis1971invited]. The sentence with *and* implies that if the addressee does not go home they will not miss the fun. The sentence with *or* implies that if the addressee goes home they will not miss the fun.

(18) Go home *and* you'll miss the fun. (If you go home, you'll miss the fun.)
(19) Go home *or* you'll miss the fun. (If you do not go home, you'll miss the fun.)

If we consider the relevant propositions as "addressee going home" and "addressee missing the fun", the first sentence has a conjunction interpretation with respect to these propositions while the second carries an exclusive interpretation. The imperative-or-declarative structure suggests that the proposition inside the imperative and the one inside the declarative will not be true at the same time: "If you go home, you won't miss the fun and if you don't go home you will miss the fun." 

<!--
@kaufmann2011interpreting analyses such imperative-connective-declarative constructions under the title "embedded imperatives". 
-->

### Alternative Unconditionals

Consider the following example of the alternative unconditional construction: "Whether you stay here or go home, you will have fun". The alternative unconditional construction has the following general schema: "Whether X or Y, Z." Sometimes the alternatives are negatives of each other: "Whether you stay here or not, you will have fun." When the alternatives are negatives of each other ("Whether X or not X, Z."), the shorter form "Whether or not X, Z" can also be used: "Whether or not you stay here, you will have fun." @rawlins2013conditionals classifies alternative unconditionals as a subtype of unconditionals along with constituent unconditionals (e.g. "Whatever happens, we win.") and headed unconditionals (e.g. "No matter what happens, we win."). He points out that unconditionals share "a certain kind of not mattering" or an indifference implication. If we schematize an unconditional construction as "unconditional-adjunct + main-proposition", the main message of the construction is that the alternatives listed in the unconditional adjunct do not make a difference with respect to the truth of the main proposition. In other words, the truth of the main proposition does not depend on the truth of the alternatives in the unconditional adjunct.

An alternative unconditional is considered to be equivalent to the conjunction of two conditional statements with the alternatives as their antecedents. This means that a construction such as "Whether X or Y, Z" is equivalent to "If X, Z, and if Y, Z." For example, "Whether it rains or snows, I won't go outside." can be paraphrased as "If it rains, I won't go outside, and if it snows I won't go outside." Notice that the unembedded connective used in the paraphrase is *and* rather than *or*. The paraphrase with *or* would be too weak: "if it rains, I won't go outside, or if it snows I won't go outside". The unconditional does not communicate that at least one of the conditional statements are true; it communicates that both are true.  Therefore, the interpretation of *or* embedded under an unconditional adjunct surfaces as conjunction rather than disjunction.

@rawlins2013conditionals's analysis of alternative unconditional has four important components. First that the whole construction is a conditional ($X\rightarrow Y$) such that the adjunct ($X$) restricts the modal base of the main clause ($Y$.) In other words, a sentence like "if you stay, you will have fun" asserts that "you will have fun" is true in the worlds where "you stay" is true.  Second, the antecedent of this conditional is a disjunction (If A or B then Y.) Third a disjunction such as "A or B" denotes the set of propositions in the disjunction ($\{A, B\}$). Fourth, the set of propositions provides sequential restrictions on the modal base in the main clause. For example, a sentence like "whether you stay or go, you will have fun" asserts that "you will have fun" is true first in worlds that "you stay" and then in worlds where "you go". The conjunctive interpretation of disjunction in alternative unconditionals follows from this sequential domain restriction. 

## Discussion

This chapter provided a brief introduction to disjunction in logic as well as previous approaches to the semantics of *or* in English. In logic, disjunction has always faced two alternate definitions: exclusive and inclusive. While different eras have assigned primacy to one or the other definition, logical systems in the past century or so have adopted the inclusive definition more often. In natural language, disjunction words like *or* are associated with (at least) five implications: inclusivity, exclusivity, conjunctivity, ignorance, and indifference. Table (\@ref(tab:implicationsall)) shows 10 example sentences in different linguistic environments and marks the presence and absence of each implication. I discussed a range of factors that may affect the interpretation of *or* including conversational (Gricean) principles, entailment environment, semantic relation of the disjuncts, syntactic category of the disjuncts, question intonation, and a range of linguistic constructions such as the ones with possibility modals, embedded imperatives, repairs, definitions, and unconditionals. 

|Example| IOR | XOR | AND | Ignorance | Indifference |
|----------------------------------------|:-:|:-:|:-:|:-----:|:-------:|
|Have some food or drinks! |$\checkmark$|X|X|X|$\checkmark$|
|Bob studied yesterday or the day before. |X|$\checkmark$|X|$\checkmark$|X|
|Students like Bob or Becky will pass. |X|X|$\checkmark$|X|$\checkmark$|
|I didn't see Bob or Becky study at all.|X|X|$\checkmark$|X|X|
|Did Bob study yesterday, or Becky? |X|$\checkmark$|X|$\checkmark$|X|
|Did either Bob or Becky study yesterday?|$\checkmark$|X|X|$\checkmark$|X|
|Study hard or you will fail! |X|$\checkmark$|X|X|X|
|Bob could study or play soccer. |X|X|$\checkmark$|X|X|
|Bob studies language acquisition or language development. |X|X|$\checkmark$|X|X|
|Becky studies phonetics or semantics; not sure which. |X|$\checkmark$|X|$\checkmark$|X|

Table:(\#tab:implicationsall) Implications of *or* for several example sentences.

This chapter also discussed two approaches to lexical meaning: lexical ambiguity approaches and Gricean approaches. The crucial difference between them is their adherence to a principle of parsimony for meanings, namely Grice's "Modified Occam's Razor". Grice's razor proposes that lexical meaning should not be multiplied beyond necessity. Lexical ambiguity approaches are more relaxed about Grice's razor and may consider different interpretations of a word as different meanings of the word. Gricean approaches strive to only assign an interpretation as lexical meaning when it cannot be explained by other independently motivated interpretive factors. For example, a lexical ambiguity account considers *or* ambiguous between the five implications mentioned earlier and possibly more. Context of the utterance helps speakers disambiguate the intended meaning of *or*. The Gricean approach proposes a single (more general) meaning for *or*  and derives its various interpretations by strengthening (making more specific) that single meaning.

<!--
The research in semantics and pragmatics shows that the interpretation of *or* is highly sensitive to its linguistic and non-linguistic environment. Acquisition theories should take this into account. We can use these factors as cues that help the acquisition as well as the interpretation of disjunction.
-->
How does research in semantics and pragmatics of disjunction inform accounts of disjunction acquisition? The literature on semantics and pragmatics of *or* provides a list of factors that affect the interpretation of a disjunction word like *or*. This list can be considered as a set of candidate cues that children may use to learn the interpretation and ultimately the meaning of a disjunction word. It is important to note that both lexical ambiguity and Gricean accounts rely on such cues to derive the intended interpretation. Ambiguity accounts use them to disambiguate among a set of learned lexical meanings while Gricean accounts need them as part of the reasoning process that derives the intended interpretation from a unified lexical meaning for *or*. While previous research in semantics and pragmatics has successfully listed some of the main factors in interpreting *or*, it has not yet addressed how these factors can potentially be weighted for their role in generating the observed interpretations. For example, it is not clear whether conversational factors or the semantics of the disjuncts play a more important role in the interpretation of a disjunction. Chapter \@ref(corpus) investigates the role of a subset of these factors in child-directed speech and Chapter \@ref(modeling) provides computational models that show how these factors can be weighted to better understand their role in deriving different interpretations of disjunction in child-directed speech.

<!--
Perhaps the most important conclusion in this section is that the interpretation of *or* relies on the linguistic and non-linguistic context of the utterance. While this is perhaps true of many (and possibly all) lexical items, *or* presents an extreme case. 

It is important to note that the Gricean approach does not reject lexical ambiguity altogether but rather treats it like a default or null hypothesis. Given multiple interpretations of a word, it can be assumed that the word is ambiguous until the interpretations can be derived using independently motivated interpretive factors and mechanisms. In that case, the lexical ambiguity of the word is rejected. It is not clear which approach is better to adopt and many arguments developed for the Gricean approach have faced theoretical challenges. For the research on linguistic meaning the choice of approach may not have major consequences and may depend on the theoretical and methodological taste of the semanticist. On the other hand, the choice of approach will have important consequences for research on the acquisition of meaning. For example, @bontly2005modified argues that while Grice's razor remains a theoretical issue in semantics and pragmatics, in language acquisition, it receives empirical support. Several studies have suggested that acquisition of meaning is conservative [@mazzocco1997children;]. Children often assume a one-to-one map between forms and meanings and find it difficult to map additional meanings to words they already know.-->
