# Cue-based Acquisition of Disjunction  {#modeling}

```{r global_options6, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, 
                      fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=TRUE, 
                      message=F, sanitize = T)
```

## Introduction

In Chapter \@ref(sempragLit), I reviewed the complexities involved in interpreting a disjunction word such as *or* in English. I showed that a disjunction can be interpreted as inclusive, exclusive, and even conjunctive. In adition to these truth-conditional interpretations, a disjunction is sometimes associated with speaker ignorance/indifference as well. Given the wide range of interpretations that *or* can have, how can children learn to interpret it correctly? In Chapter \@ref(devoLit), I presented the results of @morris2008logically that suggested children rarely hear the word *or*; and when they do, they largely hear the exclusive interpretation (75-80\% of the time). The annotation study of Chapter \@ref(corpus) replicated @morris2008logically's finding. In Chapter \@ref(devoLit), I also reported the results of recent comprehension studies that show children between the ages of three and five can interpret *or* as inclusive disjunction [@crain2012emergence]. Chapter \@ref(comprehension) confirmed that children between three and five can interpret *or* as inclusive in simple existential sentences. The finding of the comprehension studies and the corpus studies taken together present a learning puzzle: how can children learn to interpret *or* as inclusive if they mostly hear exclusive examples? This chapter provides a solution by developing a cue-based account for children's acquisition of connectives. More generally, the account proposed in this chapter is helpful for learning words with multiple interpretations when one interpretation dominates the learner's input. 

Learning from multiple cues is a common approach in language acquisition [See @monaghan2014multiple for an overview]. In the domain of function word semantics, @wynn1992children proposed a cue-based account for the acquisition of number words. In the next section, I briefly review their proposal and report their findings. The annotation study in Chapter \@ref(corpus) used a methodology similar to that of @wynn1992children and reported several cues that may help children's acquisition of the connectives *and*/*or*. In this section, I use the data in my annotation study to present a cue-based account of connective acquisition. This account provides a straightforward solution to the learning puzzle of disjunction. I support this claim using three modeling experiments. The models incorporate the proposed cues to learn decision trees that predict the interpretation of a disjunciton/conjunction with an accuracy of above 80%. I end the chapter with a discussion of the assumptions and limitations of the proposed account.

## The Cue-based Account for Number Words 

Research on children's acquisition of numeral words (e.g. *one*, *two*, *three*, etc.) has suggested that children initially know that number words more than "one" refer to precise numerosities but they do not know exactly which number refers to which word [@wynn1992children]. @bloom1997linguistic searched for linguistic cues that could help children associate numerals with quantity and numerosity. They considered two classes of cues: syntactic and semantic. Syntactic cues to word meaning were first discussed by @brown1957linguistic. He wrote: "If a part of speech has reliable semantic implications, it could call attention to the kind of attribute likely to belong to the meaning of the word ... the part of speech memebership of the new word could operate as a filter selecting for attention probably relevant features of the nonlinguistic world." He tested preschoolers with nonce constructions "to sib", "a sib", and "any sib" showing that children can use the modifying function words to decide witheher the nonce word *sib* should refer to an action, an object, or a substance. 

Semantic cues, on the other hand, are provided by the meaning of the known words in the sentence. Consider the sentence "there were several gloobs." The use of *gloob* in the plural noun phrase "several gloobs" makes it possible to infer that "gloob" is not an action or a spatial relationship but rather an entity that can have multiple instances. Using only the syntactic cues, there still remains a wide range of referential uncertainty since a gloob may be anything from an egg to an alien creature. Now consider the sentence "I ate several gloobs for breakfast." What a gloob may be is now restricted to edible entities, probably those that are suitable for breakfast. The meanings of the verb *eat* and the adverbial phrase "for breakfast" help us further narrow down the possible meanings for *gloob*. 

It is not always easy to tell whether a cue is syntactic or semantic. Here I avoid this issue by using the term "compositional cues" to refer to both syntactic and semantic cues that aid the interpretation of an unknown word. Using the term "compositional" also brings into attention the fact that syntactic and semantic cues are interrelated and do not act in an independent or unstructured way. Consider the sentence "After eating breakfast, I saw several gloobs." Even though the words *eat* and *breakfast* are present in the sentence, they do not restrict the possible meanings for *gloob* as they did before. In other words, it is not the mere presence of these words in the sentence that act as cues but rather the way they combine with the unknown word to convey the main communicative message of the utterance.

@bloom1997linguistic proposed that children learn number word meanings by attending to the compositional cues that accompany number words such as the words' ordering relative to other words, other funciton words they co-occur with, and the count-mass status of the nouns they modify. They specifically discussed four cues. Two cues could help children notice that number words pattern like quantifiers. First, similar to quantifiers, number words precede adjectives and do not follow them. Second, they participate in the "... of the Xs" construction: "one of the gloobs", "some of the gloobs", "most of the gloobs", etc. The third cue is the co-occurrence of number words with count nouns. This cue can inform learners that their meaning is restricted to the quantification of individuals. Finally, unlike other adjectives, numerals cannot be modified further using an adverb such as *very* or *too* ("very big animals" vs. "\*very two animals"). According to @bloom1997linguistic, this cue can help a learner understand that number words pick an absolute property of a set rather than a continuous one.

Using the data available in the CHILDES corpora, @bloom1997linguistic investigated the presence of cues to number word meaning in chlild directed speech for three children between the ages of one and three. They found that these children and their parents only use number words with count nouns; they do not use number words with modifiers and only use them before adjectives, not after. Finally, they found that these children and their parents use only number words and quantifiers in the partitive construction and not adjectives. The results of their corpus study shows that the compositional cues they proposed for number word acquisition are available in children's linguistic input. In the next section, I discuss some compositional cues that can help a learner limit the hypothesis space to connective meanings for coordinators such as *and*, *or*, *but*, *so*, etc.

## Cues to coordinator meanings

```{r data_prep, echo=FALSE, warning=FALSE, message=FALSE}
alex_data <- read.csv("decision_trees/annotation_data/ProvidenceData-alex.csv", nrows=100)
lily_data <- read.csv("decision_trees/annotation_data/ProvidenceData-lily.csv", nrows=99)
vio_data <- read.csv("decision_trees/annotation_data/ProvidenceData-vio.csv", nrows=102)
will_data <- read.csv("decision_trees/annotation_data/ProvidenceData-will.csv", nrows=100)
naima_data <- read.csv("decision_trees/annotation_data/ProvidenceOR-Naima.csv", nrows=241)

naima_data <- 
  naima_data %>%
  filter(!is.na(exclusivity), !is.na(intonation))

all_data <- rbind(alex_data, lily_data, vio_data, will_data, naima_data)

all_data$intonation <- as.factor(all_data$intonation)

all_data$exclusion <- fct_recode(all_data$exclusion, "Consistent" = "ELS", "Inconsistent" = "EXC")
all_data$intonation <- fct_recode(all_data$intonation, "Flat" = "0", "Rising" = "1", "Rise-Fall" = "2")
all_data$intonation <- fct_recode(all_data$intonation, "Flat" = "0", "Rising" = "1", "Rise-Fall" = "2")
```

```{r data_wrangling, echo=FALSE, warning=FALSE, message=FALSE}
raw_data <- 
  all_data %>%
  group_by(id, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX) %>%
  gather(EXIN, counts, EX:IN) %>%
  mutate(prop = counts / total) %>%
  group_by(EXIN) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

graph_data <- 
  all_data %>%
  group_by(id, intonation, exclusion, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = EX + IN) %>%
  gather(exclusivity, counts, EX:IN) %>%
  mutate(prop = counts / total) %>%
  group_by(exclusivity, intonation, exclusion) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop)) 

counts_table<-  
  all_data %>%
  group_by(intonation, exclusion, syn_level, exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX)

exclusivity_overall <-
  all_data %>%
  group_by(exclusivity) %>%
  summarise(counts= n()) %>%
  spread(exclusivity, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = IN + EX) %>%
  gather(exclusivity, counts, EX:IN) %>%
  mutate(prop = counts / total)

cds_disjunction_model <- summary(glmer(exclusivity ~ intonation + exclusion + (1|id), family="binomial", data=all_data))
```

```{r importSyntaxInfo}
annotations <- read_csv("connective_learning/3_providence_annotations/providence_merged.csv")

synTable <-
annotations %>%
  group_by(syn_level) %>%
  summarize(counts=n(), total = nrow(annotations), prop = counts/total)
```

Three important compositional cues can help learners in restricting their hypotheses to coordinator meanings. First, as pointed out by @haspelmath2007, coordination has very specific compositional properties. Coordinators combine two or more units of the same type and return a larger unit of the same type. The larger unit has the same semantic relation with the surrounding words as the smaller units would have had without coordination. These properties separate coordinators from other function words such as articles, quantifiers, numerals, prepositions, and auxiliaries which are not used to connect sentences or any two similar units for that matter. In fact, the special syntactic properties of coordinators have compelled syntactic theories to consider specific rules for coordination.  

The literature on syntactic bootstrapping suggests that children can use syntactic properties of the input to limit their word meaning hypotheses to the relevant domain [@brown1957linguistic; @gleitman1990structural; see @fisher2010syntactic for a review]. In our `r nrow(annotations)` annotations of conjunction and disjunction, we found that `r round(synTable$prop[2]*100)`% of the times *and*/*or* connected sentences or clauses. This pattern is unexpected for any other class of function words and it is possible that the syntactic distriution of coordinators cue the learners to the space of sentential connective meanings. 

Second, in our annotation study we found that *and* never occurs with inconsistent coordinands (e.g. "clean and dirty") while *or* commonly does (e.g. "clean or dirty"). The inconsistency of the coordinands can cue the learner to not consider conjunction as a meaning for the coordinator given that a conjunctive meaning would too often lead to a contradiction at the utterance level. On the other hand, choosing disjunction as the meaning avoids this problem. Third, the large scale study of Chapter \@ref(corpus) found that *or* is more likely to occur in questions than statements while *and* is more likely in statements. Since questions often contain more uncertainty while statements are more informative, it is possible that these environments bias the learner towards selecting hypotheses that match this general communicative function. Disjunction is less informative than conjunction and it is possible that the frequent appearance of *or* in questions cues learners to both its meaning as a disjunction as well as the ignorance inference commonly associated with it.

Finally, it is reasonable to assume that not all binary connective meanings shown in Figure \@ref(fig:binaryLogicalConnectivess) are as likely for mapping. For example, coordinators that communicate tautologies or contradictions seem to be not good candidates for informative communication. Similarly, if A coordinated with B simply asserts the truth of A and says nothing about B, it is unclear why it would be needed if the language already has the means of simply asserting A. It is possible that pragmatic principles already bias the hypothesis space to favor candidates that are communicatively more efficient.

```{r binaryLogicalConnectivess, fig.env="figure", fig.align="center", fig.width=5, fig.height=2, fig.cap="The truth table for the 16 binary logical connectives. The rows represent the set of situations where zero, one, or both propositions are true. The columns represent the 16 possible connectives and their truth conditions. Green cells represent true situations."}
binary_connectives<- png::readPNG("figs/binary_connective.png")
grid::grid.raster(binary_connectives)
```

Even though these findings are suggestive, they need to be backed up by further observational and experimental evidence to show that children do actually use these cues in learning connective meanings. In the next section, I turn to the more specific issue of learning the correct interpretation of *and*/*or* from the input data. Similar to the case of number words, previous research has provided insight into how children comprehend a disjunction and what they hear from their parents. The main question is how children learn what they comprehend from what they hear. I turn to this issue in the next section.

## Learning to interpret *and*/*or*: A cue-based account {#myaccount}

Previous comprehension studies as well as the one reported in Chapter \@ref(comprehension) have shown that children as early as age three can interpret a disjunction as inclusive. However, @morris2008logically's study of child-directed speech showed that exclusive interpretations are much more common than other interpretations of disjunction in children's input. In Figure \@ref(fig:interpretation), I show the results of Chapter \@ref(corpus)'s annoation study by grouping the disjunction interpretations into exclusive (EX) and inclusive (IN), i.e. non-exclusive categories. These results replicate @morris2008logically's finding and reinforce a puzzle raised by @crain2012emergence: How can children learn the inclusive interpretation of disjunction when the majority of the examples they hear are exclusive? To answer this question, I draw insight from the Gricean approach to semantics and pragmatics discussed in Chapter \@ref(sempragLit). 

```{r interpretation, fig.env='figure',fig.width=1.8, fig.height=1.8, fig.align="center",fig.cap = "Proportion of exclusive and inclusive interpretations of disjunction in child-directed speech. Error bars represent bootstrapped 95\\% confidence intervals."}
ggplot(raw_data, aes(x= EXIN, y=prop, fill=EXIN)) + 
  theme_few(base_size = 10) + 
  geom_col(width=0.7) +
  geom_linerange(aes(ymax = cih, ymin = cil)) + 
  guides(fill=FALSE) +
  labs(title = "", x = "", y = "Proportion") + 
  theme(text = element_text(family = "Times"))+
  scale_colour_solarized()
```

Research in Gricean semantics and pragmatics has shown that the word *or* is not the only factor relevant to the interpretation of a disjunction. It is not only the presence of the word *or* that makes us interpret a disjunction as inclusive, exclusive, or conjunctive, but rather the presence of *or* along with several interpretive factors such as intonation [@pruitt2013interpretation], the meaning of the disjuncts [@geurts2006exclusive], and the conversational principles governing communication [@grice1989studies]. Therefore, the interpretation and acquisition of the word *or* cannot be separated from all the factors that accompany it and shape its final interpretation. 

In the literature on word learning and semantic acquisition, form-meaning mapping is construed as mapping an isolated form such as *gavagai* to an isolated concept such as "rabbit". While this approach may be feasible for content words, it will not work for function words such as *or*. First, the word *or* cannot be mapped isolated from its formal context. As @pruitt2013interpretation showed, the intonation that accompanies a disjunction affects its interpretation. Therefore, a learner needs to pay attention to the word *or* as well as the intonation contour that accompanies the disjunction. Second, the word *or* cannot be mapped to its meaning isolated from the semantics of the disjuncts that accompany it. As @geurts2006exclusive argued, the exclusive interpretation is often enforced simply because the options are incompatible. For example, "to be or not to be" is exclusive simply because one cannot both be and not be. In addition, conversational factors play an important role in the interpretation of *or* as @grice1989studies argued. In sum, the interpretation and acquisition of function words such as *or* require the learner to consider the linguistic and nonlinguistic context of the word and map the meanings accordingly.

Previous accounts have adopted a model in which a function word such as *or* is mapped to its most likely interpretation: *or* $\rightarrow \oplus$. This model is often used in cross-situational accounts of content words. Here I argue that the direct mapping of *or* to its interpretation without consideration of its linguistic context is the primary cause of *or*'s learning puzzle. Instead, I propose that the word *or* is mapped to an interpretation in a context-dependent manner, along with the interpretive cues that accompany it such as intonation and disjunct semantics: 

[connective: *or*, Intonation: rise-fall, Disjuncts: inconsistent] $\rightarrow \oplus$

[connective: *or*, Intonation: rising, Disjuncts: consistent] $\rightarrow \lor$

Figure \@ref(fig:interpretationByIntonationAndConsistency) shows that the rate of exclusive interpretations change systematically when we break the data in Figure \@ref(fig:interpretation) down by intonation and consistency. Given a rise-fall intonation contour, a disjunction is almost always interpreted as exclusive. Similarly, if the propositions are inconsistent, the disjunction is most likely interpreted as exclusive. When either of these two features are absent, a disjunction is more likely to receive an inclusive interpretation.

In this account, it is not a single word that gets mapped to an interpretation but rather a cluster of features. This method has two advantages. First, it deals with the context dependency of disjunction interpretation. The learner knows that *or* with some intonation has to be interpreted differently from one with another. Second, it allows the learner to pull apart the contribution of *or* from the interpretive cues that often accompany it. In fact, analysis of all mapping clusters in which *or* participates and generalization over them can help the learner extract the semantics of *or* the way it is intended by Gricean semanticists. For those skeptical of such an underlying semantics for *or*, there is no need for further analysis of the mapping clusters and the meaning of *or* as a single lexical item is distributed among the many mappings in which it participates. In the next section, I implement this idea using decision tree learning.

```{r interpretationByIntonationAndConsistency, fig.env='figure', fig.width=3.5, fig.height=2.5, fig.align="center", fig.cap = "Exclusive and inclusive interpretations broken down by intonation (flat, rise, rise-fall) and consistency. Error bars represent bootstrapped 95\\% confidence intervals."}
ggplot(graph_data, aes(x= exclusivity, y=prop, fill=exclusivity)) + 
  geom_col(width=0.7) +
  geom_linerange(aes(ymax = cih, ymin = cil)) +
  guides(fill=FALSE) +
  theme_few(base_size = 10) + 
  labs(title = "", x = "", y = "Proportion") + 
  scale_colour_solarized() + facet_grid(exclusion~intonation) + theme(text = element_text(family = "Times"))
```


## Modeling Using Decision Tree Learning {#DecisionTrees}

A decision tree is a classification model structured as a hierarchical tree with nodes, branches, and leaves [@breiman2017classification]. The tree starts with an initial node, called the root, and branches into more nodes until it reaches the leaves. Each node represents the test on a feature, each branch represents an outcome of the test, and each leaf represents a classification label. Using a decision tree, observations can be classsifed or labeled based on a set of features. For example, we can make a decision tree to predict whether a food item is a fruit or not based on its color (green or not) and shape (round or not). An example decision tree is the following: at the root, the model can ask whether the item is green or not. If yes, the model creates a leaf and labels the item as "not fruit". If no, the model creates another node and asks if the item is round. If yes, the item is classified as a "fruit" and if not it is classifed as "not fruit". 

Decision trees have several advantages for modeling cue-based accounts of semantic acquisition. First, decision trees use a set of features to predict the classification of obsevations. This is analogous to using cues to predict the correct interpretation of a word or an utterance. Second, unlike many other machine learning techniques, decision trees result in models that are interpretable. Third, the order of decisions or features used for classification is determined based on information gain. Features that appear higher (earlier) in the tree are more infomrative and helpful for classification. Therefore, decision trees can help us understand which cues are probably more helpful for the acquisition and interpreation of a word.

Decision tree learning is the construction of a decision tree from labled training data. In this section, I apply decision tree learning to our annotated data by constructing random forests [@ho1995random; @breiman2001random]. In random classification forests multiple decision trees are constructed on subsets of the data and the decisions are made by taking the majority vote. Next section discusses the methods used in constructing the random forests for interpreting connectives *or*/*and*.

### Methods

The random forest models were constructed using python's Sci-kit Learn package [@pedregosa2011scikit]. The annotated data had a feature array and a connective interpretation label for each connective use. Connective interpreations included exclusive (XOR), inclusive (IOR), conjunctive (AND). The features or cues used included all other annotation categories: intonation, consistency, syntactic level, utterance type, and communicative function. All models were trained with stratified 10-Fold cross-validation to reduce overfitting. Stratified cross-validation maintains the distribution of the initial data in the random sampling to build cross validated models. Maintaining the data distribution ensures a more realistic learning environment for the forests. First a grid search was run on the hyperparamter space to establish the number of trees in each forest and the maximum tree depth allowable. The default number of trees for the forests was set to 20, with a max depth of eight and a minimum impurity decrease (i.e. gini decrease) of 0. Decision trees were fit with high, medium, and low minimum gini decrease values. High gini decrease values tested trees that did not use any features for branching. Such trees represent the traditional approach to mapping that directly maps words to their most likely interpretation. Medium minimum gini decrease values allowed for very conservative branching. Such trees represent a cue-based account in which only one or two most informative features are used for interpreation. Finally low minimum gini decrease values allow for less conservative and more elaborate trees with more features. Such trees represent an account that allows for multiple cues to be used for interpretation. 

### Results

I first present the results of the random forests trained to classify exclusive and inclusive interpretations of disjunction. Figure \@ref(fig:binaryNoFeatures) shows the decision tree with high minimum gini decrease. As expected, a learner that does not use any cues would interpret *or* as exclusive. Figure \@ref(fig:binaryLowFeatures) shows the most accurate decision tree with medium minimum gini decrease. The tree has learned to use intonation and consistency to classify disjunctions as exclusive or inclusive. As expected, if the intonation is rise-fall or the disjuncts are inconsistent, the interpretation is exclusive. Otherwise, if the intonation is not rise-fall and the disjuncts are consistent, then the disjunction is classified as inclusive. Figure \@ref(fig:binaryHighFeatures) presents the most accurate decision tree with low minimum gini decrease. In addition to using intonation and consistency like the previous tree, the tree has learned to use communicative functions further identify exclusive interpretations. Conditionals (e.g. be careful or you'll get hurt) and descriptions are found more likely to be exclusive. Figure \@ref(fig:binaryLearningCurve) shows a cross-validated learning curve of the forest with low minimum gini decrease on increasingly large training sets. The forest surpasses mean accuracy of 80% after training on about 20 data points.

```{r binaryNoFeatures, fig.cap="Tree grown with minimum impurity decrease of 0.1. The tree always classifies examples of disjunction as exclusive.", fig.width=2}
binaryNone <- readJPEG("random_forests/exin_0.1.jpg")
grid::grid.raster(binaryNone)
```

```{r binaryLowFeatures, fig.cap="Tree grown with minimum impurity decrease of 0.05. The tree classifies examples of disjunction with rise-fall intonation as exclusive (intonation > 1.5). If the intonation is not rise-fal but the disjuncts are inconsistent (consistency < 0.5), then the disjunction is still classified as exclusive. However, if neither of these two holds, the disjunction is classified as inclusive."}
binaryLow <- readJPEG("random_forests/exin_0.05.jpg")
grid::grid.raster(binaryLow)
```

```{r binaryHighFeatures, fig.cap="Tree grown with minimum impurity decrease of 0.01. The tree classifies examples of disjunction with rise-fall intonation as exclusive (intonation > 1.5). If the intonation is not rise-fal but the disjuncts are inconsistent (consistency < 0.5), then the disjunction is still classified as exclusive. Even if the disjuncts are conditional but the communicative function is conditional (speech_act_conditional > 0.5), the classification is again exclusive. A similar decision is made for the communicative function of description and the rest of the disjunctions that do not satisfy any of these conditions are classified as inclusive.", fig.asp=1.5}
binaryHigh <- readJPEG("random_forests/exin_0.01.jpg")
grid::grid.raster(binaryHigh)
```

```{r binaryLearningCurve, fig.cap="Cross-validated learning curve of the forest with binary exclusive-inclusive classification and with low minimum gini decrease."}
binaryCurve <- readPNG("random_forests/exin_curve.png")
grid::grid.raster(binaryCurve)
```

Next I use decision tree learning for learning conjunction as well as disjunction and distinguish between inclusive, exclusive, and conjunctive interpretations. Figure \@ref(fig:ternaryNoFeatures) shows the example tree with high minimum gini decrease, which only uses the presence of the words *or*/*and* as the cue to interpret conjunction and disjunction. As expected, the tree interprets a coordination with *and* as a conjunction and one with *or* as exclusive disjunction. Figure \@ref(fig:ternaryLowFeatures) shows the highest scoring decision tree with medium minimum gini decrease. In addition to the presence of *and*/*or*, the tree uses intonation, consistency, and communicative function to distinguish exclusive and inclusive uses of disjunction. A disjunction that is not rise-fall, inconsistent, and conditional is classified as inclusive. Figure \@ref(fig:ternaryHighFeatures) presents the highest scroing decision tree with low minimum gini decrease. This tree is almost identical to the tree in Figure \@ref(fig:ternaryLowFeatures), except that descriptions are also classified as exclusive. Figure \@ref(fig:ternaryCurve) shows a cross-validated learning curve of the forest with low minimum gini decrease on increasingly large training sets. The forest surpasses mean accuracy of 80% after training on about 40 data points. 

```{r ternaryNoFeatures, fig.cap="Tree grown on conjunctions and disjunctions with minimum impurity decrease of 0.05. The tree uses the words and/or and classifies them as conjunction and exclusive disjunction."}
TernaryNone <- readJPEG("random_forests/exinand_0.05.jpg")
grid::grid.raster(TernaryNone)
```

```{r ternaryLowFeatures, fig.cap="Tree grown on conjunctions and disjunctions with minimum impurity decrease of 0.015. In addtion to the words and/or, the tree uses intonation, consistency, and the communicative function of conditionals to detect exclusive uses of disjunction."}
TernaryLow <- readJPEG("random_forests/exinand_0.015.jpg")
grid::grid.raster(TernaryLow)
```

```{r ternaryHighFeatures, fig.cap="Tree grown on conjunctions and disjunctions with minimum impurity decrease of 0.005. After using the words and/or, intonatio, and consistency, the tree uses communicative functions of conditional and descreiptions to detect exclusive disjunctions."}
TernaryHigh <- readJPEG("random_forests/exinand_0.005.jpg")
grid::grid.raster(TernaryHigh)
```

```{r ternaryCurve, fig.cap="Cross-validated learning curve of the forest with binary exclusive-inclusive classification and with low minimum gini decrease."}
TernaryCurve <- readPNG("random_forests/exinand_curve.png")
grid::grid.raster(TernaryCurve)
```

## Discussion

In this chapter, I discussed two accounts for the acquisition of function words. The first account is similar to cross-situational content word acquisition, in that words are isolated and directly mapped to their most likely meanings. The second account is what I call cue-based context-dependent mapping in which words are mapped to meanings conditional on a set of present cues in the context. I showed that the first account's direct mapping of forms to meanings runs into a learning problem for the disjunction word *or*. Under this account, the input statistics supports an exclusive interpretation for *or* while in comprehension studies we find children interpreting *or* as inclusive. I showed that the cue-based account resolves this problem by allowing *or* to be mapped to its interpretation according to the contextual cues that disambiguate it. The results of my decision tree learning experiments on the annotated data from child-directed speech suggested that such an approach can successfully learn the exclusive and inclusive interpretations of a disjunction. More broadly, the cue-based context-dependent mapping account proposed in this chapter is useful for the acquisition of ambiguious words and interpretations that are consistent but relatively infrequent in child-directed speech. 

<!--
A major limitation of the model is that it only from about 500 data points. This is substantially fewer number of examples than what children hear until they reach the age of 3. It is reasonable to expect that the model's performance, especially for conjunctive (AND) and negative (NOR) interpretations improves with more example data that allows the model to learn the features that correlate systematically with uncommon interpretations. 

Explain how much you expect these to generalize to other connective words or more generally function words.
-->
<!--
It is also important to think about the role of attention in learning. Currently models assume that each example has equal siginficance to learning and the learner pays equal attention to them. This does not match the intuitive understanding of learning. We dedicate more resources and attention to tasks we have not yet mastered while skills or knowledge we already possess becomes part of the daily routine and we do not need to assign learning resources to them.

In the case of conjunction and disjunction, a large part of the input is very similar and repetetive. Children can learn their meaning and structure quickly and future exmples are not informative or challenging. It is reasonable to assume that children dedicate their resources to the more challenging instances and focus on mastering them. In such a model, learning is surprisal driven. If a usage is novel and unfamiliar, the learner assigns more learning resources to it and focuses on learning them. This way, the learner can move from extremely frequent instances in the input to the rare ones. Under such a model, frequency determines order of acquisition (to some degree) but not whether a construction is learned or not. Under the right circumstances even a one time occurance of a word or construction can result in successful learning. 
-->

<!--
Bloom, Lois; Lahey, Margaret ; Hood, Lois ;  Lifter, Karin,  &  Fiess, Kathleen. 1980. Complex sentences: Acquisition of syntactic connectives 
and the semantic relations they encode." Journal of Child Language 7, no. 2, 235-261.

Clark, Eve V.  1971. On the acquisition of the meaning of before and after. Journal of Verbal Learning & Verbal Behavior 10, no. 3, 266-275.

Coker, Pamela L.  1978. Syntactic and semantic factors in the acquisition of before and after. Journal of Child Language 5, no. 2. 261-277.

Eisenberg, Ann R.  1985.  Learning to describe past experiences in conversation. Discourse Processes 8, no. 2, 177-204.

Emerson, Harriet F.  1979. Children's comprehension of ‘because’in reversible and non-reversible sentences. Journal of Child Language 6, no. 2, 279-300.

Evers-Vermeul, Jacqueline, & Sanders, Ted .  2009.  The emergence of Dutch connectives; how cumulative cognitive complexity explains
 the order of acquisition. Journal of Child Language 36, no. 4, 829-854.

Feagans, Lynne.  1980.  Children's understanding of some temporal terms denoting order, duration, and simultaneity." Journal of 
Psycholinguistic Research 9, no. 1, 41-57.

Ferreiro, Emilia, & Sinclair, Hermina .  1971.  Temporal relationships in language. International Journal of Psychology 6, no. 1, 39-47.

Friedman, William J., & Seely, Pamela B.  1976. "The child's acquisition of spatial and temporal word meanings." Child Development , 47, 1103-1108.

Johnson, Helen L.  1975. The meaning of before and after for preschool children. Journal of Experimental Child Psychology 19, no. 1, 88-99.

Peterson, Carole, & McCabe, Allyssa .  1987.  The connective ‘and’: Do older children use it less as they learn other connectives?.Journal of Child Language 14, no. 2, 375-381.

Silva, Marilyn N.  1991. Simultaneity in children's narratives: the case of when, while and as. Journal of Child Language 18, no. 3,  641-662.

Stevenson, Rosemary J., & Pollitt, Caroline .  1985.  The acquisition of temporal terms. Journal of Child Language 14, no. 3,  533-545.
-->
