---
bibliography: dissertation_biblio.bib
output:
  pdf_document: default
  word_document:
    fig_caption: yes
---
# Comprehension of Disjuntion in Preschool Children {#comprehension}

```{r comments4, eval=FALSE, echo=FALSE}
# Introduction: 
#   * Find more citation for the claim that children interpret "or" as inclusive.
#   * Systmetaic convention of italics vs. "" for marking words in the object langauge and guessing categories.
#   * Explain the ignorance and exclusivity implicatures of "or" and that they should be separated. (in the semantics and pragmatics chapter?)
# Change binary and ternary to 2AFC and 3AFC

# Study1:
#   * Revise study 1 to emphasize that adults do not reject disjunctive guesses. Therefore, probably better to say that their interpretation is inclusive overall yet showing signs of infelicity when both disjuncts are true.

# Study2: 
#   * Figure 4.13: Remove A little right
#   * Report exclusions

# Study3:
#   * multilevel logistic for feedback categories, test correction increase
#   * Agreement Kappa for Judgment task and open-ended feedback
# Are there children that systematically treat conjunction and disjunction similarly?

# Write General Discussion
#   * Implicature:
#     +not clear on the real rate of exclusivity implicatures with "or". Previous studies often do not separete ignorance from exclusivity.
#   * Important: point out the issue with "the access to alternative" account in your setup.
# 
# * A discussion of possible confounds in the experiment: 
#     + An important factor in the rate of implicature is the speaker commitment in their guesses and predictions. There is no reason to believe that adults and children assume the same level of commitment in a guessing game and therefore they may judge things more leniently or

######## Mike's Comments:

# layout the implicature vs. infelicity argument
  #clarify what you mean by implicature rate

# Analogy, imagine measuring line orientation: adults and children are asked to detect a change in line orientation. The ability to detect the orientation depends on: contrast of the stimulus, the speed, asking children did the orientation of the line change, children may think did it change a lot but adults may 

### My own corrctions:
# Figure 4.2 could use labels simple, conjunctive, disjunctive

```

```{r global_options4, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, 
                      fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=TRUE, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(grid)
library(irr)
library(lme4)
library(gtable)
library(png)
library(jpeg)
library(ggplot2)
library(xtable)
library(tidyverse)
library(bootstrap)
library(ggthemes)
library(forcats)
```

```{r bootstrapping}
## for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}
```

```{r preliminary_wrangling}
## reading the processed data
data_main <- read.csv("connective_comprehension/study2/2_processed_data/cardgame_data_processed.csv")

data_main <-
  data_main %>%
  filter(trial_type!= "X.YorZ", trial_type!= "XY.YorZ", trial_type!= "XY.YandZ") %>%
  separate(trial_type, c("Card", "Guess"), sep="\\.", remove=FALSE)

data_adults <-
  data_main %>%
  filter(age_category=="Adults")

data_children <-
  data_main %>%
  filter(age_category=="Children")

data_main_dv3 <- 
  data_main %>%
  filter(dv_type==3)

data_main_dv2 <-
  data_main %>%
  filter(dv_type==2)

n_adults <- data_adults$sid %>% unique() %>% length()  
n_adults2 <- filter(data_main_dv2, age_category=="Adults")$sid %>% unique() %>% length() 
n_adults3 <- filter(data_main_dv3, age_category=="Adults")$sid %>% unique() %>% length()
```

## Introduction

This chapter investigates the differences between adults and children's interpretation of the connectives "and" and "or". Previous research has suggested that adults and children might differ in their interpretation of *or* in two ways. First, unlike adults, children might interpret *or* as logical conjunction, akin to *and* [@Singh2016; @tieu2016]. Second, children might interpret *or* as inclusive disjunction while adults interpret it as exclusive [@crain2012emergence]. In this chapter, I present three studies that assess adults and children's understanding of "and" and "or" using a guessing game paradigm. These studies show that four-year-olds' interpretation of conjunction and disjunction may not be as different from adults as previously supposed.

Study 1 tested adults' interpretations of logical connectives in the context of a guessing game using Two and Three-Alternative Forced Choice judgment tasks (2AFC and 3AFC). The results showed that adults interpret *and* and *or* differently. They interpreted *and* as conjunction and *or* as inclusive disjunction. However, in the task with three alternatives (3AFC) adults did not consider a disjunction felicitious when both disjuncts were true. Comparing the 2AFC and 3AFC results, we find that the felicity of disjunctive statements is sensitive to the measurement. 2AFC task systematically underestimated judgments of felicity and better approximated truth judgments compared to the 3AFC task. This finding is intuitive given that more options provide the opportunity to espress more nuances of linguistic interpretation.

Study 2 investigated children's judgments in the same guessing game as study 1 using a 3AFC task. We used three alternatives to give children a better chance of expressing their pragmatic knowldge and judgments of felicity [-@katsos2011pragmatic]. The study also analyzed and categoried children's open-ended spontatneous feedback to the guesser. Both the 3AFC judgments and the categories of open-ended responses showed that four-year-olds differentiated *or* from *and*. While children's judgments in the 3AFC task showed no sign of infelicity for disjunctive guesses when both disjuncts were true, their open-ended feedback showed that children find such guesses infelicitous. In their open-ended feedback, children expressed that a conjunction in such cases would be more appropriate.

Study 3 used the same paradigm as study 2, but focused on replicating children's open-ended responses and contrasting them with the results of a 2AFC task. Similar to study 2, both truth judgments and open-ended feedback showed that children differentiated *or* from *and*. The 2AFC task showed no evidence that children find disjunctions with true disjuncts infelicitous. However, children's judgments did not differ significantly from those of adults in the 2AFC task of study 1. Similar to study 2, children's open-ended feedback suggested that when both disjuncts are true, children find a disjunctive statement infelicitous and the conjunctive alternative more appropriate. Overall, the results of study 2 and 3 show that forced-choice judgemnet tasks underestimate children's pragmatic competence. Therefore, using open-ended elicitation and analysis of children's feedback **along with** forced choice judgment tasks may provide a better understandindg of children's true semantic and pragmatic knowledge.

Studies reported here also fill two gaps in the literature. First, previous research has focused on children's interpretation of *and* and *or* in complex sentences -- for example with other logical words such as quantifiers *every* and *none*. Here we test children and adults' understanding of *and* and *or* in simple existential sentences like "*There is a cat or a dog*." Second, previous research has mostly tested children and adults using 2AFC truth judgment tasks [@crain1998investigations]. Here we report adults and children's judgments on both 2AFC and 3AFC tasks. We also use children's open-ended spontaneous feedback to develop relevant analytical response categories and we replicate the findings in a following pre-registered study. Katsos \& Bishop [-@katsos2011pragmatic] argued that 3AFC judgment tasks are better suited for assessing children's pragmatic competence. Here we show that even 3AFC tasks can underestimate children's pragmatic knowldge and children's open-ended elicited responses can provide valuable insights not available in forced choice judgment tasks.

## Study 1: Adult Judgments

The goal of this study was to examine adults' interpretations of *and* and *or* in a guessing game as a benchmark for children's interpretations. We designed the study as a card game in which participants saw a picture card, read a description about the card, and had to evaluate the description. In test trials, the descriptions contained the conjunction word *and* and the disjunction word *or*. We tested adults in both two-alternative and three-alternative forced choice tasks (2AFC and 3AFC). This study found that adults interpreted *and* as a conjunction and *or* as an inclusive disjunction. They also considered *or*-statements infelicitous when both disjuncts were true. We also found that the 2AFC and 3AFC tasks register different aspects of adult interpretations: the 2AFC task captures adult intuitions on the basic semantics of the connectives while the 3AFC task is sensitive to pragmatic infelicities as well.

### Methods

#### Materials and Design

```{r stimuli, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=3, fig.height=1.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Cards used in the connective guessing game."}
cards_img <- jpeg::readJPEG("connective_comprehension/study1/0_methods/Stimuli/cards.jpg")
grid::grid.raster(cards_img)
```

The study included six cards with cartoon images of a cat, a dog, and an elephant (Figure \@ref(fig:stimuli)). The study was designed based on the type of cards participants saw and the type of guesses they heard. There were two types of cards: cards with only one animal on them and the ones with two animals. There were three types of guesses: simple (e.g. *There is a cat*), conjunctive (e.g. *There is a cat and a dog*), and disjunctive (e.g. *There is a cat or a dog*). In each guess, the animal labels used in the guess and the animal images on the card may have no overlap (e.g. Image: dog, Guess: *There is a cat or an elephant*), a partial overlap (e.g. Image: Cat, Guess: *There is a cat or an elephant*), or a total overlap (e.g. Image: cat and elephant, Guess: *There is a cat or an elephant*). Crossing the number of animals on the card, the type of guess, and the overlap between the guess and the card results in 12 different possible trial types. We chose 8 trial types (Figure \@ref(fig:trials)), balancing the number of one-animal vs. two-animal cards, simple vs. connective guesses, and expected true vs. false trials. 

```{r trials, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=4, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Trial types represented by example cards and guesses."}
trials_img <- jpeg::readJPEG("figs/trial-types.jpg")
grid::grid.raster(trials_img)
```

#### Participants and Procedure

```{r exampleTrial, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=4, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.cap = "An example trial in the online study."}
trial_img <- jpeg::readJPEG("figs/example_trial.jpg")
grid::grid.raster(trial_img)
```

We used Amazon’s Mechanical Turk (MTurk) for recruitment and the online platform Qualtrics for data collection and survey design. The task took about 5 minutes on average to complete. `r n_adults` English speaking adults participated. `r n_adults2` of them were assigned to a 2AFC judgment task and `r n_adults3` to a 3AFC judgment task. In the 2AFC task, participants had to judge using the options *wrong* and *right*. In the 3AFC task they had to choose betweeen *wrong*, *kinda right*, and *right*. The two conditions were identical otherwise.

The experiment had three phases: introduction, instruction, and test. In the introduction, participants saw the six cards and read that they will play a guessing game with them. Then a blindfolded cartoon character named Bob appeared on the screen and they were told that in each round of the game, they will see a card and Bob is going to guess what animal is on the card. We emphasized that Bob cannot see anything. We asked participants to judge whether Bob's guess was right. In the instruction phase, participants saw an example trial where a card with the image of a dog was shown and the following sentence was written above Bob's head: *Bob: There is a cat on the card*. All participants correctly responded with *wrong* and proceeded to the test phase.^[Randomization code, data, and R code for analyses available at (https://github.com/jasbi/cogsci2017).]

In the test phase, participants saw one trial per trial type. Within each trial type, the specific card-guess scenario was chosen at random. The order of trial types was also randomized. At the end of the study, participants received $0.4 as compensation. Figure \@ref(fig:exampleTrial) shows an example test trial.

Study | N | Age | Mode | DV |
|---|---|---|---|---|
Study 1 - Part 1 | `r n_adults2`| Adults | Online (Mturk) | Wrong, Right|
Study 1 - Part 2 | `r n_adults3`| Adults | Online (Mturk) | Wrong, Kinda Right, Right|
Table: (\#tab:study1info) Summary of Study 1 Methods

### Results 

```{r plots}
prop_table <- 
  data_main %>% 
  group_by(sid, Card, Guess, age_category, dv_type, response) %>%
  summarise(counts= n()) %>%
  spread(response, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = Wrong + Right + `Kinda Right`) %>%
  gather(response, counts, `Kinda Right`:Wrong) %>%
  mutate(prop = counts / total) %>%
  group_by(Card, Guess, age_category, dv_type, response) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

# order the levels as Wrong, Kinda Right, and Right
prop_table$response <- relevel(as.factor(prop_table$response), ref="Wrong")

prop_table$Guess <- fct_recode(prop_table$Guess,  Z="Y")

prop_table  <- rename(prop_table, Options = dv_type, `Age Category` = age_category)

prop_table$Guess <- 
  prop_table$Guess %>%
  factor(levels = c("Z", "X", "YandZ", "YorZ", "XandY", "XorY")) %>%
  fct_recode(Elephant = "Z", Cat = "X", `Cat and Dog`="XandY", `Cat or Dog`="XorY", `Dog and Elephant` = "YandZ", `Cat or Elephant` = "YorZ")
```

#### Two Alternative Judgments

```{r binaryAdultsPlot, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Adults' two-alternative forced choice judgments in the connective guessing game."}
adults_plot2<-
  prop_table %>%
  filter(Options==2, `Age Category`=="Adults") %>%
  filter(response!="Kinda Right") %>%
  ggplot(aes(x=response, y=prop, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(Card~Guess) +
  labs(x="", y="")+
  theme_few() +
  theme(text = element_text(size=17)) +
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE)

cat <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/cat_card.jpg") %>%
  rasterGrob(width=0.9)

cat_dog <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/catdog_card.jpg")%>%
  rasterGrob(width=0.9)

adults_plot_g <- ggplot_gtable(ggplot_build(adults_plot2))

strips <- grep("strip-r", adults_plot_g$layout$name)

new_grobs <- list(cat, cat_dog)

adults_plot_g <- with(adults_plot_g$layout[strips,],
          gtable_add_grob(adults_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

adults_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(adults_plot_g)
```

Figure \@ref(fig:binaryAdultsPlot) shows the results for the 2AFC task. I start from the two left columns which have simple guesses and serve as controls. The results seem straightforward: If the animal mentioned in the guess was not on the card (e.g. elephant), participants judged the guess "wrong"; if the animal was on the card (e.g. cat), participants judged the guess "right". Moving to the next two columns of Figure \@ref(fig:binaryAdultsPlot), we see that the results for our test conditions, namely conjunction and disjunction, match the expectations for logical conjunction and (inclusive) disjunction: An *and*-guess (e.g. cat and dog) is "wrong" if only one of the animals is on the card and "right" if both are on the card. An *or*-guess (e.g. cat or dog) is "right" whether one or both animals are on the card.

#### Three Alternative Judgments

```{r ternaryAdultsPlot, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Adults' three-alternative forced choice judgments in the connective guessing game."}
adults_plot3<-
  prop_table %>%
  filter(Options==3, `Age Category`=="Adults") %>%
  ggplot(aes(x=response, y=prop, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(Card~Guess) +
  labs(x="", y="")+
  theme_few() +
  theme(axis.text.x = element_text(angle=45, hjust = 1, vjust = 1), text = element_text(size=17)) +
  scale_fill_manual(values = c("red4", "grey","springgreen3"), guide=FALSE)

cat <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/cat_card.jpg") %>%
  rasterGrob(width=0.9)

cat_dog <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/catdog_card.jpg")%>%
  rasterGrob(width=0.9)

adults_plot_g <- ggplot_gtable(ggplot_build(adults_plot3))

strips <- grep("strip-r", adults_plot_g$layout$name)

new_grobs <- list(cat, cat_dog)

adults_plot_g <- with(adults_plot_g$layout[strips,],
          gtable_add_grob(adults_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

adults_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(adults_plot_g)
```

```{r Study1TernaryStats}
responses_XY_X_3 <- filter(data_adults, dv_type==3, trial_type=="XY.X")$response
  
XY_X_3_binomTest <-binom.test(x=table(responses_XY_X_3)[1], n=length(responses_XY_X_3), p=0.5, alternative = "greater", conf.level = 0.95)

responses_X_XoY_3 <- filter(data_adults, dv_type==3, trial_type=="X.XorY")$response
  
X_XoY_3_binomTest <-binom.test(x=table(responses_X_XoY_3)[2], n=length(responses_X_XoY_3), p=0.5, alternative = "greater", conf.level = 0.95)

responses_X_XnY_3 <- filter(data_adults, dv_type==3, trial_type=="X.XandY")$response

X_XnY_3_binomTest <-binom.test(x=table(responses_X_XnY_3)[3], n=length(responses_X_XnY_3), p=0.5, alternative = "greater", conf.level = 0.95)

responses_XY_XoY_3 <- filter(data_adults, dv_type==3, trial_type=="XY.XorY")$response

XY_XoY_3_binomTest <-binom.test(x=table(responses_XY_XoY_3)[1], n=length(responses_XY_XoY_3), p=0.5, alternative = "greater", conf.level = 0.95)
```

Figure \@ref(fig:ternaryAdultsPlot) shows the results for the 3AFC judgment task. For four trial types, the results are identical to the 2AFC task: if the animal mentioned in the guess was not on the card (e.g. elephant), participants judged the guess "wrong". If the animal mentioned (e.g. cat) was the only animal on the card, participants judged the guess "right". Finally, if there were two animals and the puppet mentioned them using *and* (e.g. cat and dog), all participants considered the guess "right".

The four remaining trial types showed different patterns of judgments than the ones in the 2AFC task. If the animal mentioned (e.g. cat) was only one of the animals on the card, participant judgments were divided between "right" and "kinda right" (Table \@ref(tab:binomStats), row 1). Also, most adults considered a conjunctive guess (e.g. cat and dog) "wrong", when only one of the animals was on the card (Table \@ref(tab:binomStats), row 2). However, some considered it *kinda right*. When both animals were on the card everyone agreed that the guess was "right" - similar to the results for such trials in the 2AFC task.

With respect to disjunctive guesses like "cat or dog", if the card had only one of the animals, most adults considerd the guess "right" while some considered it "kinda right" (Table \@ref(tab:binomStats), row 3). It is possible that the adults who considered such guesses "kinda right" were sensitive to the underinformative nature of a disjunctive guess when a simple guess would have been more appropriate. If both animals were on the card, adults were split between "kinda right" and "right" responses (Table \@ref(tab:binomStats), row 4). The choice of "kinda right" over "right" in such trials can be interpreted as a sign that adults were sensitive to the infelicity of a disjunction when conjunction was more appropriate. However, the scalar reasoning with *and* and *or* is subtle and in section \@ref(implicature) we discuss the nature of this reasoning in the context of our guessing game.

|Trial Type | $n_{_{right}}/n_{_{total}}$ | $\hat{p}_{_{right}}$ | $p_{_{null}}$ | P-value | $95\%~CI$|
|--------------------|----------|---------|--------|--------|---------|--------|
|Two Animals - Simple | `r table(responses_XY_X_3)[1]`/`r length(responses_XY_X_3)` | `r round(XY_X_3_binomTest$estimate,2)` | `r XY_X_3_binomTest$null.value` | `r round(XY_X_3_binomTest$p.value, 2)`| `r round(XY_X_3_binomTest$conf.int[1],2)`-`r round(XY_X_3_binomTest$conf.int[2],2)`|
|One Animal - AND | `r table(responses_X_XnY_3)[1]`/`r length(responses_X_XnY_3)`| `r round(X_XnY_3_binomTest$estimate,2)` | `r X_XnY_3_binomTest$null.value` | `r round(X_XnY_3_binomTest$p.value, 3)` |`r round(X_XnY_3_binomTest$conf.int[1],2)`-`r round(X_XnY_3_binomTest$conf.int[2],2)`|
|One Animal - OR| `r table(responses_X_XoY_3)[1]`/`r length(responses_X_XoY_3)` |`r round(X_XoY_3_binomTest$estimate,2)` | `r X_XoY_3_binomTest$null.value` | `r round(X_XoY_3_binomTest$p.value, 2)` | `r round(X_XoY_3_binomTest$conf.int[1],2)`-`r round(X_XoY_3_binomTest$conf.int[2],2)`|
|Two Animals - OR|`r table(responses_XY_XoY_3)[1]`/`r length(responses_XY_XoY_3)`|`r round(XY_XoY_3_binomTest$estimate,2)`|`r XY_XoY_3_binomTest$null.value`|`r round(XY_XoY_3_binomTest$p.value, 2)`|`r round(XY_XoY_3_binomTest$conf.int[1],2)`-`r round(XY_XoY_3_binomTest$conf.int[2],2)`|

Table: (\#tab:binomStats) Exact One-Sided Binomial Test

```{r adultsTests, eval=FALSE, fig.env="Table", fig.pos=t, fig.align="center", fig.cap="Some caption"}
(, $n_{_{right}}/n_{_{total}}=$ , $\hat{p}_{_{right}}$ = , $p_{_{null}}$ = , P-value = , $95\%~CI$ = )
```

<!--#### Individual Response Profiles -->

### Discussion

The example sentences bellow show the common interpretations of conjunctive and disjunctive assertions [@aloniStanford].

* Bob is sad *and* angry.
    + Both are true. (Truth Conditional Meaning)

* Bob is sad *or* angry.
    + At least one of the two is true. (Truth Conditional Meaning)
    + Speaker doesn't know which is true. (Ignorance Inference)
    + At most one of the two is true. (Exclusivity Inference)

A conjunctive assertion implies that both propositions are true while a disjunctive assertion implies that at least one is true. These two inferences follow from the classical truth-conditional account of conjunction and disjunction. They constitute the semantics of *and* \& *or*. However, a disjunctive assertion often has two additional inferences: an ignorance inference and an exclusivity inference. These additional inferences are often classified as pragmatic meaning. This section discusses the semantics and pragmatics of *and* \& *or* in the context of study 1's guessing game. [^5]

[^5]: See @gutzmann2014 for a comprehensive discussion of the definitions and boundaries of semantics and pragmatics. Here our definitions and assumptions are close to those of @gazdar79's. 

#### The Semantics of the Connectives

Let us assume that  the semantics of *and* and *or* in simple declarative sentences like "there is a cat and/or a dog" is captured by the logical operators conjunction and inclusive disjunction respectively. A conjunction is true when both conjuncts are true and false otherwise. An inclusive disjunction is true when at least one disjunct is true and false otherwise. Let us also assume that a false statement is judged as "wrong" and a true statement as "right". In the context of study 1, this purely semantic (i.e. truth-conditional) account has two main predictions: 1. Conjunctive guesses like "cat and dog" are wrong when only one of the animals is on the card. 2. Disjunctive guesses are always right because in all such trials at least one of the animals is present on the card. Figure \@ref(fig:binaryAdultsPlot) shows that in 2AFC judgments, both predictions are borne out. In other words, judgments with two alternatives closely match the predictions of a purely semantic or truth conditional account of the connectives *and* and *or*. We can conclude that adults interpreted "right" and "wrong" roughly as "true" and "false" in the 2AFC task.

However, in the 3AFC task, judgments deviated from a purely semantic account in four trial types: 1. disjunction trials with one animal 2. disjunction trials with two animals, 3. conjunction trials with one animal, and 4. the trials with simple guesses when two animals were on the card. Participants often used the third option *kinda right* in these trial types. Other trial types obtained identical results in 2AFC and 3AFC tasks. The comparison of forced choice judgments with two and three alternatives suggests that two alternatives better capture the truth-conditional meaning of the connectives, but underestimate adult pragmatics reasoning.

#### The Pragmatics of the Connectives {#implicature}

```{r scalarTests}
data_adults$binary_response <- fct_recode(data_adults$response, Wrong = "Kinda Right")

or_trials <- filter (data_adults, trial_type == "X.XorY" | trial_type == "XY.XorY")

disjunctTrialsModel <- summary(glmer(binary_response ~ trial_type + (1 + trial_type|sid), family= "binomial", data=filter(or_trials, dv_type==3)))
```

```{r exclusivity, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=4, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Example of cards referred to by a conjunction, inclusive disjunction, and exclusive disjunction."}
excl_diagram <- jpeg::readJPEG("figs/exclusivity_diagram.jpg")
grid::grid.raster(excl_diagram)
```

A disjunctive assertion like "cat or dog" gives rise to an ignorance inference and an exclusivity inference. The ignorance inference is the inference that the speaker does not know which disjunct actually holds. For example in figure \@ref(fig:exclusivity), the speaker does not know whether the card has a cat on it or a dog - s/he is uncertain between cards 1, 2, and 3. The guessing game in this study controls for the ignorance inference by keeping the guesser blindfolded. Therefore, all the disjunctive guesses are evaluated in a context where participants know that the guesser is ignorant of the animals on the cards - both the number of them on the card and their identity. The exclusivity inference is the inference that only one of the disjuncts holds and **not both**. In figure \@ref(fig:exclusivity), a disjunction like "cat or dog" only refers to cards 2 and 3 if it is accompanied with an exclusivity inference. 

Since @grice1975logicconvo, this exclusive interpretation of "or" has been (at least partly) attributed to pragmatic reasoning on the speaker's connective choice. This pragmatic reasoning goes like this: conversational participants are required to make their utterances as informative as possible. In the context of making predictions and guessing, a guesser is required to make their guess as specific (i.e. informative) as possible. [^7] The conjunction word *and* is more specific and informative than the disjunction word *or* [@horn1989natural]. For example in figure \@ref(fig:exclusivity), "cat and dog" picks card 1 while "cat or dog" refers to cards 1, 2, and 3. If a speaker intends to refer to card 1, s/he should use *and* and say "cat and dog". S/he did not use *and*, so s/he probably does not intend to refer to card 1. Following this line of reasoning, we can **exclude** the possibility that the speaker intends to refer to card 1. The term **exclusivity implicature** captures this pragmatic reasoning that results in excluding the possibility of both disjuncts being true. 

[^7]: When you ask someone to predict the outcome of a coin toss, a guess like "it will be heads or tails" does not count as a felicitous guess or prediction.

The goal in this section is to lay out the structure of pragmatic reasoning in our experimental setup and explain how they can be manifested in the results of our experimental studies. There are three main components to the pragmatic reasoning in the guessing game: 1. the assumptions of the game. 2. sensitivity to (under)informativity, and 3. the pragmatic reasoning on the speaker's choice of connectives. Following Katsos \& Bishop [-@katsos2011pragmatic], we have considered **sensitivity to informativeness** as a precondition for **derivation of scalar implicatures**. First we start with the assumptions of the guessing game.

* **Guessing Game Assumptions**:
    + **Ignorance**: the guesser does not know the number or identity of the animals on the card.
    + **Specificity**: A guesser is required to be as specific as possible, ideally referring to a single card.

As explained before, ignorance of the guesser is explicitly explained to the participants. However, specificity is an assumption that was implicit in the game[^8]. All the guesses used in the experiment can pick a single specific card except for disjunctive ones. Conjunctive guesses like "cat and dog" pick specific cards. The simple ones like "cat" can be strengthened pragmatically to mean "only a cat", and pick a specific card. However, Disjunctive ones like "cat or dog" pick two cards in their most specific (exclusive) sense. Therefore, they are always underinformative and violate the specificity assumption.

[^8]: Making this assumption explicit is both hard for the children's age range and almost impossible when disjunctive guesses are used. Disjunctive guesses are always underinformative and never pick a specific card.

* **Sensitivity to Informativeness**: The guesser said "cat or dog" which is underinformative and picks card 1, 2, and 3.
    + **Violation Assumption**: the guesser is violating the specificity requirement.

Participants can detect the underinformativity of disjunctive guesses, notice the violation of specificity, and then decide whether they would like to tolerate this violation or punish it. It should be pointed out that it is hard to distinguish between "tolerating the specificity violation" and simply revising the specificity assumption of the game to avoid a violation. For example, participants may assume that the goal of the game is saying something true about the cards rather than being as specific as possible. In either case, the prediction is that adults who tolerate violation or revise specificity would judge disjunctive guesses as "right". However, if participants assume specificity and decide to not tolerate its violation, they will judge all disjunctive guesses to have some degree of infelicity. Since an underinformative guess is still technically correct, participants may not punish such a guess with a "wrong" response and prefer an intermediate option for like "kinda right". This is what we find in study 1 results. With two alternatives, not many adults judge infelicity with disjunctive guesses and we see almost no "wrong" responses. With three alternatives, "kinda right" responses pop up. Adults responses are split between "kinda right" and "right". 

```{r implicaturePlot, fig.width=6,fig.env="figure",fig.align="center", fig.cap="Adult responses to dicjunction guesses where both discjuncts where true with 2 and 3 options."}
disjunctionPlot <-
  prop_table %>%
  filter(Guess=="Cat or Dog", `Age Category`=="Adults") %>%
  ggplot(aes(x=response, y=prop)) +
  geom_bar(stat = "identity") +
  facet_grid(Card ~ Options, scales = "free") + 
  theme_few() + 
  labs(x="", y="") + 
  geom_linerange(aes(ymin=cil,ymax=cih))

disjunctionPlot_g <- ggplot_gtable(ggplot_build(disjunctionPlot))

strips <- grep("strip-r", disjunctionPlot_g$layout$name)

new_grobs <- list(cat, cat_dog)

disjunctionPlot_g <- with(disjunctionPlot_g$layout[strips,],
          gtable_add_grob(disjunctionPlot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

disjunctionPlot_g$widths[[7]] <- unit(2.5,"cm")

grid.draw(disjunctionPlot_g)
```

If detecting and reacting to underinformativity is the whole story, then disjunctive guesses should show similar degrees of infelicity, regardless of how many animals are on the card. The results of the 3AFC task suggest otherwise. A logistic mixed-effects model with the random intercepts and slopes for subjects and fixed effect of disjunction type found that when comparing disjunctive guesses in the 3AFC task, participants are more likely to choose *kinda right* than *right* when both animals are on the card ($\beta$=`r round(disjunctTrialsModel$coefficients[2],2)`, $z$=`r round(disjunctTrialsModel$coefficients[6],2)`, $p$=`r round(disjunctTrialsModel$coefficients[8],2)`). In other words, participants judged further infelicity with disjunctive guesses that had both disjuncts as true. Therefore, it is reasonable to assume that when both disjuncts where true, some participants went through the following pragmatic reasoning. 

* **Reasoning on Alternatives**: Why did the guesser choose the underinformative connective *or* rather than the more informative one *and*?
    + **Resolution Assumption**: speaker is trying to be as specific as possible by resolving the issue of how many animals are on the card.
        + **Exclusivity Implicature**: Given the resolution hypothesis, if the speaker had decided that two animals were on the card, s/he should have said "cat *and* dog". He did not, so he had decided that only one animal is on the card and **not both**.

```{r rates}
adult_kindaWrong <- filter(prop_table, Guess=="Cat or Dog", Card=="XY", `Age Category`=="Adults", response=="Kinda Right")$prop[2]
  
adult_wrong  <- filter(prop_table, Guess=="Cat or Dog", Card=="XY", `Age Category`=="Adults", response=="Wrong")$prop[1]
```

How does the exclusivity implicature affect participant judgments in the experimental setting? One possibility is that excluding the correct response pragmatically is treated like cases of excluding the right response semantically. For example, guessing "elephant" when there is a cat on the card. The prediction is that disjunctive trials with true disjuncts should receive "wrong" responses. However, this prediction is not borne out. Such disjunctive trials are almost never judged as "wrong". 

Alternatively, it is possible that adults differentiate incorrect pragmatics from incorrect semantics  (i.e. falsehood) and punish incorrect pragmatics less than incorrect semantics. This conclusion is supported by the response patterns across trial types (figure \@ref(fig:ternaryAdultsPlot)). Trial types that have received a "wrong" response are those that are false. Pragmatically infelicitous trial types, namely simple guesses like "cat" or disjunctive guesses like "cat or dog" when both animals are on the card, receive "kinda right" responses. In other words, adults consider false utterances as "wrong" guesses but infelicitous utterances do not reach the level of being "wrong"; they are still right even though not completely right. This explains why the rates of infelicity (avoiding the "right" alternative) differ between 2AFC and 3AFC trials in disjunctive trials with true disjuncts (`r round(adult_wrong,2)` vs. `r round(adult_kindaWrong,2)`). 

<!--
|Problem | Strict Response | Tolerant Response |
|--------|----------------------|------------------|
|Violating Informativity | Kinda Right | Right |
|Excluding the Correct Card (Pragmatically) | Wrong | Kinda Right / Right|

Table: (\#tab:judgmentPrediction) Predictions of Judgments for Scalar Reasoning in Study 1.
-->

```{r implicatureStats, eval=FALSE}

test <- glmer(response ~ Card + (1|sid), family = "binomial" , data = filter(data_main_dv2, Guess=="XorY"))
summary(test)

factor(filter(data_main_dv2, Guess=="XorY")$response)
```

## Study 2: Children's three-alternative judgments and open-ended corrective feedback

```{r}
n_study2 <- length(unique(data_children$sid))

ageInfo_study2 <- c(floor(min(data_children$age_year)), 
                    round((min(data_children$age_year) - floor(min(data_children$age_year)))*12),
                    floor(max(data_children$age_year)),
                    round((max(data_children$age_year) - floor(max(data_children$age_year)))*12),
                    floor(mean(data_children$age_year)),
                    round((mean(data_children$age_year) - floor(mean(data_children$age_year)))*12)
                    )
```

The goal of this study was to examine children's interpretations of *and* and *or* in our guessing game and compare them to those of our adults. Since the 3AFC judgment task in study 1 proved better at capturing the nuances in adults' pragmatic reasoning, we decided to first test children using three alternatives. We also analyzed children's open-ended comments to the guesses in the experimental context. Both three-alternative judgments and the analysis of children's open-ended responses found that children differentiate *and* and *or* statements. The judgment task suggested that children do not consider disjunctive guesses when both disjuncts are true infelicitous. Yet, the analysis of their open-ended feedback showed that they find such guesses infelicitous. We conclude that the 3AFC judgment task can underestimate children's pragmatic competence. 

### Methods


| Study | N | Age | Mode | DV |
|--------------|---|-------------|------------|--------------------------|
| Study 1 - Part 1 | `r n_adults2`| Adults | Online (Mturk) | Wrong, Right|
| Study 1 - Part 2 | `r n_adults3`| Adults | Online (Mturk) | Wrong, Kinda Right, Right|
| Study 2 | `r n_study2` | `r ageInfo_study2[1]`;`r ageInfo_study2[2]`-`r ageInfo_study2[3]`;`r ageInfo_study2[4]` (M = `r ageInfo_study2[5]`;`r ageInfo_study2[6]`) | Study Room | Circle (wrong), Little Star (little right), Big Star (right) |
Table: (\#tab:study2info) Summary of Study 1 and Study 2 Methods

#### Materials and Design

We used the same set of cards and linguistic stimuli as the ones in study 1. There were 8 trial types and 2 trials per trial type for a total of 16 trials. We made two changes to make the experiment more suitable for children. First, instead of the fictional character Bob, a puppet named Jazzy played the guessing game with the children. Jazzy had a sleeping mask on his eyes during the game (Figure \@ref(fig:jazzy)). Second, a pilot study showed that a scale with three alternatives is better understood and used by children if it is presented in the form of rewards to the puppet rather than verbal responses such as "wrong", "a little bit right", and "right", or even hand gestures such as thumbs up, middle, and down. Therefore, we placed a set of red circles, small blue stars, and big blue stars in front of the children. These tokens were used to reward the puppet after each guess. During the introduction, the experimenter explained that if the puppet is right, the child should give him a big star, if he is a little bit right, a little star, and if he is not right, the puppet gets a red circle. 

```{r jazzy, fig.align='center', fig.env = 'figure' , fig.cap="Jazzy with and without the sleeping mask."}
jazzy_img <- png::readPNG("connective_comprehension/figs/Jazzy.png")
grid::grid.raster(jazzy_img)
```

#### Participants and Procedure

We recruited `r n_study2` English speaking children from the Bing Nursery School at Stanford University. Children were between `r ageInfo_study2[1]`;`r ageInfo_study2[2]` and `r ageInfo_study2[3]`;`r ageInfo_study2[4]` years old (Mean = `r ageInfo_study2[5]`;`r ageInfo_study2[6]`). The experiment was carried out in a quiet room and the sessions were videotaped. There was a small table and two chairs in the room. Children sat on one side of the table and the experimenter and the puppet on the other side facing the child. The groups of circles, small stars, and big stars were placed in front of the child from left to right. A deck of six cards was in front of the experimenter. Similar to study 1 with adults, participants sat through three phases: introduction, instruction, and test. 

The goal of the introduction was to show the cards to children and make sure they recognize the animals and know their names. The experimenter showed the cards to the children and asked them to label the animals. All children recognized the animals and could label them correctly. In the instruction phase, children went through three example trials. The experimenter explained that he is going to play with the puppet first so that the child can learn the game. He removed the six introduction cards and placed a deck of three cards face-down on the table. From top to bottom (first to last), the cards had the following images: cat, elephant, cat and dog. He put the sleeping mask on the puppet's eyes and explained that he is going to guess what animal is on the cards. He then picked the first card and asked the puppet: "*What do you think is on this card?*" The puppet replied with "*There is a dog*". The experimenter showed the cat-card to the child and explained that when the puppet is *not right* he gets a circle. He then asked the child to give the puppet a circle. Rewards were collected by the experimenter and placed under the table to not distract the child. The second trial followed the same pattern except that the puppet guessed *right* and the experimenter invited the child to give the puppet a big star. In the final trial, the puppet guessed that there is a cat on the card when the card had a cat and a dog on it. The experimenter said that the puppet was *a little right* and asked the child to give him a little star.

|Card | Guess | Reward |
|-----|-----|-----|
|CAT | there is a cat! | Circle |
|ELEPHANT | there is an elephant!| Big Star | 
|CAT-DOG | there is a dog! | Little Star |
Table: (\#tab:instruction) Instruction Trials.

In the test phase, the experimenter removed the three instruction cards and placed a deck of 16 randomized cards on the table. The experimenter explained that it was the child's turn to play with the puppet. The test phase followed the pattern described in the instruction phase. You can find the randomization code as well as the details of the methods on the online repository for this study. 

#### Offline Annotations {#feedbackCoding}

During analysis of the videos, children's linguistic feedback to the puppet after each guess was categorized into four types: 1. None, 2. Judgments, 3. Descriptions, and 4. Corrections. The first category referred to cases where children did not say anything and only rewarded the puppet. Judgments referred to linguistic feedback such as *you are right!*, *yes*, *nope*, or *you winned*. Such feedback only expressed judgments and complemented the rewards. Descriptions were cases that the child simply mentioned what was on the card: *cat!*, *dog and elephant!*, *There is a cat and a dog!* etc. Finally, corrections referred to feedback that provided additional linguistic elements that acted like corrections to what the puppet had said. Examples include: *Just a cat!*, *Both!*, *The two are!*, *Only cat*, *cat AND dog* (with emphasis placed on *and*). In trials where the child provided both judgments as well as descriptions or corrections, we placed the feedback into the more informative categories, namely description or correction.

```{r child_adult_data, fig.env="figure", fig.pos = "t", fig.align = "center", fig.width=3.5, fig.height=3.5, fig.cap = "Results for Adults' and Children's Conjunction and Disjunction Trials. Error bars represent 95\\% confidence intervals.", eval=FALSE }
# graph_data %>%
#   separate(trial_type, into = c("animal", "connective"), sep="\\.") %>%
#   ggplot(aes(x=response, y=prop,fill=age_category)) + 
# geom_bar(stat="identity", width = 0.6, position="dodge") + 
# geom_errorbar(aes(ymax = cih, ymin= cil), position = position_dodge(.6), width=0.2) + facet_grid(connective~animal) + 
#  guides(fill=FALSE) + labs(title = "", x = "", y = "") + theme_few()
```

### Results 

```{r childrenTernaryPlot, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Children's ternary judgments in the connective guessing game."}
prop_table$response <- fct_recode(prop_table$response, `A Little Right`="Kinda Right")

child_plot_tvjt <-
  prop_table %>%
  filter(`Age Category`=="Children") %>%
  ggplot(aes(x=response, y=prop, fill=response)) + 
  geom_bar(stat="identity", width=0.6) + 
  facet_grid(Card~Guess) +
  geom_linerange(aes(ymin=cil, ymax=cih)) +
  labs(x="",y="") +
  theme_few() +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1), text = element_text(size=17)) + 
  scale_fill_manual(values = c("red4", "grey","springgreen3"), guide=FALSE)

child_plot_tvjt_g <- ggplot_gtable(ggplot_build(child_plot_tvjt))

strips <- grep("strip-r", child_plot_tvjt_g$layout$name)

new_grobs <- list(cat, cat_dog)

child_plot_tvjt_g <- with(child_plot_tvjt_g$layout[strips,],
          gtable_add_grob(child_plot_tvjt_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

child_plot_tvjt_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(child_plot_tvjt_g)
```

Figure \@ref(fig:childrenTernaryPlot) shows the results for children's 3AFC judgments. Starting from the left column, we see that if the mentioned animal was not on the card (e.g. elephant), children judged the guess "wrong". If the mentioned animal (e.g. cat) was the only animal on the card, children judged the guess "right". Here we are going to ignore the results for trial types in which the mentioned animal was one of the animals on the card. The reason is that such trials were used in the instruction phase to introduce the "little bit right" guesses, and the results are potentially biased by the instructions.

In conjunctive guesses (e.g. *cat and dog*), when only one of the mentioned animals was on the card, children judged the guess as "wrong" or "a little bit right". However, if both animals are on the card, they judged the conjunctive guess as "right". In disjunctive guesses (e.g. *cat or dog*), when only one of the mentioned animals was on the card, children considered the guess "right" or "kinda right". If both animals were on the card, the disjunctive guess was considered "right".

Comparing the conjunction and disjunction trials (last two columns of figure \@ref(fig:childrenTernaryPlot)), we see that overall, children distinguish between *and* and *or* guesses in cases where one animal is on the card but not when both are. Given that the one-animal conjunction trials and the one-animal disjunction trials differ in truth conditions, the difference in response patterns suggests that children at this age have a different semantic knowledge for *and* and *or*. The two-animal conjunction and two-animal disjunction trials do not differ in truth values, and the responses also show no difference.

In the one-animal and two-animal trials (figure \@ref(fig:childrenTernaryPlot) rows), children show different response patterns when the guess contains the conjunction word *and* but not when *or* is used. Since the truth values of one-animal and two-animal trials differ for conjunctive guesses but not disjunctive ones, the results suggest that children are sensitive to the semantic differences between *and* and *or*. The similarity of the disjunctive guesses in one-animal and two-animal trials can also be interpreted as a lack of pragmatic reasoning in children. \@ref(fig:childAdultComp) compares the results for children and adults' 3AFC judgments in the conjunction and disjunction trials. The major difference between adults and children's responses is visible in the disjunctive trials where both animals are on the card; in other words trials where both disjuncts are true. In the next section, we use statistical modeling to compare adults and children's three-alternative responses more systematically. 

```{r childAdultComp, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Comparison of Adults' and Children's ternary judgments."}
child_adult_comp <-
  prop_table %>%
  filter(Options==3, Guess!="Elephant", Guess!="Cat") %>%
  ggplot(aes(x=response, y=prop, fill=`Age Category`)) + 
  geom_bar(stat="identity", width=0.6, position = "dodge") + 
  facet_grid(Card~Guess) +
  geom_linerange(aes(ymin=cil, ymax=cih), position = position_dodge(width=0.6)) +
  labs(x="",y="") +
  theme_few() +
  scale_fill_manual(values = c("grey", "red4"))

child_adult_comp_g <- ggplot_gtable(ggplot_build(child_adult_comp))

strips <- grep("strip-r", child_adult_comp_g$layout$name)

new_grobs <- list(cat, cat_dog)

child_adult_comp_g <- with(child_adult_comp_g$layout[strips,],
          gtable_add_grob(child_adult_comp_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

child_adult_comp_g$widths[[7]] <- unit(2.5,"cm")

grid.draw(child_adult_comp_g)
```

#### Analysis and Statistical Modeling

```{r rstanModel}
adult_samples <- read_csv("connective_comprehension/study2/3_analysis/adult_samples")
child_samples <- read_csv("connective_comprehension/study2/3_analysis/child_samples")

adult_comps <- filter(adult_samples, 
                 Parameter %in% c("alpha","beta_x_or",
                                  "beta_x_and","beta_xy_or"))

child_comps <- filter(child_samples, 
                 Parameter %in% c("alpha","beta_x_or",
                                  "beta_x_and","beta_xy_or"))

acomp_tab <- adult_comps %>%
  spread(Parameter, value) %>%
  mutate(`b(AND,2)-b(AND,1)` = alpha - beta_x_and,
         `b(OR,1)-b(AND,1)` = beta_x_or - beta_x_and, 
         `b(OR,2)-b(OR,1)` = beta_xy_or - beta_x_or) %>%
  select(-alpha, -beta_x_and, -beta_x_or) %>%
  rename(`b(OR,2)`=beta_xy_or) %>%
  gather(Parameter, value, `b(AND,2)-b(AND,1)`, `b(OR,1)-b(AND,1)`, 
         `b(OR,2)-b(OR,1)`, `b(OR,2)`) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99)) 

ccomp_tab <- child_comps %>%
  spread(Parameter, value) %>%
  mutate(`b(AND,2)-b(AND,1)` = alpha - beta_x_and,
         `b(OR,1)-b(AND,1)` =  beta_x_or - beta_x_and, 
         `b(OR,2)-b(OR,1)` = beta_xy_or - beta_x_or) %>%
  select(-alpha, -beta_x_and, -beta_x_or) %>%
  rename(`b(OR,2)`=beta_xy_or) %>%
  gather(Parameter, value, `b(AND,2)-b(AND,1)`, `b(OR,1)-b(AND,1)`, 
         `b(OR,2)-b(OR,1)`, `b(OR,2)`) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99)) 

ccomp_tab[,"Age"] <- "child"
acomp_tab[,"Age"] <- "adult"
comp_tabs <- rbind(ccomp_tab, acomp_tab)
comp_tabs$Age <- factor(comp_tabs$Age)
```

```{r stanModelPlot, fig.env="figure", fig.pos = "t", fig.align = "center", fig.width=4.5, fig.height=2.5, fig.cap = "Coefficients capturing the relevant comparisons across conditions in ternary judgments in Study 1 and 2. Error bars represent 99\\% regions of highest posterior density."}

ggplot(comp_tabs, aes(y=mean,x=Parameter, color=Age)) + 
  geom_point(position = position_dodge(width = .1)) + 
  geom_hline(yintercept=0, linetype="dashed") + 
  geom_linerange(aes(ymin=hdp_lower, ymax=hdp_upper), 
                 position = position_dodge(width = .1)) + 
  labs(y="Difference in Parameter Means", x="Trial Type Contrasts") + 
  scale_colour_solarized() + 
  theme_few() + 
  coord_flip()
```

We used the R package *{rstan}* for Bayesian statistical modeling. We fit separate ordinal mixed-effects logistic models for children's and adults' judgments. The response variable had three ordered levels: *wrong*, *kinda right*, and *right*. The trial types *One-Animal-OR*, *Two-Animals-OR*, *One-Animal-AND* constituted the (dummy-coded) fixed effects of the model with *Two-Animals-AND* set as the intercept. The model also included by-subject random intercepts. The priors over trial types and the random intercepts were set to $\mathcal{N}(0,10)$. We also included parameters $C_1$ and $C_2$, the two cutpoints delimiting the logistic for 1) *wrong* and *kinda right* and 2) *kinda right* and *right* responses, drawn with the prior $\mathcal{N}(0,1)$.^[We used a tight prior in this case to decrease posterior correlations between cutpoints and intercept.] All four chains converged after 3000 samples (with a burn-in period of 1500 samples)

We make inferences based on the highest-posterior density (HPD) intervals for the coefficients estimated from each model. Because predictors are dummy-coded, we can examine contrasts of interest by computing the difference between coefficients for pairs of conditions we wish to contrast. Figure \@ref(fig:stanModelPlot) shows the contrasts of interest: *b(OR, 2)-b(OR,1)*  represents the difference between the estimated coefficients for the disjunction trials with two animal on the card and those with only one; *b(OR, 2)* represents the difference between the estimated coefficients for the conjunction trials with two animals and the disjunction trials with two animals; and so on.

Overall, adults' and children's estimated coefficients are similar in sign to one another, though adults are more extreme. Comparing the conjunction trials (*b(AND, 2)-b(AND,1)*), children and adults showed a strong preference for the cards with two animals rather than one. At the same time, given two animals on the card, children and adults showed a preference for *and* rather than *or* (*b(OR, 2)*). However, with only one animal on the card, children and adults preferred a disjunctive guess (*b(OR, 1)-b(AND,1)*). These results are compatible with the truth conditions of conjunction and disjunction.

The main difference between adults and children shows up in the contrast between the disjunctive trial types: two animals vs. only one (*b(OR, 2)-b(OR,1)*). On average, children rated disjunction trials with two animals higher than those with only one. Adults on the other hand showed the opposite pattern: they rated disjunction trials with two animals lower. This pattern is compatible with current accounts of pragmatic development that suggest a lack of implicatures in children's interpretations. The idea is that while adults strengthen the disjunctive guess "cat or dog" to "cat or dog but not both", children simply interpret it as "cat or dog or both". Therefore, adults are going to rate trials where both disjuncts are true lower. 

The slight preference children show for cards with two animals when the guess is disjunctive is also compatible with the account proposed by @Singh2016 and @tieu2016. However, the effect is much smaller here than what was reported in their studies. The comparison with conjunction trials makes it clear that overall, children are not interpreting *or* as a conjunction. The effect in this study can be more accurately described as a preference for both disjuncts being true rather than a conjunctive interpretation of disjunction. We will further discuss the nature of this preference in the general discussion section.

#### Children's open-ended feedback

```{r chisq-tests}
criticals <- 
  data_main_dv3 %>% filter(!is.na(response),  age_category== "Children", 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY"))
criticals$trial_type <- factor(criticals$trial_type)

ctable <-
  criticals %>% select(feedback, trial_type) %>%
  group_by(trial_type, feedback) %>%
  summarise (counts = n()) %>%
  spread(trial_type, counts)
ctable <- as.data.frame(ctable)

chisq_1andor <- chisq.test(ctable[,2], p= ctable[,3], rescale.p = T)
chisq_and12 <- chisq.test(ctable[,2], p = ctable[,4], rescale.p = T)
chisq_1and2or <- chisq.test(ctable[,2], p = ctable[,5], rescale.p = T)
chisq_1or2and <- chisq.test(ctable[,3], p= ctable[,4], rescale.p = T)
chisq_or12 <- chisq.test(ctable[,"X.XorY"], 
                         ctable[,"XY.XorY"], 
                         rescale.p = TRUE)
chisq_2andor <- chisq.test(ctable[,"XY.XandY"], 
                           p=ctable[, "XY.XorY"], rescale.p = T)

```

|Category | Definition | Examples |
|---|----------------------------------|-------------------------------|
|None | no feedback provided to the puppet, only reward | | 
|Judgment | the child said yes/no, you are right, etc. | "No!" , "You are right Jazzy!" | 
|Description | mentioned the animal(s) on the card | "elephant", "cat and dog" |
|Correction | used focus particles like "only"/"just", emphasized "and" or used "both" | "only cat", "just elephant", "Both!", "cat AND dog!"|

Table: (\#tab:feedbackCat) Definitions and Examples for the Feedback Categories.

As explained in section \@ref(feedbackCoding), we also categorized and annotated children's spontaneous and free form verbal reactions to the puppet's guesses. Table \@ref(tab:feedbackCat) summarizes the definitions and examples for each category and figure \@ref(fig:feedbackData) shows the results.

```{r feedbackData, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4,   fig.cap = "Children's open-ended Feedback.Error bars represent 95\\% confidence intervals." }

study2_feedback <-
  data_children %>%
  group_by(sid, trial_type, feedback) %>%
  summarise(counts= n()) %>%
  spread(feedback, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = None + Judgement + Description + Correction) %>%
  gather(feedback, counts, Correction:None) %>%
  mutate(prop = counts / total) %>%
  group_by(trial_type, feedback) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop)) %>%
  separate(trial_type, into = c("animal", "connective"), sep="\\.")

study2_feedback$feedback <- factor(study2_feedback$feedback, levels = c("None", "Judgement", "Description", "Correction"))

study2_feedback$connective <- 
  study2_feedback$connective %>%
  factor() %>%
  fct_relevel("Z") %>%
  fct_recode(Cat = "X", Elephant = "Z", `Cat or Dog`="XorY", `Cat and Dog`="XandY")
  
study2_feedbackPlot <-
  ggplot(study2_feedback, 
       aes(x=feedback, y=prop, fill=feedback)) + 
  geom_bar(stat="identity", width = 0.6, position="dodge") + 
  geom_linerange(aes(ymax = cih, ymin = cil)) + 
  facet_grid(animal~connective) + 
  labs(title = "", x = "", y = "") + 
  theme_few() + 
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = .5)) + 
#  scale_fill_solarized(guide=FALSE)
  scale_fill_manual(values = c("gray", "red2", "red3", "red4"), guide=FALSE)

study2_feedbackPlot_g <- ggplot_gtable(ggplot_build(study2_feedbackPlot))

strips <- grep("strip-r", study2_feedbackPlot_g$layout$name)

cat <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/cat_card.jpg") %>%
  rasterGrob(width=0.9)

cat_dog <- 
  readJPEG("connective_comprehension/study1/0_methods/stimuli/catdog_card.jpg")%>%
  rasterGrob(width=0.9)

new_grobs <- list(cat, cat_dog)

study2_feedbackPlot_g <- with(study2_feedbackPlot_g$layout[strips,],
          gtable_add_grob(study2_feedbackPlot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

study2_feedbackPlot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(study2_feedbackPlot_g)
```

We should first point out that each trial type shows similar number of "None" feedback. The reason is that some children remained more or less silent throughout the experiment and only provided rewards to the puppet. After all the main measurement of this experiment was children's forced-choice rewards and the analysis of their open-ended feedback was secondary. In the next study we focus on children's open-ended feedback. In our discussion and analysis here we do not comment on the "None" category anymore and focus on the other three categories.

Starting with the leftmost column, we see that when the guessed animal (i.e. elephant) is not on the card, children either provide judgments like "No!" or describe what is on the card like "cat" or "cat and dog". However, when the guessed animal is the only animal on the card (i.e. cat), most children provide a positive judgment like "Yes". When the guessed animal is only one of the animals on the card, children describe what is on the card like "cat and dog". Corrections were rare for all these four control trial types.

Let's move to critical trial types with conjunction and disjunction. Children showed a high rate of corrections and description when the guess used "and" but only one of the animals was on the card. In their corrections, children used the focus particles "just" and "only" like "just a cat" or "only a cat". However, in trial types where conjunction was used and both animals were on the card, children predominantly provided positive judgments like "Yes!" and "You are right". Considering disjunctive guesses like "cat or dog", when only one of the animals was on the card, the vast majority of children simply described what was on the card, for example "cat". However, when both animals were on the card, children corrected the puppet by saying "both" or emphasizing "and" like "cat AND dog". 

We performed chi-squared goodness-of-fit tests to compare the feedback distributions in our critical conditions with "and" and "or". Here we focus on those trials (the four bar charts on the right of figure \@ref(fig:feedbackData)). Children's linguistic feedback fell into three patterns. First, the one-animal conjunctive and two-animal disjunctive (top left and bottom right) trials show higher proportion of *Corrections* than the other trial types. These are trials that the guesses are either false or infelicitous. In the conjunction trials, a comparison of the feedback distribution in one-animal and two-animal conditions was statistically significant ($\chi$(`r chisq_and12$parameter`, `r sum(chisq_and12$observed)`) = `r  round(chisq_and12$statistic,2)`, p < .0001), indicating different feedback for false sentences. A similar numerical trend was present in the disjunction trials, but was not significant ($\chi$(`r chisq_or12$parameter`, `r sum(chisq_or12$observed)`) = `r round(chisq_or12$statistic,2)`, p = `r round(chisq_or12$p.value,2)`). 

Second, the one-animal disjunctive trials (top right) showed the highest proportion of *Descriptions*. These are trials in which the guess is correct but not specific enough: it leaves two possibilities open. These trials were significantly different than the one-animal trials for conjunction ($\chi$(`r chisq_1andor$parameter`, `r sum(chisq_1andor$observed)`) = `r round(chisq_1andor$statistic,2)`, p < .0001). Finally, the two-animal conjunctive trials (bottom left) showed the highest proportion of *Judgments* such as *you are right!*. This is not surprising given that in these trials represent the most optimal guessing scenario. These trials had a significantly different feedback distribution from the matching disjunction trials ($\chi$(`r chisq_2andor$parameter`, `r sum(chisq_2andor$observed)`) = `r round(chisq_2andor$statistic,2)`, p < .0001).

### Discussion

In study 2, we used a 3AFC judgment task to test children's comprehension of logical connectives *and* and *or*. We compared these results to those found in the 3AFC judgment task of study 1 with adults. The general comparison showed that adults and children have similar patterns of judgments, except when both disjuncts are true. In such cases, adults judge the disjunctive guess as not completely right while most children find it completely right. There was even a slight preference among children to rewarded the puppet more in such cases, compared to cases of disjunction when only one disjunct is true.

To consider another measure of children's comprehension, we also looked at children’s spontaneous open-ended feedback to the guesses. Our analysis suggested that children recognize false and infelicitous utterances with the connectives and provide appropriate corrective feedback. As expected from an adult-like understanding of connectives, children corrected the puppet most when there was only one animal on the card and the guess was conjunctive, or when there were two animals on the card and the guess was disjunctive.  Perhaps the most important finding was that children increased their corrective feedback in disjunctive guesses where both disjuncts were true, compared to those with only one disjunct true. This is contrary to the the results of the 3AFC judgment task which suggested that children do not find any infelicity with disjunctive guesses when both disjuncts are true.

The analysis of children's open-ended feedback raises two important issues. First, as mentioned before, it runs counter to what the 3AFC judgment task suggests with respect to cases of exclusivity implicatures (i.e. trials with disjunction when both disjuncts are true). The forced-choice judgment task suggests that children find such cases as unproblematic while the analysis of their spontaneous feedback shows that they provide more corrections to the puppet. Second, a common explanation for why children fail at deriving implicatures is that they cannot access the stronger alternative to the disjunction word *or*, namely *and* [@barner2011accessing]. However, in the context of our guessing game, some children explicitly mentioned the word *and* as what the puppet should have said instead of *or*. Interestingly, these children continued to reward the puppet and considered the guess "right". This raises the possibility that children's forced-choice truth value judgments, whether with two or three alternatives, do not properly reflect their knowledge of pragmatics. In study 3, we decided to make children's open-ended feedback the main dependent variable of study and directly compare them to a 2AFC truth judgment in the same task. This design allowed us to directly compare children's open-ended and forced choice responses. 

## Study 3: Children's 2AFC judgments and open-ended feedback

```{r study3Wrangling}
data_study3 <- read.csv("connective_comprehension/study3/2_processed_data/study2_data")

#filter(feedback!="0", tvjt!="YN", !is.na(tvjt)) # Exclusions

tvjt_exclusion <-
  data_study3 %>%
  filter(tvjt=="YN" | is.na(tvjt))

feedback_exclusion <-
  data_study3 %>%
  filter(feedback=="0")

data_study3$Guess <- 
  data_study3$Guess %>%
  factor(levels = c("Z", "X", "XandY", "XorY")) %>%
  fct_recode(Elephant = "Z", Cat = "X", `Cat and Dog`="XandY", `Cat or Dog`="XorY")

data_study3$tvjt <- 
  data_study3$tvjt %>%
  factor(levels = c("N", "Y")) %>%
  fct_recode(Yes = "Y", No = "N")

n_study3 <- length(unique(data_study3$id))

ageInfo_study3 <- c(floor(min(data_study3$age*0.00273973)), # Min age in year
                    round((min(data_study3$age*0.00273973)-floor(min(data_study3$age*0.00273973)))*12), #Min age months
                    floor(max(data_study3$age*0.00273973)), # Max age year
                    round((max(data_study3$age*0.00273973)-floor(max(data_study3$age*0.00273973)))*12), # Max age months
                    floor(mean(data_study3$age*0.00273973)), # Mean age year
                    round((mean(data_study3$age*0.00273973)-floor(mean(data_study3$age*0.00273973)))*12)) # Mean age months
```

This study used the same paradigm as study 2 but focused on children's open-ended feedback and aimed at replicating the findings in study 2. The main hypothesis was that four-year-olds provide corrective feedback to the puppet if both disjuncts are true, but they do not consider this infelicity to be grave enough to render the guess "wrong". The main hypothesis along with relevant analyses and predictions were preregistered in an "As Predicted" format[^1]. The study used a 2AFC judgment task to compare with the open-ended feedback results. We expected children to provide corrective feedback to the puppet when both disjuncts were true, yet consider the guess "right" and not reflect this infelicity in truth value judgments. This is what the study found.

[^1]: The As Predicted pdf document is accessible at https://aspredicted.org/x9ez2.pdf.

### Methods
|Study | N | Age | Mode | DV |
|-------------|---|----------------|------------|---------------------------|
|Study 1 - Part 1 | `r n_adults2`| Adults | Online (Mturk) | Wrong, Right|
|Study 1 - Part 2 | `r n_adults3`| Adults | Online (Mturk) | Wrong, Kinda Right, Right|
|Study 2 | `r n_study2` | `r ageInfo_study2[1]`;`r ageInfo_study2[2]`-`r ageInfo_study2[3]`;`r ageInfo_study2[4]` (M = `r ageInfo_study2[5]`;`r ageInfo_study2[6]`) | Study Room | Circle (Wrong), Little Star (Little Right), Big Star (Right) |
|Study 3 | `r n_study3` | `r ageInfo_study3[1]`;`r ageInfo_study3[2]`-`r ageInfo_study3[3]`;`r ageInfo_study3[4]` (M = `r ageInfo_study3[5]`;`r ageInfo_study3[6]`) | Study Room | Yes (Right)/No (Wrong) - Open-ended Feedback |
Table: (\#tab:study3info) Summary of Study 1, 2, and 3 Methods

#### Materials and Design

Study 3 was similar to study 2 but differed on how children provided their judgments. Based on our findings in study 2, we decided to focus on verbal judgments and feedback, instead of rewards. Two different ways of measuring children's judgments were used. First, we encourages children to provide verbal feedback to the puppet. Children were asked to say "yes" when the puppet was right and "no" when he was not. They were also encouraged to help the puppet say it better when he was not right. After children were done with this initial open-ended feedback to the puppet, we asked them a forced choice yes/no judgment question: "Was Jazzy (the puppet) right?". This question elicited a "yes" or "no" response for each trial independent from the earlier open-ended response. Using both measures enabled us to compare open-ended and forced choice judgments.

#### Participants and Procedure

We recruited `r n_study3` English speaking children from the Bing Nursery School at Stanford University. Children were between `r ageInfo_study3[1]`;`r ageInfo_study3[2]` and `r ageInfo_study3[3]`;`r ageInfo_study3[4]` years old (Mean = `r ageInfo_study3[5]`;`r ageInfo_study3[6]`). The setup and procedure was similar to study 2, except there were no rewards on the table. As before, participants sat through three phases: introduction, instruction, and test. The introduction phase made sure children knew the names of the animals on the cards. In the instruction phase, there were four training trials shown in table \@ref(tab:instructionStudy2).

Similar to study 2, the experimenter put the sleeping mask on the puppet's eyes and explained that Jazzy (the puppet) is going to guess what animal is on the cards. He then picked the first card and asked the puppet: "*What do you think is on this card?*" The puppet replied with "*There is a dog*". The experimenter showed the cat-card to the child and said: when Jazzy is *not right*, tell him "No". He then asked the child to say "no" to the puppet. The second trial followed the same pattern except that the puppet guessed *right* and the experimenter invited the child to say "yes" to the puppet. There were two more instruction trials and then the test phase began. It contained 16 randomized trials, half of which contained guesses with the words "and" and "or". You can find the randomization code as well as the details of the methods on the online repository for this study.

|Card | Guess | Response |
|-----|-----|-----|
|CAT | there is a dog! | No! |
|ELEPHANT | there is an elephant!| Yes! | 
|DOG-ELEPHANT | there is a cat! | No! |
|DOG | there is a dog! | Yes! |
Table: (\#tab:instructionStudy3) Instruction Trials for Study 3.

### Results

```{r study3JudgementTable}
study3_tvjt <-
  data_study3 %>%
  filter(tvjt!="YN", !is.na(tvjt)) %>% #exclusion
  group_by(id, Animal, critical, Guess, tvjt) %>%
  summarise(counts= n()) %>%
  spread(tvjt, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = No + Yes) %>%
  gather(tvjt, counts, No:Yes) %>%
  mutate(prop = counts / total) %>%
  group_by(Animal, Guess, tvjt, critical) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))
```

Here we first take a look at the results of the 2AFC judgement task for each trial type and compare them to those of the adults' in study 1. Then we investigate children's open-ended responses and compare them to the forced choice responses obtained in the same trials.

For the 2AFC judgments we excluded `r nrow(tvjt_exclusion)` trials (out of total `r nrow(data_study3)`) where children either did not provide a Yes/No response or provided both (i.e. "Yes and No"). The exclusions were almost equally distributed among different types of guesses and cards. In our analysis of children's open-ended feedback, we excluded `r nrow(feedback_exclusion)` trials (out of total `r nrow(data_study3)`) where children either did not provide any feedback or their feedback could not be categorized into our existing categories.

#### Two-Alternative Forced Choice Judgments

```{r Study3tvjtPlot, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Children's binary truth value judgments."}
tvjt_plot <-
  study3_tvjt %>%
  ggplot(aes(x=tvjt, y=prop, fill=tvjt)) + 
  geom_bar(stat="identity", width = 0.6, position="dodge") + 
  geom_linerange(aes(ymax = cih, ymin = cil)) + 
  facet_grid(Animal~Guess) + 
  guides(fill=FALSE) + 
  labs(title = "", x = "", y = "") + 
  theme_few() + 
  theme(text = element_text(size=15)) + 
  scale_fill_manual(values = c("red3", "springgreen3"))

tvjt_plot_g <- ggplot_gtable(ggplot_build(tvjt_plot))

strips <- grep("strip-r", tvjt_plot_g$layout$name)

new_grobs <- list(cat, cat_dog)

tvjt_plot_g <- with(tvjt_plot_g$layout[strips,],
          gtable_add_grob(tvjt_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

tvjt_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(tvjt_plot_g)
```

```{r binaryStats}
data_study3$Guess <- 
  data_study3$Guess %>%
  fct_recode(Z= "Elephant", X = "Cat", XandY = "Cat and Dog", XorY = "Cat or Dog")

binaryStats_children <-
  data_study3 %>%
  filter(tvjt!="YN", !is.na(tvjt)) %>%
  select(Animal, Guess, tvjt, id) %>%
  rename(response="tvjt", sid="id", Card = "Animal")

binaryStats_children[,"age_category"] <- "Children"

binaryStats_children$response <- recode(binaryStats_children$response, Yes="Right", No="Wrong") # change yes no to right and wrong

binaryStats_adults <-
  filter(data_adults, dv_type==2) %>%
  select(sid, response, age_category, Card, Guess)

binaryStats <- bind_rows(binaryStats_adults, binaryStats_children)
binaryStats$response <- factor(binaryStats$response, levels=c("Wrong", "Right"))

oneDisjunct_model <- summary(glmer(response ~ age_category + (1|sid), family= "binomial", data=filter(binaryStats, Card=="X", Guess=="XorY")))

twoDisjunct_model <- summary(glmer(response ~ age_category + (1|sid), family= "binomial", data=filter(binaryStats, Card=="X", Guess=="XandY")))

childDisj_model <- summary(glmer(response ~ Card + (1|sid), family= "binomial", data=filter(binaryStats, age_category=="Children", Guess=="XorY")))
```

Figure \@ref(fig:Study3tvjtPlot) shows children's 2AFC judgments. Starting from the leftmost column, when the guessed animal (e.g. elephant) was not on the card, children considered the guess "wrong". When the guessed animal (e.g. cat) was the only animal on the card, children considered the guess "right". However, if the guessed animal (e.g. cat) was only one of the animals on the card, children were equally split between "wrong" and "right" judgments. On the other hand, almost all adults considered such guesses "right" in their 2AFC judgments (Figure \@ref(fig:binaryAdultsPlot)). In such trial types, children seem to interpret the guess "there is a cat" as "there is **only** a cat", while adults do not. This difference between children and adults is unexpected for a theory of meaning acquisition that assumes children are overall more logical/literal interpreters than adults. 

Moving to the trials with *and* and *or*, we see that children's judgments are similar to those of adults'. You can see figure \@ref(fig:BinaryPlotComp) for the direct comparison of adults' and children's 2AFC judgments. In trials with conjunction, when only one of the animals was on the card, most children considered the guess "wrong". This is similar to adults and only different in extent: adults more consistently and unanimously rejected such guesses than children. A mixed effects logistic regression with the fixed effect of age category (adult vs. child) and random effect of subject found no significant difference between adults' and children's responses in such trials (Table \@ref(tab:statsStudy3), Conjunction - One Animal). 

|Trial Data | Coefficient | Standard Error | Z-Value | P-value |
|-----|-----|-----|----|-----|
|Conjunction - One Animal | `r round(oneDisjunct_model$coefficients[2],2)` | `r round(oneDisjunct_model$coefficients[4],2)` | `r round(oneDisjunct_model$coefficients[6],2)` | `r round(oneDisjunct_model$coefficients[8],2)` |
|Disjunction - One Animal | `r round(twoDisjunct_model$coefficients[2],2)` | `r round(twoDisjunct_model$coefficients[4],2)` | `r round(twoDisjunct_model$coefficients[6],2)` | `r round(twoDisjunct_model$coefficients[8],2)` |

Table: (\#tab:statsStudy3) Mixed effects logistic models for disjunction trials in 2AFC judgments of adults and children, using `glmer` in R's \{LME4\} package. Formula: $Response \sim Age Category + (1|Subject)$.

In conjunctive guesses where both animals were on the card, both children and adults considered the guess "right" unanimously. In disjunctive trials when only one of the animals was on the card, most children considered the guess "right". This is again similar to adults but differs from them in extent: adults more consistently and unanimously judged such guesses as "right". Yet, a mixed effects logistic regression with the fixed effect of age category (adult vs. child) and random effect of subject found no significant difference between adults' and children's responses in such trials (Table \@ref(tab:statsStudy3), Disjunction - One Animal). Adults and children showed almost identical patterns of judgments in trials where there was two animals on the card and the guess used the connective "or". Children and adults did not differ in their rate of rejecting disjunctive guesses when both disjuncts were true.

```{r BinaryPlotComp, fig.env="figure", fig.pos="t", fig.align='center', fig.width=8, fig.height=4, fig.cap="The comparision of the 2AFC judgment task for conjunction and disjunction trials in adults (study 1) and children (study 3)."}

child_binary <-
  study3_tvjt %>% 
  rename(Card="Animal", response="tvjt") %>%
  select(-critical) %>%
  filter(Guess!="Elephant", Guess!="Cat")

child_binary[,"Age Category"] <- "Children" # add age category 
child_binary$response <- recode(child_binary$response, Yes="Right", No="Wrong") # change yes no to right and wrong

adult_binary <-
  prop_table %>%
  filter(Options==2, response!="Kinda Right", `Age Category`=="Adults", Guess!="Cat", Guess!="Elephant") %>% ungroup() %>% select(-Options)

adult_binary$response <- fct_drop(adult_binary$response)

binary_comp <-
  bind_rows(adult_binary, child_binary)

binary_comp$response <- fct_relevel(binary_comp$response, "Wrong")

binary_compPlot <-
  binary_comp %>%
  ggplot(aes(x=response, y=prop, fill=`Age Category`)) + 
  geom_bar(stat = "identity", position = "dodge", width=0.6) + 
  geom_linerange(aes(ymin=cil,ymax=cih), position = position_dodge(width=0.6)) +
  facet_grid(Card~Guess) + 
  labs(x="",y="") +
  theme_few() +
  scale_fill_manual(values = c("grey", "red4"))

binary_compPlot_g <- ggplot_gtable(ggplot_build(binary_compPlot))

strips <- grep("strip-r", binary_compPlot_g$layout$name)

new_grobs <- list(cat, cat_dog)

binary_compPlot_g <- with(binary_compPlot_g$layout[strips,],
          gtable_add_grob(binary_compPlot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

binary_compPlot_g$widths[[7]] <- unit(2.5,"cm")

grid.draw(binary_compPlot_g)
```

Finally, there is a small but significant preference in children's judgments of disjunctive statements for both disjuncts to be true. Comparing the disjunctive trials with one animal and two animals on the card, a mixed-effects logistic model with the fixed effect of disjunction type and the random effect of subjects found that children have a slight preference for both animals to be on the card ($b$= `r round(childDisj_model$coefficients[2],2)`, $se$= `r round(childDisj_model$coefficients[4],2)`, $z$= `r round(childDisj_model$coefficients[6],2)`, $p < 0.001$ ). We found a similar small trend in children's three-alternative judgments in study 2. While this preference is quite small compared to the other effects observed in these studies, it nevertheless shows a minor difference between children and adults judgments and we will discuss it in more detail in section \@ref(conjunctive) of our general discussion.

#### Open-ended Feedback

```{r feedbackStudy3, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4,   fig.cap = "Children's Open-ended Feedback.Error bars represent 95\\% confidence intervals." }

study3_feedback <-
  data_study3 %>%
  filter(feedback!="0") %>%
  group_by(id, Animal, Guess, feedback) %>%
  summarise(counts= n()) %>%
  spread(feedback, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = J + D + C) %>%
  gather(feedback, counts, C:J) %>%
  mutate(prop = counts / total) %>%
  group_by(Animal, Guess, feedback) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

study3_feedback$feedback <- 
  study3_feedback$feedback %>%
  factor() %>%
  fct_recode(Judgment = "J", Correction = "C", Description = "D") %>%
  fct_relevel(c("Judgment", "Description","Correction"))

study3_feedback$Guess <- 
  study3_feedback$Guess %>%
  factor() %>%
  fct_relevel("Z") %>%
  fct_recode(Cat = "X", Elephant = "Z", `Cat or Dog`="XorY", `Cat and Dog`="XandY")
  
study3_feedbackPlot <-
  ggplot(study3_feedback, 
       aes(x=feedback, y=prop, fill=feedback)) + 
  geom_bar(stat="identity", width = 0.6, position="dodge") + 
  geom_linerange(aes(ymax = cih, ymin = cil)) + 
  facet_grid(Animal~Guess) + 
  labs(title = "", x = "", y = "") + 
  theme_few() + 
  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = .5)) + 
#  scale_fill_solarized(guide=FALSE)
  scale_fill_manual(values = c("red2", "red3", "red4"), guide=FALSE)

study3_feedbackPlot_g <- ggplot_gtable(ggplot_build(study3_feedbackPlot))

strips <- grep("strip-r", study3_feedbackPlot_g$layout$name)

new_grobs <- list(cat, cat_dog)

binary_compPlot_g <- with(study3_feedbackPlot_g$layout[strips,],
          gtable_add_grob(study3_feedbackPlot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

study3_feedbackPlot_g$widths[[7]] <- unit(2.5,"cm")

grid.draw(study3_feedbackPlot_g)
```

Figure \@ref(fig:feedbackStudy3) shows the distribution of children's feedback to the puppet in study 3. You can refer to table \@ref(tab:feedbackCat) to refresh your memory on the definitions and examples of feedback categories. There were no "None" responses in this study since the study explicitly asked children to provide feedback to the puppet. The distribution of the responses in the other three categories (Judgment, Description, and Correction) shows a successful replication of study 2. 

Children's feedback distribution shows four main patterns. First when the puppet guessed an animal not on the card (e.g. elephant), we see a split pattern between negative judgments like "no!" or just descriptions of what was on the card like "cat!". Children provided no corrections in such trials. Second, almost all children responded with positive judgments like "yes" when the puppet's guess accurately matched what was on the card. This was the case in trials where there was only one animal on the card (e.g. cat) and the puppet mentioned it, and trials where there were two animals and the puppet mentioned both with a conjunction (e.g. cat and dog). Third, children provided the highest number of corrective feedback in trials where the guess was either false or infelicitous. These included three trial types: the ones where there were two animals on the card but the puppet only guessed one (e.g. cat); the ones where the puppet guessed two animals with conjunction (e.g. cat and dog) but only one of them was on the card; and the ones where there were two animals and the puppet guessed both but used a disjunction (e.g. cat or dog). The fourth general pattern was unique to disjunctive trials with only one animal on the card. In such trials, almost all children simply named the animal on the card (e.g. "Cat!").

```{r judgmentProptable}
judgment_data <- 
  data_study3 %>%
  filter(judgment!="YN") %>% # excluding the child that said yes and no.
  group_by(id, Animal, Guess, judgment) %>%
  summarise(counts= n()) %>%
  spread(judgment, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = `0` + N + Y) %>%
  gather(judgment, counts, `0`:Y) %>%
  mutate(prop = counts / total) %>%
  group_by(Animal, Guess, judgment) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

judgment_data$judgment <- fct_recode(judgment_data$judgment, No="N", `-`="0", Yes="Y")
judgment_data$judgment <- fct_relevel(judgment_data$judgment, "No", "-", "Yes")
```

```{r study3JudgmentPlot, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=8, fig.height=4, set.cap.width=T, num.cols.cap=1, fig.cap = "Children's open-ended feedback to the puppet's guesses."}
judgment_plot <-
  judgment_data %>%
  ggplot(aes(x=judgment, y=prop, fill=judgment)) + 
  geom_bar(stat = "identity") + 
  geom_linerange(aes(ymax=cih,ymin=cil)) +
  facet_grid(Animal~Guess) + 
  theme_few() + labs(x="", y="") + scale_fill_manual(values= c("red3", "grey", "springgreen3")) + guides(fill=FALSE)

judgment_plot_g <- ggplot_gtable(ggplot_build(judgment_plot))

strips <- grep("strip-r", judgment_plot_g$layout$name)

new_grobs <- list(cat, cat_dog)

judgment_plot_g <- with(judgment_plot_g$layout[strips,],
          gtable_add_grob(judgment_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

judgment_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(judgment_plot_g)
```

Figure \@ref(fig:study3JudgmentPlot) breaks down children's open-ended feedback based on whether children said "yes", "no", or said something else. Responses that were not yes/no judgments are grouped in a middle category shown with a dash. The goal here is to compare children's open-ended judgments with their forced choice judgments shown in figure \@ref(fig:Study3tvjtPlot). Children's open-ended judgments and their forced choice judgments in study 3 show similar patterns for all types of guesses except for one: disjunctive ones. In trials that the puppet guessed using "or", the vast majority of children refused to provide a yes/no judgment when they were not forced to. Instead, they described the animal on the card or provided corrections to the puppet's guess if the disjunction was infelicitous. 

One way to interpret these results is that disjunctive guesses (with at least one disjunct true) are considered neither right nor wrong by almost all children. When children are forced to provide wrong-right responses in the experimental context, some may conform to the adult patterns of judgment and some may not. However, it is possible that such deviations from adult judgments do not reflect differences in the comprehension of disjunction, but rather differences in how children map their adult-like comprehension onto the notions of "right" and "wrong" in a forced choice judgment task.

```{r feedbackFillData}
feedback_data <- 
  data_study3 %>%
  filter(feedback!="0")

feedback_data$Guess <- 
  feedback_data$Guess %>%
  factor() %>%
  fct_relevel("Z") %>%
  fct_recode(Cat = "X", Elephant = "Z", `Cat or Dog`="XorY", `Cat and Dog`="XandY")
feedback_data <- unite(feedback_data, Feedback, c(feedback, correction), sep="")

feedback_data$judgment <- fct_relevel(feedback_data$judgment, "N", "0", "Y", "YN")
feedback_data$judgment <- fct_recode(feedback_data$judgment, No="N", `-`="0", Yes="Y")

feedback_data$Feedback <- fct_relevel(feedback_data$Feedback, "J0", "D0", "CE" , "CI")
feedback_data$Feedback <- fct_recode(feedback_data$Feedback, `Yes/No` = "J0", `Just/Only Cat`="CE", `Both/Cat AND Dog`="CI", `Cat (and Dog)`="D0")
```

```{r feedbackFillPlot, fig.env="figure", fig.pos="t", fig.align='center', fig.width=8, fig.height=4, fig.cap="Children's open-ended feedback. The x-axis shows whether children said yes or no and the color fill shows what children said other than yes/no."}
feedback_plot <-
  feedback_data %>%
  filter(judgment!="YN") %>%
  ggplot(aes(x=judgment, fill=Feedback, order=Feedback)) + geom_bar() + facet_grid(Animal~Guess) + theme_few() + labs(x="", y="") + scale_fill_manual(values= c("grey", "gold", "blue3", "red3"))

feedback_plot_g <- ggplot_gtable(ggplot_build(feedback_plot))

strips <- grep("strip-r", feedback_plot_g$layout$name)

new_grobs <- list(cat, cat_dog)

feedback_plot_g <- with(feedback_plot_g$layout[strips,],
          gtable_add_grob(feedback_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

feedback_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(feedback_plot_g)
```

Figure \@ref(fig:feedbackFillPlot) is similar to figure \@ref(fig:study3JudgmentPlot) but it uses color-fill to show what children said in addition to yes or no. The gray color represents the trials where children only said yes/no and nothing else. The yellow-fill represents descriptions where children mentioned the animal on the card (i.e. "cat!"). The blue fill represents children's corrective feedback that used the exclusive focus particles  *just* or *only* (i.e. "just a cat!"). Such a corrective feedback suggests that the guess included an animal that needed to be excluded. Finally the red fill represents the inclusive corrective feedback that emphasized the word *and* or said *both* (i.e. "both", "cat AND dog"). Such corrective feedback indicates that two animals need to be included.

Starting from the leftmost column, when the puppet mentioned an animal that was not on the card (e.g. elephant), children responded with a simple "no" or "no" followed by what **was** on the card (e.g. no! elephant!). When there was only one animal on the card and the puppet mentioned the animal (e.g. cat), children responded with a simple "yes". However, when the card had two animals (e.g. cat and dog) and the puppet only mentioned one of them (e.g. cat), children were likely to provide inclusive feedback: say *both* or emphasize "cat *AND* dog". However, in such trials children were equally split between saying yes, no, or neither.

Moving to the trials with conjunctive and disjunctive guesses, when there was only one animal on the card (e.g. cat) and the puppet used a conjunction (e.g. cat and dog), children were likely to say a simple "no" or say "no" and follow it with "only/just" (e.g. no, just a cat). Some children did not mention "no" but still responded with "only/just". When the card had two animals on it and the puppet mentioned both using *and*, children responded with a simple "yes". In trials with disjunctive guesses like "cat or dog", children avoided yes/no responses. Instead, when the card had only one animal, children mentioned the animal on the card. When the card had both animals, children said *both* or emphasized the word *AND* like "cat *AND* dog". 

```{r inclusionTable}
inclusion_table <- 
  feedback_data %>%
  filter(judgment!="YN") %>%
  group_by(id, Animal, Guess, Feedback) %>%
  summarise(counts= n()) %>%
  spread(Feedback, counts) %>%
  replace(is.na(.), 0) %>%
  mutate(total = `Yes/No` + `Cat (and Dog)` + `Just/Only Cat` + `Both/Cat AND Dog`) %>%
  gather(Feedback, counts, `Yes/No`:`Both/Cat AND Dog`) %>%
  mutate(prop = counts / total) %>%
  group_by(Animal, Guess, Feedback) %>%
  summarize(cih = ci.high(prop),
            cil = ci.low(prop),
            prop = mean(prop))

inclusion_table$Feedback <- fct_relevel(inclusion_table$Feedback, "Yes/No", "Cat (and Dog)", "Just/Only Cat", "Both/Cat AND Dog")
```

```{r correctivePlot, fig.env="figure", fig.pos="t", fig.align='center', fig.width=8, fig.height=4, fig.cap="Children's feedback categories in disjunction trials."}
inclusion_plot <-
  inclusion_table %>%
#  filter(Guess == "Cat or Dog") %>%
  ggplot(aes(x=Feedback, y=prop, fill=Feedback)) + geom_bar(stat="identity") +
  geom_linerange(aes(ymin=cil, ymax=cih)) + facet_grid(Animal~Guess) + labs(x="", y="") + theme_few() + scale_fill_manual(values=c("grey", "gold", "blue3" ,"red3")) + theme (text = element_text(size=15), axis.text.x = element_text(angle=30, hjust = 1, vjust = 1)) + guides(fill=FALSE)

inclusion_plot_g <- ggplot_gtable(ggplot_build(inclusion_plot))

strips <- grep("strip-r", inclusion_plot_g$layout$name)

inclusion_plot_g <- with(inclusion_plot_g$layout[strips,],
          gtable_add_grob(inclusion_plot_g, new_grobs,
                          t=t, l=l, b=b, r=r, name="cards"))        

inclusion_plot_g$widths[[11]] <- unit(2.5,"cm")

grid.draw(inclusion_plot_g)
```

Figure \@ref(fig:correctivePlot) shows the same feedback data in \@ref(fig:feedbackFillPlot) but uses the x-axis to also show the proportion of feedback categories other than yes/no judgments. The goal here is to better see the trial types with corrective feedback (blue and red). These trial types include: 1. conjunction when only one conjunct is true, 2. disjunction when both disjuncts are true, 3. simple guesses (e.g. "there is a cat") when two animals are on the card. These trial types involve guesses that are either false or infelicitous.

Furthermore, the type of corrective feedback provided matches the type of mistakes made in the guesses. With conjunctive guesses like "cat and dog" when there is only one cat on the card, children provide exclusive corrections (e.g. just/only a cat), suggesting that the other animal should have been excluded. With disjunctive guesses like "cat or dog" and simple guesses like "cat" when there are two animals on the card, children provide inclusive feedback, indicating that another animal should have also been included. This is particularly notable in the case of disjunction since both animals are mentioned, but children still emphasize that the connective *AND* should have been used, or that *BOTH* mentioned animals are on the card.

```{r eval=FALSE}
judg_agree <-
  data_study3 %>%
  select(Animal, Guess, tvjt, judgment)
  
judg_agree$tvjt <- as.character(judg_agree$tvjt)

judg_agree$tvjt[is.na(judg_agree$tvjt)] <- "0"

judg_agree$judgment <- 
  fct_recode(judg_agree$judgment,Yes="Y", No="N", `0`="YN")

agree_table <-
  judg_agree %>%
  group_by(Animal,Guess) %>%
  select(tvjt, judgment) %>%
  agree()

total_agree <-
judg_agree %>%
  select(judgment, tvjt) %>% 
  agree()

total_kappa <-
judg_agree %>%
  select(judgment, tvjt) %>% 
  kappa2()
```

### Discussion

Study 3 measured children's comprehension of logical connectives by asking them to judge a puppet's guess in two ways: open-ended and two-alternative forced choice. First, we asked children to say "yes" to the puppet if he was right and "no" if he was wrong. However, we allowed children to provide any form of feedback they desired. Second, we followed children's open-ended feedback with a two-alternative forced choice question: "Was the puppet right?" This way, we measured children's comprehension in two different ways in the same trial. Ideally, both measures should show similar findings. We showed that the findings were similar with conjunctive guesses but not disjunctive ones. Children avoided a clear-cut binary right/wrong feedback with disjunction and preferred to provide more nuanced feedback when they were allowed.

The 2AFC responses followed an expected pattern: conjunctive guesses with only one conjunct true were considered wrong while those with both conjuncts true were judged right. Disjunctive guesses were judged right whether one or both disjuncts were true. We found no significant difference in the 2AFC task between the responses of children and those of adults in study 1.

Children's open-ended feedback in study 3 replicated those of study 2. We found that children provided more corrective feedback in false and infelicitous trials than true and felicitous ones. The corrective feedback was tailored to the puppet's mistake. If the puppet used a conjunction when only one animal was on the card, children pointed out that the other animal should have been excluded using the exclusive adverbials *just* and *only*. If the puppet used a disjunction when both animals were on the card, children stressed the connective *and* or the determiner "both", implying that both animals should be included. 

While the 2AFC results suggested that children take no issue with disjunctive guesses when both disjuncts are true, the analysis of their corrective feedback showed that children provide appropriate corrections in such cases and emphasize that the connective *and* would have provided a better guess. Taken both measures together, we can conclude that even though children are aware of the problem with such guesses, they do not consider them as *wrong*. This is similar to what we reported with adults in study 1. 

## General Discussion

<!-- Summary--->

We reported three studies on adults and four-year-old's comprehension of the logical connectives *and* and *or*. The first study used two- and three-alternative forced choice judgment tasks with adults. In the 2AFC task, adult interpretations closely matched the semantic accounts of *and* and *or* as conjunction and inclusive disjunction. The judgments did not register robust signs of pragmatic infelicities. However, the 3AFC judgment task, showed signs of pragmatic infelicities, especially in disjunctive guesses with true disjuncts. When both disjuncts were true, participants were more likely to choose "kinda right" rather than "right".

The second study used a 3AFC judgment task with four-your-old children. It also included an exploratory analysis of children's open-ended verbal feedback to the puppet in the experimental setting. Children's interpretations were similar to those of adults in the 3AFC task and only differed in one case: pragmatically infelicitous disjunctions. When both disjuncts were true, adults tended to judge disjunctive guesses as "kinda right". We took this to provide evidence for the pragmatic infelicity of such guesses. This was not the case for children as they judged such disjunctive statement as "right". However, our analysis of children's open-ended feedback showed that children took issue with such statements and provided appropriate corrective feedback.

In the third study, we focused on eliciting open-ended verbal feedback from children and followed it with a 2AFC task. In the 2AFC task, children's responses reflected the semantics of connectives as conjunction and inclusive disjunction. There was no significant difference between children and adults in two-alternative judgments. Since the 2AFC task proved to be a good indicator of semantic knowledge, it is reasonable to conclude that adults and children showed similar semantic knowledge of the connectives. The analysis of children's open-ended feedback replicated the findings in study 2. Children provided more corrective feedback in false and pragmatically infelicitous trials with logical connectives. The comparison of the 2AFC task and children's open-ended responses shows that children are sensitive to the infelicity of disjunctions with true disjuncts, even though they consider them to be "right" guesses.

Overall, we did not find major differences in adults' and four-year-old children's interpretation of logical connectives *and* and *or* in the context of our guessing game. However, there were two minor differences. First, we found that in both 2AFC and 3AFC judgment tasks, children showed a small preference for disjunctions with both disjuncts true rather than only one. Adults on the other hand showed the opposite pattern: they preferred disjuncts with only one disjunct true. Second, in both 2AFC and 3AFC judgment tasks, children rated disjunctions with both disjuncts true better than adults. That is, they considered utterances like "there is a cat or a dog" when both animals were on the card "right" more often than adults. Here we discuss these two differences and their potential causes in more detail.

###Preference for True Disjuncts {#conjunctive}

First, we found a small preference for both disjuncts being true rather than only one among some children. This effect is similar in nature to what @Singh2016 and @tieu2016 report but not in magnitude. In our study this effect is quite small while @Singh2016 and @tieu2016 seem to have found bigger effects. Based on this, @Singh2016 proposed that a large group of children at this age-range have a pragmatically driven conjunctive interpretation of disjunction. In short, due to a non-adult like alternative set to the connective *or*, children strengthen a disjunctive statement pragmatically and derive a conjunction. Our studies do not provide support for this proposal. In both 2AFC and 3AFC judgments, children clearly differentiated between disjunctive and conjunctive guesses. Furthermore, the analysis of children's open-ended feedback showed clearly different response patterns for conjunction and disjunction. More importantly, the open-ended feedback in disjunctive guesses showed the opposite pattern to what the conjunctive hypothesis predicts: children took issue with disjunctions that had both disjuncts true and provided more corrective feedback in such cases. Therefore, it is possible that this preference is a product of forced choice tasks rather than children's comprehension of connectives.

However, even assuming that this small preference for true disjuncts is not due to the measurement method, it can be accounted for by several other hypotheses that have not been successfully ruled out yet. First, the conjunctive interpretation may not be due to a faulty pragmatic computation, but rather a default conjunctive interpretation when the connective is not properly heard/understood or is unknown. To check this hypothesis, it may be possible to test children's comprehension of novel or noisy connectives. A novel coordination like *cat dax dog* with *dax* as a nonce connective may be interpreted as a conjunction. Such results would suggest that in studies with high cognitive demand, children may default to a conjunctive interpretation if they miss the relevant connective. Second, the conjunctive preference may be due to some children's preference for the linguistic labels to match the animals on the card (or more generally linguistic description and state of the world). This hypothesis is consistent with the results in the other trial type that had mismatch in the number of animals and the guess while the guess was technically true: simple guesses (e.g. there is a cat) when two animal (e.g. cat and dog) were present on the card. Children were equally split between "wrong" and "right" in their judgments while adults considered such guesses as "right". In light of these alternative explanations, we are hesitant to attribute this small preference to a pragmatically driven conjunctive interpretation of disjunction.

###Lack of Infelcity with True Disjuncts 

The second minor difference between adults and children was in the 3AFC judgment task: in disjunctive trials (e.g. "cat or dog") with two animals (e.g. cat and dog), adults were more likely to choose "kinda right" than children. Children mostly chose "right". This response pattern is often taken to mean that children find no infelicity with such disjunctions or that they do not "derive an exclusivity implicature". This lack of infelicity/implicature is consistent with the generalization in the literature on scalar items: that children are more likely than adults to interpret terms literally  and that children do not compute implicatures or judge infelicity to the same **rate** that adults do [@pouscoulous2009going, @katsos2014scalar]. But why is that?

There have been three major proposals to account for children's low rate of implicatures: 1. processing [@reinhart2004processing, @pouscoulous2007developmental] 2. non-adult-like lexical entry (@barner2011accessing, @horowitz2017trouble) and 3. pragmatic tolerance [@katsos2011pragmatic]. Here we show that none of these accounts provide a satisfactory explanation of the results in this study.  

**1. Processing Account** First, the processing accounts locate the issue in children's processing capacities such as working memory. They suggest that pragmatic computations are cognitively taxing and children lack the appropriate processing resources to carry them out appropriately. A prediction of processing accounts (at least in their current format) is that children will show reduced implicature computations for all types of implicatures - scalar or ad-hoc. This prediction is not borne out in our experimental results. In study 3, children were much more likely than adults to call a simple guess (e.g. "cat") "wrong" in the 2AFC task if there were two animals on the card (e.g. "cat and dog"). Processing accounts would not predict that children may derive more implicatures than adults but this is what we find if we carry on with the traditional interpretation of the judgment task.

**2. Non-adult Like Lexicon** Several proposals blame the structure of the child's lexicon for the alleged failure in deriving implicatures/infelicity. The assumption is that the child's lexical entry for scalar items must include three elements for successful derivation: 1. the semantics of the weak term (e.g. *some*, *or*) 2. the semantics of the strong term (e.g. *all*,*and*); and possibly 3. a scale that recognizes the stronger term as an alternative to the weaker one (e.g. <*some*, *all*>, <*or*, *and*>). Each of these elements have been pinpointed as the source of the problem in previous studies [@horowitz2017trouble, @katsos2011pragmatic, @barner2011accessing]. However none of them seem to apply to the results reported here.

If children in this study lack the semantics of the connective *or*, we can expect them to either perform at chance or default to a conjunctive interpretation. Neither prediction is borne out in study 2 and 3. Furthermore, children's free-form linguistic feedback suggests good understanding of disjunction. So this explanation seems unlikely. The problem cannot be that children do not know the meaning of *and* either. Children's performances in both study 2 and 3 for conjunction trials show that they understand its meaning very well. Finally, while it is possible that children lacked the appropriate lexical scale and could not access the stronger alternative, this explanation cannot be the whole story. Several children in both study 2 and 3 stressed the word *and* suggesting that the puppet should have used the stronger term instead. However, they judged the puppet's guess as "right". If children could not access the stronger term, they could not mention it in their feedback either.

**3. Pragmatic Tolerance** Katsos \& Bishop [-@katsos2011pragmatic] suggested that children tend to tolerate pragmatic infelicities more than adults. They showed that when children are provided with a 2AFC judgment task, they tolerate the infelicity of *some* when *all* applies but when they are presented with a 3AFC task they choose the middle option and report this infelicity. Similar to the processing account, the pragmatic tolerance account predicts that scalar and ad-hoc implicatures are similarly affected. However, we did not find similar results to that of Katsos \& Bishop [-@katsos2011pragmatic]. When children were presented with a 3AFC task, they chose the highest reward (and not the middle option) for uses of *or* when *and* applied. Second, and more importantly, we found different patterns for ad-hoc and scalar implicatures as mentioned before. This is not predicted by the tolerance account unless we assume that children are more tolerant towards violations of scalar implicatures than ad-hoc ones. While tolerance may not be the source of the problem here, we believe that a number of discussions including Katsos \& Bishop [-@katsos2011pragmatic] and later Katsos [-@katsos2014scalar] point to an important factor: the role of measurement in our estimates of children's pragmatic capacity. 

We believe several observations in our studies here provide support for the hypothesis that methodological issues, and more specifically issues of measurement contribute to the differences found between adults and children in pragmatic capacities. First, study 1 showed that even for adults, our estimates of adult infelicity rates may differ based on the number of alternatives in the forced choice task. A 2AFC task tends to underestimate adults' rates of pragmatic infelicity. Second, children's open-ended linguistic feedback in the experimental context better reflected their sensitivity to pragmatic nuances than the forced-choice judgment tasks. Third, children showed a higher rate of infelicity judgments for cases of ad-hoc implicatures (simple guesses with two animals on the card) than adults. While a difference in sensitivity to ad-hoc vs. scalar implicatures has been reported and argued for before [@stiller2015ad; @horowitz2017trouble], a higher sensitivity than adults is not predicted by any of the current accounts. 

In order to better understand the differences between adults and children's pragmatic capacities, we need to first have a good understanding of how our measurements affect our estimates of adults and children's performance in our experimental tasks. It is quite possible that children are not more capable of making exhaustive inferences than adults and not less capable of making scalar inferences either. They simply have different construal of the wrong-right scale and what the forced-choice task is about. The concepts "right" and "wrong" are as much subject to developmental change and differences between adults and children as scalar items that we are interested in. It is reasonable to assume that children's understanding of what constitutes as "right" or "wrong" does not conform to that of adults. However, we do not have a good understanding of what these differences are and how they affect our estimates of children's pragmatic abilities. It is important to point out that such issues of measurement can be the culprit behind both children's seemingly slight preference for true disjuncts explained earlier and their lack of infelicity when both disjuncts are true.

**A General Approach for Measuring Implicature/Infelicity Rate**

Methodological issues are nothing new in studies of children's semantic and pragmatic development. Developing the best measures of children's linguistic capacities has always been a major concern for researchers in the field. Our goal here is to propose some future steps that can address the methodological concerns in children's pragmatic development.

As Pouscoulous \& Noveck [-@pouscoulous2009going] and Katsos [-@katsos2014scalar] explain, the central issue is **the rate** at which children and adults manifest pragmatic reasoning in the experimental setting. No one doubts children's capacity to perform such computations at all. At issue is the extent to which children and adults compute implicatures. And the claim is that children perform such computations less often than adults; or that children do not perform such computations when adults normally would. In the previous section, we discussed the proposed factors in the literature that can cause this difference including processing demands, the structure of the lexicon, tolerance, as well as the way we measure adults and children's comprehension. As Katsos [-@katsos2014scalar] points out, it is reasonable to assume that all these factors play a part to some degree. What matters is the degree to which each factor contributes to the outcome. 

```{r implicatureGraph, fig.env = "figure", fig.pos = "t", fig.align='center', fig.width=5, fig.height=5, set.cap.width=T, num.cols.cap=1, fig.cap = "Factors that affect pragmatic computations and our estimates of these computations in the experimental setting."}
implicature_img <- jpeg::readJPEG("figs/implicature.jpg")
grid::grid.raster(implicature_img)
```

Figure \@ref(fig:implicatureGraph) shows the factors that affect pragmatic computations as well as our observations of the rate of pragmatic computations in an experiment. First it is important to distinguish between factors that affect pragmatic computations and those that affect our observed rate in the experimental setting. As we showed in experiment 1, given the number of alternatives in the forced choice task (2AFC vs. 3AFC), we may get different estimates of adults' infelicity judgments, yet it is unreasonable to assume that there is a difference in adults' pragmatic capacities in these two tasks. A similar situation exists when we compare children's forced choice measures of infelicity and their open-ended feedback. In disjunctive trials where both disjuncts are true, the forced choice tasks show no sign of infelicity while the open ended responses show that children are sensitive to the infelicity of disjunction when a conjunction would have been more appropriate.  

## Conclusion

To conclude, we showed that children and adults do not differ substantially in their **semantic** knowledge of the logical connectives. The results were highly consistent with the current accounts that posit the semantics of *and* as logical conjunction and *or* as logical (inclusive) disjunction. With respect to pragmatic knowledge, our three-alternative forced choice judgment task showed that adults are sensitive to the infelicity of disjunctive statements when both disjuncts are true. We showed that while the three-alternative judgment task failed to register such a sensitivity for children, our systematic analysis of children's verbal open-ended feedback showed that children are sensitive to pragmatic infelicities and can provide appropriate corrections to infelicitous utterances with logical connectives.


<!-- * In both binary and ternary judgment task, children do not show any sign of an exclusivity inference. There can be two explanations: 1. children's pragmatic comeptence is not developed. 2. they don't have access to "and" as an alternative. 3. children and adults differ on what they count as "wrong" or "right".
    + The first explanation does not work. Children show sensitivity to the vilation of an exhaustive reading of a simple guess even more than adults.
    + The second does not work because children explicitly mentioned "and" as the alternative.
    + There is tentative evidence to suggest the last option may be true. what are those?

To our knowldge the issue of measurment has remained under-explored. It is possible that the researcher's methods of measurement are an important contributor to this effect. Reasons 1. study 1 binary and ternary judgments as well as study 2 and 3 differences between the judgment tasks and the analysis of children's open-ended verbal feedback. 

Truth value judgment tasks rely on the participants' understanding of "right" and "wrong" to measure the semantics or pragmatics of a word of interest such as "and" and "or". In other words, we are using a concept to measure another. Our measurements are going to be as accurate as the tool that we use. 

It is reasonable to doubt that children and adults have a similar understanding of what constitutes as "right" or "wrong" in a given context.

  + To our knowledge, no study has systematically investigated issues in measuring implicatures in experimental settings. The working assumption has been that adults and children do not differ in what they consider *right* or *wrong* in a judgment task. Our data suggests otherwise.
  
  + For adults, the forced-choice binary task seems to underestimate the adult pragmatic inferences and reflect semantic knowledge better. We see that the forced-choice ternary task does a better job of reflecting some of the pragmatic knowledge. Yet, even the forced choice ternary tasks leaves a lot out. For children, both forced-choice measures did not reflect sensitivity to the infelicity of *or* but open-ended feedback did.
 
 Children did not only show a non-adult like pattern in the case of scalar inferences as percieved in the field. they also showed non-adult like behavior with respect to the ad-hoc ones. and in fact in the opposite direction. This is predicted if we assume that children's understanding of "right" and "wrong" is not adult like and will differ substantially in edge-cases, pragmatically loaded guesses.
 
  * Second, in study 3 we saw that even though x number of children explicitly correct the puppet when the puppet uses "or" instead of "and", they continue to consider the puppet to be "right". This is a strong evidence suggesting that children do not consider such violations as "wrong" even though they are sensitive to them and understand that "and" would have been the better one to use.

 Forced choice experimental tasks rely on a linking account of how theoretical concepts such as "truth" or "felicity" would map into intuitive concepts such as being "right" or "wrong". Most accounts assume that the linking account remains relatively constant not only among adult participants, but rather between children and adults. In other words, what adults consider right and wrong in linguistic descriptions, children would also find right or wrong. Our study provides tentative evidence against this assumption and suggests that serious investigation is required before we can comfortably proceed with it. We plan to test this hypothesis more systematically in a future study.

-->
