---
title: "Connective_Game_Analysis"
author: "Masoud Jasbi"
date: "August 18, 2016"
output: html_document
---
  
```{r packages, include=FALSE}
require(tidyverse)
require(ggplot2)
library(forcats)
library(rstan)
library(ggmcmc)
library(ggthemes)
## reading the processed data
data_main <- read.csv("../2_processed_data/cardgame_data_processed.csv")

## isolating the trials with 3 values as DV
data_main_dv3 <- data_main %>%
  filter(dv_type==3)
```

```{r rstan-data}
df_stan_adult <- as_tibble(data_main_dv3) %>%
  filter(age_category == "Adults", dv_type== 3) %>%
  select(-age_category, -dv_type, -age_year, -feedback) %>%
  mutate(response = fct_relevel(response, "Wrong","Kinda Right")) %>%
  filter(!is.na(response), 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY")) %>%
  mutate(sid = factor(sid), 
         trial_type = fct_relevel(factor(trial_type), "XY.XandY"))
```

```{r rstan-model}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // one-animal conjunction. two-animals-conjunction is the intercept
  int x_or[N]; // one-animal disjunction
  int xy_or[N]; // two-animals disjunction
}

parameters {
  vector[J] sub_intercept; //subject intercepts
  real alpha; // 
  real beta_x_and; // one-animal 
  real beta_x_or; // 
  real beta_xy_or; // 
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);

  c ~ normal(0, 1);

  for (j in 1:J)
    sub_intercept[j] ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + sub_intercept[subj[n]], c);
}
"

stan_data <- list(subj = as.integer(df_stan_adult$sid),
                  response = as.integer(df_stan_adult$response),
                  x_and = as.integer(df_stan_adult$trial_type == "X.XandY"),
                  x_or = as.integer(df_stan_adult$trial_type == "X.XorY"),
                  xy_or = as.integer(df_stan_adult$trial_type == "XY.XorY"),
                  N = nrow(df_stan_adult),
                  J = length(unique(df_stan_adult$sid)),
                  K = 3)

cardgame_stan <- stan(model_code = cardgame_model,
                       data = stan_data, 
                       warmup = 1000,
                       iter = 2000, chains = 4)
```

Extract the betas:

```{r betas}
stanadult_samples <- ggs(cardgame_stan) 

adult_model <- filter(stanadult_samples, 
                 Parameter %in% c("alpha","beta_x_or",
                                  "beta_x_and","beta_xy_or"))

adult_model %>%  
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99)) %>%
  knitr::kable(digits = 2)

adultmodel_plot <- adult_model %>%
  spread(Parameter, value) %>%
  mutate(and_mean = alpha - beta_x_and,
         x_mean = beta_x_and - beta_x_or, 
         or_mean = beta_x_or - beta_xy_or) %>%
  select(-alpha, -beta_x_and, -beta_x_or) %>%
  gather(Parameter, value, and_mean, x_mean, or_mean, beta_xy_or) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99)) %>%
  knitr::kable(digits = 2)

  ggplot(adultmodel_plot, aes(x=Parameter,y=mean)) + geom_point() + geom_errorbar(aes(ymin=hdp_lower,ymax=hdp_upper), position = position_dodge(.6), width=0.2)

adultmodel_plot_2 <- 
  adult_model %>%
  spread(Parameter, value) %>%
  gather(Parameter, value, alpha, beta_x_and, beta_x_or, beta_xy_or) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99))

  ggplot(adultmodel_plot_2, aes(x=Parameter,y=mean)) + geom_point() + geom_errorbar(aes(ymin=hdp_lower,ymax=hdp_upper), position = position_dodge(.6), width=0.2) + theme_few()


```

Does intercept trade off with cutpoints?

```{r}
ggmcmc:ggs_pairs(cardgame_samples)

```

```{r}
df_stan_adult %>%
  group_by(trial_type, response) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```

Explore chains

```{r}

ggmcmc(filter(cardgame_samples, !str_detect(as.character(Parameter), "sub_intercept")))
```

Old age model. 

```{r}
cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  real age_year[N];
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // xy and is the intercept
  int x_or[N];
  int xy_or[N];
}

parameters {
  // vector[3] beta; // fixed intercept and slopes
  // vector[J] u; //subject intercepts
  // real<lower=0> sigma_e; //error sd
  // real<lower=0> sigma_u; //subj sd
  real alpha;
  real beta_age;
  real beta_x_and;
  real beta_x_or;
  real beta_xy_or;
  real beta_age_x_and;
  real beta_age_x_or;
  real beta_age_xy_or;
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_age ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);
  beta_age_x_and ~ normal(0, 10);
  beta_age_x_or ~ normal(0, 10);
  beta_age_xy_or ~ normal(0, 10);

  c ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + beta_age * age_year[n] + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + 
        beta_age_x_and * age_year[n] * x_and[n] + 
        beta_age_x_or * age_year[n] * x_or[n] + 
        beta_age_xy_or * age_year[n] * xy_or[n], c);
}
"
```

To test this we look at the difference in paramter estimate iteration by iteration, the 95% confidence intervals for each iteration. 95% confidence interval on the difference between the two.
Report the means for the three betas with the confidence intervals and then a table for the differences with their confidence intervals
'


