---
title: "Connective_Game_Analysis"
author: "Masoud Jasbi"
date: "August 18, 2016"
output: html_document
---
  
```{r packages, include=FALSE}
require(tidyverse)
require(ggplot2)
library(forcats)
library(rstan)
library(ggmcmc)
```

```{r raw-data}
## reading the processed data
data_main <- read.csv("../2_processed_data/cardgame_data_processed.csv")

## isolating the trials with 3 values as DV
data_main_dv3 <- data_main %>%
  filter(dv_type==3)
```

```{r rstan-child-data}
df_stan <- as_tibble(data_main_dv3) %>%
  filter(age_category == "Children") %>%
  select(-age_category, -dv_type) %>%
  mutate(response = fct_relevel(response, "Wrong","Kinda Right"), 
         age_centered_years = as.numeric(scale(age_year, scale=FALSE))) %>%
  filter(!is.na(response), 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY")) %>%
  mutate(sid = factor(sid), 
         trial_type = fct_relevel(factor(trial_type), "XY.XandY"))
```

```{r rstan-adult-data}
df_stan_adult <- as_tibble(data_main_dv3) %>%
  filter(age_category == "Adults", dv_type== 3) %>%
  select(-age_category, -dv_type, -age_year, -feedback) %>%
  mutate(response = fct_relevel(response, "Wrong","Kinda Right")) %>%
  filter(!is.na(response), 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY")) %>%
  mutate(sid = factor(sid), 
         trial_type = fct_relevel(factor(trial_type), "XY.XandY"))
```

```{r rstan-model}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // one-animal conjunction. two-animals-conjunction is the intercept
  int x_or[N]; // one-animal disjunction
  int xy_or[N]; // two-animals disjunction
}

parameters {
  vector[J] sub_intercept; //subject intercepts
  real alpha; // 
  real beta_x_and; // one-animal 
  real beta_x_or; // 
  real beta_xy_or; // 
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);

  c ~ normal(0, 1);

  for (j in 1:J)
    sub_intercept[j] ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + sub_intercept[subj[n]], c);
}
"
```

```{r chid-sampling}
stan_childata <- list(subj = as.integer(df_stan$sid),
                  response = as.integer(df_stan$response),
                  x_and = as.integer(df_stan$trial_type == "X.XandY"),
                  x_or = as.integer(df_stan$trial_type == "X.XorY"),
                  xy_or = as.integer(df_stan$trial_type == "XY.XorY"),
                  N = nrow(df_stan),
                  J = length(unique(df_stan$sid)),
                  K = 3)

childmodel_stan <- stan(model_code = cardgame_model,
                       data = stan_childata, 
                       warmup = 1500,
                       iter = 3000, chains = 4)

child_samples <- ggs(childmodel_stan)
write_csv(child_samples, "child_samples")
```

```{r}
stan_adultdata <- list(subj = as.integer(df_stan_adult$sid),
                  response = as.integer(df_stan_adult$response),
                  x_and = as.integer(df_stan_adult$trial_type == "X.XandY"),
                  x_or = as.integer(df_stan_adult$trial_type == "X.XorY"),
                  xy_or = as.integer(df_stan_adult$trial_type == "XY.XorY"),
                  N = nrow(df_stan_adult),
                  J = length(unique(df_stan_adult$sid)),
                  K = 3)

adultmodel_stan <- stan(model_code = cardgame_model,
                       data = stan_adultdata, 
                       warmup = 1500,
                       iter = 3000, chains = 4)

adult_samples <- ggs(adultmodel_stan)
write_csv(adult_samples, "adult_samples")
```

Extract the betas:

```{r betas}
#adult_samples <- read_csv("adult_samples")

fixefs <- filter(child_samples, 
                 Parameter %in% c("alpha","beta_x_or",
                                  "beta_x_and","beta_xy_or"))

sumtab <- fixefs %>%
  spread(Parameter, value) %>%
  mutate(and_mean = alpha - beta_x_and,
         x_mean = beta_x_and - beta_x_or, 
         or_mean = beta_x_or - beta_xy_or) %>%
  select(-alpha, -beta_x_and, -beta_x_or) %>%
  gather(Parameter, value, and_mean, x_mean, or_mean, beta_xy_or) %>%
  group_by(Parameter) %>%
  summarise(mean = mean(value), 
            hdp_lower = quantile(value, .01),
            hdp_upper = quantile(value, .99)) 

  ggplot(sumtab, aes(x=Parameter,y=mean)) + geom_point() + geom_errorbar(aes(ymin=hdp_lower,ymax=hdp_upper), width=0.2)
```

Does intercept trade off with cutpoints?

```{r}
ggmcmc:ggs_pairs(cardgame_samples)

```

```{r}
df_stan %>%
  group_by(trial_type, response) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n))
```

Explore chains

```{r}

ggmcmc(filter(cardgame_samples, !str_detect(as.character(Parameter), "sub_intercept")))
```

Old age model. 

```{r}
cardgame_model <- "
data {
  int K; // number of responses
  int<lower=1> N; // number of trials
  int<lower=1> J; // number of subs
  real age_year[N];
  int<lower=1, upper=3> response[N];
  int<lower=1, upper=J> subj[N];
  int x_and[N]; // xy and is the intercept
  int x_or[N];
  int xy_or[N];
}

parameters {
  // vector[3] beta; // fixed intercept and slopes
  // vector[J] u; //subject intercepts
  // real<lower=0> sigma_e; //error sd
  // real<lower=0> sigma_u; //subj sd
  real alpha;
  real beta_age;
  real beta_x_and;
  real beta_x_or;
  real beta_xy_or;
  real beta_age_x_and;
  real beta_age_x_or;
  real beta_age_xy_or;
  ordered[K-1] c; // cutpoints
}

model {
  alpha ~ normal(0, 10);
  beta_age ~ normal(0, 10);
  beta_x_and ~ normal(0, 10);
  beta_x_or ~ normal(0, 10);
  beta_xy_or ~ normal(0, 10);
  beta_age_x_and ~ normal(0, 10);
  beta_age_x_or ~ normal(0, 10);
  beta_age_xy_or ~ normal(0, 10);

  c ~ normal(0, 10);

  for (n in 1:N)
    response[n] ~ ordered_logistic(alpha + beta_age * age_year[n] + 
        beta_x_and * x_and[n] + beta_x_or * x_or[n] + beta_xy_or * xy_or[n] + 
        beta_age_x_and * age_year[n] * x_and[n] + 
        beta_age_x_or * age_year[n] * x_or[n] + 
        beta_age_xy_or * age_year[n] * xy_or[n], c);
}
"
```

To test this we look at the difference in paramter estimate iteration by iteration, the 95% confidence intervals for each iteration. 95% confidence interval on the difference between the two.
Report the means for the three betas with the confidence intervals and then a table for the differences with their confidence intervals
'

```{r chisq-test}
criticals <- 
  data_main_dv3 %>% filter(!is.na(response), age_category == "Children", 
         trial_type %in% c("X.XorY","X.XandY","XY.XorY","XY.XandY"))
criticals$trial_type <- factor(criticals$trial_type)

ctable <-
  criticals %>% select(feedback, trial_type) %>%
  group_by(trial_type, feedback) %>%
  summarise (counts = n()) %>%
  spread(trial_type, counts)
ctable <- as.data.frame(ctable)

chisq_1andor <- chisq.test(ctable[,2], p= ctable[,3], rescale.p = T)
chisq_and12 <- chisq.test(ctable[,2], p = ctable[,4], rescale.p = T)
chisq_1and2or <- chisq.test(ctable[,2], p = ctable[,5], rescale.p = T)
chisq_1or2and <- chisq.test(ctable[,3], p= ctable[,4], rescale.p = T)
chisq_or12 <- chisq.test(ctable[,3], ctable[,5], rescale.p = T)
chisq_2andor <- chisq.test(ctable[,4], p=ctable[,5], rescale.p = T)

```

